### 2015-10-08 manuscript/acronyms.tex
move acronyms to their own file

\acrodef{EV}[EV]{Extracellular Vesicle}
\acrodef{SVM}{Support Vector Machine}
\acrodef{ROC}{Receiver Operating Characteristic}
\acrodef{KNN}{K Nearest Neighbors}
\acrodef{SRA}{Sequence Read Archive}
\acrodef{PRG2}[\textit{PRG2}]{Proteoglycan 2, bone marrow}
\acrodef{FPKM}{Fragments Per Kilobase of exon per Million reads}
\acrodef{CSH1}[\textit{CSH1}]{chorionic somatomammotropin hormone 1}
\acrodef{CSH2}[\textit{CSH2}]{chorionic somatomammotropin hormone 2}
\acrodef{CSHL1}[\textit{CSHL1}]{chorionic somatomammotropin hormone-like 1}
\acrodef{GH2}[\textit{GH2}]{growth hormone 2}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manuscript.Rnw"
%%% End: