### 2016-01-12 manuscript/acronyms.tex
add GH acronym

\acrodef{EV}[EV]{Extracellular Vesicle}
\acrodef{SVM}{Support Vector Machine}
\acrodef{ROC}{Receiver Operating Characteristic}
\acrodef{KNN}{K Nearest Neighbors}
\acrodef{SRA}{Sequence Read Archive}
\acrodef{PRG2}[\textit{PRG2}]{Proteoglycan 2, bone marrow}
\acrodef{FPKM}{Fragments Per Kilobase of exon per Million reads}
\acrodef{CSH1}[\textit{CSH1}]{chorionic somatomammotropin hormone 1}
\acrodef{CSH2}[\textit{CSH2}]{chorionic somatomammotropin hormone 2}
\acrodef{CSHL1}[\textit{CSHL1}]{chorionic somatomammotropin hormone-like 1}
\acrodef{GH2}[\textit{GH2}]{growth hormone 2}
\acrodef{CEA}{carcinoembryonic antigen}
\acrodef{HSD3B1}[\textit{HSD3B1}]{3-β Hydroxysteroid dehydrogenase 1}
\acrodef{GSE}{Gene Set Enrichment}
\acrodef{OU}{Ornstein-Uhlenbeck}
\acrodef{OUCH}{\ac{OU} Comparative Hypotheses}
\acrodef{TSI}{Tissue Specificity Index}
\acrodef{SRA}{Sequence Read Archive}
\acrodef{GH}{growth hormone}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manuscript.Rnw"
%%% End: