### 2016-02-21 presentation/placenta_viviparous_presentation.Rnw
add presentation

\documentclass[ignorenonframetext]{beamer}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage{qrcode}
\usepackage{cleveref}
\usepackage[absolute,overlay]{textpos}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}

\usepackage{texshade}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usepackage{xspace}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\renewcommand*{\bibfont}{\tiny}
\usepackage{mathtools,cancel}

\renewcommand{\CancelColor}{\color{red}} %change cancel color to red

\makeatletter
\let\my@cancelto\cancelto %copy over the original cancelto command
\newcommand<>{\cancelto}[2]{\alt#3{\my@cancelto{#1}{#2}}{\mathrlap{#2}\phantom{\my@cancelto{#1}{#2}}}}
% redefine the cancelto command, using \phantom to assure that the
% result doesn't wiggle up and down with and without the arrow
\makeatother

\usepackage{multirow}
\usepackage{array}

\mode<presentation>{ 
  \usetheme{CambridgeUS}
  \usecolortheme{crane}
%   % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
  \definecolor{ilboldblue}{HTML}{002058}
  \definecolor{ilboldorange}{HTML}{E87722}
  \definecolor{ilblue}{HTML}{606EB2}
  \definecolor{ilorange}{HTML}{D45D00}
%   \setbeamercolor{alerted text}{fg=ilboldorange}
%   \setbeamercolor*{palette primary}{fg=ilboldblue,bg=ilorange!70!white}
%   \setbeamercolor*{palette secondary}{fg=ilboldblue,bg=ilorange!60!white}
%   \setbeamercolor*{palette tertiary}{bg=ilblue!60!white,fg=ilboldorange}
%   \setbeamercolor*{palette quaternary}{fg=ilboldblue,bg=ilorange!80!white}
%   
%   \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}
%   
%   \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
%   \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
%   \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
%   \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}
%   
%   % \setbeamercolor*{titlelike}{parent=palette primary}
%   \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
%   \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!20!white}
%   \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}
%   
%   \setbeamercolor*{separation line}{}
%   \setbeamercolor*{fine separation line}{}
%   \setbeamercovered{transparent}  
  \logo{\begin{tikzpicture}% Pale figure
      {\node[opacity=0.734]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
        };}%
    \end{tikzpicture}}
}

\title[Placenta Gene Expression]{The evolution of gene expression in the term placenta of viviparous mammals}
\author[Don L Armstrong \textit{et al.}]{Don L Armstrong, Michael R.
  McGowen, Amy Weckle, Jason Caravas, Dalen Agnew, Kurt Benirschke,
  Sue Savage-Rumbaugh, Eviatar Nevo, Chong J Kim, Günter P. Wagner,
  Roberto Romero, and Derek E. Wildman}
\institute[IGB]{Institute for Genomic Biology, Computing Genomes 
  for Reproductive Health, University of Illinois, Urbana-Champaign}
\subject{Placenta Expression}

\date{\today}
\renewcommand\cite[1]{\autocite{#1}}
% \pagestyle{fancy}

% species names
\newcommand{\smonkey}[0]{\textit{A. fusciceps}\xspace}
\newcommand{\cow}[0]{\textit{B. taurus}\xspace}
\newcommand{\dog}[0]{\textit{C. familiaris}\xspace}
\newcommand{\armadillo}[0]{\textit{D. novemcinctus}\xspace}
\newcommand{\horse}[0]{\textit{E. caballus}\xspace}
\newcommand{\elephant}[0]{\textit{L. africana}\xspace}
\newcommand{\human}[0]{\textit{H. sapiens}\xspace}
\newcommand{\opossum}[0]{\textit{M. domestica}\xspace}
\newcommand{\mouse}[0]{\textit{M. musculus}\xspace}
\newcommand{\galili}[0]{\textit{N. galili}\xspace}
\newcommand{\sheep}[0]{\textit{O. aries}\xspace}
\newcommand{\bonobo}[0]{\textit{P. paniscus}\xspace}
\newcommand{\carmeli}[0]{\textit{S. carmeli}\xspace}
\newcommand{\pig}[0]{\textit{S. scrofa}\xspace}

% top 50 gene table explanation. Note that if you modify this, you
% should probably also modify the human table explanation, which is
% slightly different.
\newcommand{\tableexplanation}[2]{The top 50 genes by \ac{FPKM} in #1
  which are not annotated as Mt\_rRNA, Mt\_tRNA, rRNA or whose \human
  ortholog is a housekeeping gene. Gene symbol is the gene symbol or
  Ensembl gene id. Human Symbol is the 1:1 human ortholog or the best
  human alignment when enclosed in parenthesis. Human Name is the HUGO
  name corresponding to the human symbol. A complete list of all genes
  is in \cref{#2}.}

\newcommand{\genetreeexplanation}[1]{Gene tree and heatmap of
  expression of genes related to #1. The gene tree was obtained from
  Ensembl~\cite{Cunningham.ea2015:Ensembl2015} and genes from species
  not included in this study or having expression by \ac{FPKM} less
  than 1 were excluded from the plotted tree. Heatmap cells in white
  are from species where there was no expression of the relevant
  transcript (or a species where that transcript does not apply). The
  species names are abbreviated with the first letter of the genus and
  first two letters of the species (for example, \textit{Homo sapiens}
  is \textit{H.sa.}). Genes with duplicate gene names in the same
  species (or no gene named defined) are listed by the Ensemble gene
  ID.}

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.8\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 0)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")
library("scales")
library("magrittr")
library("tidyr")

@

<<source.to.latex,echo=FALSE,message=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R")))>>=
for (a in paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R"))) {
    source(a)
}
@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
setkey(hugo.nomenclature,"Approved Symbol")
@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

@

<<plot.defaults,echo=FALSE,results="hide">>=
grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)

@

\IfFileExists{figures/relevant_xkcd.png}{\frame[plain]{\centering \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{figures/relevant_xkcd.png}

    \url{https://xkcd.com/1605/}}}

\frame[plain]{\titlepage}

\section{Background}

\begin{frame}{Function and Evolution of Placenta}
\begin{itemize}
\item Placenta functions in the transfer of nutrients, oxygen, and
  waste between mother and offspring
\item Placenta is highly variable structure and form in mammals, even though major
  function is conserved
\item Placenta likely arose before the origin of therian mammals
\item Form relatively well characterized, but diversity of molecular
  environment of the placenta is only recently being understood
\item Possible role of evolutionary interplay between maternal and
  fetal strategies
\end{itemize}
\end{frame}

\begin{frame}{Impacts on Human Health?}
\begin{itemize}
\item Humans possess a deeply invasive trophoblast layer that can
  penetrate the myometrium of the uterus
  \cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
\item Defects in placental growth can lead to pre-eclampsia, which can
  cause hypertension, proteinuria, and mortality of mother and/or
  infant \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
\item Pre-eclampsia has been seen in \textit{Pan troglodytes}
\cite{Stout.ea1969:Glomerularcapillaryendothelialswelling}, and
\textit{Gorilla gorilla}
\cite{Thornton.ea1992:Convulsionsinpregnancyin}
\item Far higher incidence in \textit{Homo sapiens}, with as many as 
  4\% of pregnancies United States affected
  \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
\end{itemize}
\end{frame}

\begin{frame}{Overall Hypotheses}
\begin{itemize}
\item Gene expression patterns correlate with placenta morphology
\item Altered expression of genes correlates with invasiveness, and
  therefore pre-eclampsia
\item Maternal/Fetal evolutionary strategies will be reflected in gene
  expression patterns
\end{itemize}
\end{frame}

\begin{frame}{Species Studied}
\begin{figure}
<<species.tree,fig.width=8,fig.height=6,echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Nannospalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.5,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.5,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.5,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.5,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.5,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.5,node=6,width=0.12)
      ## sheep
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.5,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.5,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.5,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.5,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.5,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.5,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.5,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.5,node=13,width=0.18)
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{Variation in Gestational Periods and Litter Size}
\tiny
\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Gestation (days) & Litter Size  & Source \\
    \midrule
    \textit{Ateles fusciceps}     & 224 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Bos taurus}           & 278 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Canis familiaris}     & 61 & 4 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Dasypus novemcinctus} & 150 & 4 & \citet{Smith1984:amazingarmadillo} \\
    \textit{Equus caballus}       & 336 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation}\\
    \textit{Loxodonta africana}   & 644 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Homo sapiens}         & 270 & 1.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Monodelphis domestica}& 15 & 7.1 & \citet{Harder.ea1993:Gestationandplacentationin} \\
    \textit{Mus musculus}         & 20 & 6.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Nannospalax galili}   & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals} \\
    \textit{Ovis aries}           & 151 & 1.35 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Pan paniscus}         & 223 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental} \\
    \textit{Spalax carmeli}       & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals}\\
    \textit{Sus scrofa}           & 114 & 5.8 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \bottomrule
  \end{tabular}
%  \caption{Gestational periods in days and average liter sizes for studied species.}
%  \label{tab:gestational_periods}
\end{table}
\end{frame}

\begin{frame}{Variations in Placenta Morphology}
\begin{table}
\begin{tabular}{c c c}
  \toprule
  Species & Interhemal Membrane & Pattern \\
  \midrule
  \smonkey & hemochorial & trabecular \\
  \cow     & epitheliochorial & villous \\
  \dog     & endotheliochorial & labyrinthine \\
  \armadillo & hemochorial & trabecular \\
  \horse & epitheliochorial & villous \\
  \elephant & endotheliochorial & labyrinthine \\
  \human & hemochorial & villous \\
% http://onlinelibrary.wiley.com/doi/10.1002/jez.1402660511/pdf
  \opossum & epitheliochorial & choriovitelline \\
  \mouse & hemotrichorial & labyrinthine \\
  \galili & hemotrichorial & labyrinthine \\
  \sheep & epitheliochorial & villous \\
  \bonobo & hemochorial & villous \\
% http://www.biolreprod.org/content/early/2014/10/17/biolreprod.114.124818.short
  \carmeli & hemotrichorial & labyrinthine \\
  \pig & epitheliochorial & diffuse \\
  \bottomrule
\end{tabular}
\end{table}
\end{frame}

\section{Methods}

\subsection{Tissue/Extraction/Sequencing}

\begin{frame}{Collection of Placental Tissue}
\begin{itemize}
\item Collected from 8 mammalian species
\item Only fetal portion was sampled
\item Sequenced using GAII with an insert size of 230 bp
\item Four additional species and additional \mouse and \human samples
  were obtained from SRA
\end{itemize}
\end{frame}

\subsection{Transcriptome Assembly, Alignment, and Quantification}

\begin{frame}{Assembly, Alignment, Annotation and Quantification}
  \begin{itemize}
  \item QC Using FASTQC v0.11.2
  \item Transcriptomes without suitable reference (\smonkey, \galili,
    \carmeli)
    \begin{itemize}
    \item assembled using Trinity
    \item annotated using Diamond against \human or \mouse
    \end{itemize}
  \item Aligned to the reference (or assembled genome) using STAR
  \item Expression values were quantified using Cufflinks
  \end{itemize}
\end{frame}

\subsection{Additional Methods}
\begin{frame}{Additional Methods}
  All of the code used to produce every analysis presented in this
  poster is available at
  \url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of
  the sequences used are or will be deposited in SRA.
\end{frame}

\section{Results}

\subsection{Transcriptome Statistics}

\begin{frame}{Transcriptome Mapping Statistics: Reads}
\begin{figure}
<<transcriptome.mapping.figure,fig.width=10,fig.height=8,echo=FALSE>>=
print(ggplot(star.logs[grepl("input reads$",field),list(reads=as.numeric(value),species=capfirst(species))],
             aes(x=species,y=reads))
      + coord_flip()
      + geom_boxplot()
      + scale_y_log10(breaks=c(1e5,1e6,1e7,1e8,1e9,1e10),labels=trans_format("log10", math_format(10^.x)))
      + ylab("Reads per Sample")
      + xlab("Species")
      + theme_bw()
      + theme(legend.position="top",
              axis.text.x = element_text(size=10,face="italic"))
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{Transcriptome Mapping Statistics: Unmapped}
\begin{figure}
<<transcriptome.mapping.figure.unmapped,fig.width=8,fig.height=8,echo=FALSE>>=
star.logs.unmapped <- 
    star.logs[grepl("%",field) & 
                  grepl("unmapped",field) & 
                      !grepl("too many mismatches",field),]
star.logs.unmapped <- 
    star.logs.unmapped[,field:=capfirst(gsub("\\% of reads unmapped: ","",field))]
print(ggplot(star.logs.unmapped,
             aes(x=capfirst(species),y=as.numeric(gsub("%","",value))))
      + geom_boxplot()
      + facet_grid(field~.)
      + ylab("Percent of Reads")
      + xlab("Species")
      + theme_bw()
      + coord_flip()
      + theme(legend.position="top",
              axis.text.x = element_text(face="italic"))
      )
@ 
\end{figure}
\end{frame}

<<top.genes.function,echo=FALSE,results="hide">>=

gene.expression <- function(species.name,gene,complete=FALSE,units=TRUE) {
    species.name <- gsub("_"," ",species.name)
    fpkm <- combined.fpkm[species==species.name &
                              (gene_short_name==gene | 
                               gene_id==gene),
                          mean_fpkm][1]
    if (units || complete)
        fpkm <- formatC(fpkm,digits=0,format="f")
    if (units) {
        fpkm <- paste0(fpkm,
                       "\\ \\text{FPKM}")
    }
    if (complete) {
        if (any(grepl("^ENS.+G\\d+$",gene))) {
            fpkm <- paste0("\\text{",gene,"} = ",
                           fpkm)
        } else {
            fpkm <- paste0("\\text{\\textit{",gene,"}} = ",
                           fpkm)
        }
    }
    if (units || complete) {
        fpkm <- paste0("$",fpkm,"$")
    }
    return(fpkm)
}

@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE) {
    temp.table <- 
        head(placenta.lineage.specific.analysis[analysis.name==analysis &
                                                p.value <= 0.05][order(p.value),
                                                                 list(human_name,p.value,fdr,fc),],
             n=5)
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[order(`p (t-test)`),]
        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     columns=c("c","c","p{1.5in}","c","c","c")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}

@

\subsection{Core Placenta Transcriptome}

\begin{frame}{Core Placenta Transcriptome}
\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species))))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(),
              axis.text.y = element_text(face="italic"),
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{Core placenta transcriptome}
\begin{itemize}

\item We considered a gene to be core to the placenta transcriptome if
  a 1:1 ortholog of that gene was expressed with
  $\mathrm{FPKM} \ge 10$ in all species studied and was not a human
  housekeeping gene according to
  \citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}.

\item Multiple components of annexin complexes are included in the core
transcriptome, including \textit{ANXA2}, \textit{ANXA1},
\textit{S100A11}, \textit{S100A10}.

\item ANXA1 is involved in the resolution of the inflammation response
  through inhibition of phospholipase A2 activity and signaling
  through the formyl peptide receptor family, and has previously been
  hypothesized to be important for the maintenance of the
  anti-inflammatory during pregnancy
  \cite{Hutchinson.ea2011:Molecularregulatorsofresolution}.

\item ANXA2 is known to interact with S100A10, S100A6, S100A11, and
  S100A4 \cite{Liu.ea2015:AnnexinA2complexeswith}, and there is
  evidence for its interaction with S100P as well; it is involved in
  cell-cell interaction as well as vesicle trafficking and von
  Willibrand Factor secretion.

\item The expression regulation of another core gene, \textit{TIMP3},
  has been implicated in pre-eclampsia
  \cite{Ma.ea2014:Down-regulationofTIMP3leads,Xiang.ea2013:PromoterhypomethylationofTIMP3}.
\end{itemize}

\end{frame}

\subsection{Primate Placenta Transcriptomes}

\begin{frame}{Primate Placenta Transcriptome}
\begin{table}
  \footnotesize
<<primate.lineage.specific,echo=FALSE,results="asis">>=
pri.all.lin.spec <- placenta.lineage.specific.table("pri.all")
setkey(pri.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed significantly in the primate
  order as compared to all other analyzed species.}
\label{tab:primate_lineage_specific}
\end{table}
\end{frame}

\begin{frame}{PSG1}
\begin{figure}
<<psg.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
plot.label.size <- 3
axis.text.size <- 3.5

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PSG1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,min.fpkm=1,
                 vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"),
                 axis.text.size=axis.text.size,
                 fontsize=plot.label.size,
                 )
@ 
\end{figure}
\end{frame}

\begin{frame}{CTSL}
\begin{figure}
<<ctsl.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
plot.label.size <- 3
axis.text.size <- 3.5
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CTSL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(gene.tree,gene.tree.table,gene.tree.expression,min.fpkm=1,
                 subtree=c("ENSMODG00000002844","ENSMUSG00000055679"),
                 vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=axis.text.size,
                 offset.ratio=1
                 )

@ 
\end{figure}
\end{frame}

\begin{frame}{PAG2}
\begin{figure}
<<pag2.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PAG2" & species=="bos taurus",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=1,
                 subtree=c("ENSSSCG00000006792",
                           "ENSBTAG00000046085"),
                 vp=viewport(layout.pos.col=2,layout.pos.row=2,clip="off"),
                 axis.text.size=axis.text.size,
                 fontsize=plot.label.size,
                 offset.ratio=0.6
                 )

@ 
\end{figure}
\end{frame}

\subsection{Evolution of Placental Hormones}

Reproductive hormones are central to the regulation of pregnancy and
their source, timing, and quantity can vary widely from species to
species \cite{Carter2012:Evolutionofplacentalfunction}. Expression and synthesis of hormones is
perhaps the most well-documented physiological process in the placenta
in many species. Here we quantified the extreme disparity in genes
with hormone products or enzymatic products that directly produce
hormones in multiple mammalian species (\cref{fig:hormone_genes}).

\begin{frame}{Galectin Expression}
\begin{figure}
<<hormone.genes.plot,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

gene.expression.nozero <- function(genes) {
    temp <-
        data.matrix.dt(dcast(combined.fpkm[human_name %in% genes,],
                             human_name~species,
                             value.var="FPKM",
                             fun.aggregate=function(x){x[1]}))
    temp <-
        temp[apply(temp,1,function(x){sum(x,na.rm=TRUE)})>5,]
    return(temp)
}

### hormone genes
hormone.genes <- c("CGA","CGB","CGB1","CGB2","CGB5","CGB8","CSH1","CSH2","CSHL1","CYP19A1",
                   "GH2","GHRH","HSD3B1","HSD3B2","HSD17B2","HSD17B1","NPPB","NPPC","PRL","RLN1")
hormone.expression.nozero <-
    gene.expression.nozero(hormone.genes)

hor.exp.long <- melt(log10(hormone.expression.nozero+1)[rev(1:nrow(hormone.expression.nozero)),])
colnames(hor.exp.long)[1:3] <- c("gene","species","fpkm")
hor.exp.long <- hor.exp.long[hor.exp.long$gene!="IGF2",]

### galectin genes
galectin.genes <- unique(grep("^LGALS",combined.fpkm[,human_name],value=TRUE))

galectin.expression.nozero <-
    gene.expression.nozero(galectin.genes)

gal.exp.long <- melt(log10(galectin.expression.nozero+1)[rev(1:nrow(galectin.expression.nozero)),])
colnames(gal.exp.long)[1:3] <- c("gene","species","fpkm")

### IGF related genes

igfrelated.genes <- unique(grep("^IGF",combined.fpkm[,human_name],value=TRUE))

igfrelated.expression.nozero <-
    gene.expression.nozero(igfrelated.genes)

igf.exp.long <- melt(log10(igfrelated.expression.nozero+1)[rev(1:nrow(igfrelated.expression.nozero)),])
colnames(igf.exp.long)[1:3] <- c("gene","species","fpkm")

igf.exp.long$type <- "IGF Related"
gal.exp.long$type <- "Galectin"
hor.exp.long$type <- "Hormone"

### scalerange <- range(c(igf.exp.long$fpkm,gal.exp.long$fpkm,hor.exp.long$fpkm),na.rm=TRUE)
### gradientends <- scalerange + rep(c(0,100,200), each=2)
### colorends <- c("white", "red", "white", "green", "white", "blue")
### 
### gal.exp.long$fpkm <- gal.exp.long$fpkm+100
### hor.exp.long$fpkm <- hor.exp.long$fpkm+200

combined.long <-
    rbind(igf.exp.long,
          gal.exp.long,
          hor.exp.long)
combined.long$species <-
    factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(combined.long$species)),
           levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(names(species.names))))

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
##         scale_fill_gradientn(colours = colorends, values = rescale(gradientends)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 300, hjust = 0, vjust=1,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long[combined.long$type=="Galectin",], aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\caption{Expression of Hormone Genes, Galectins, and IGF related genes}
\label{fig:hormone_genes}
\end{figure}
\end{frame}

\begin{frame}{Hormone Expression}
\begin{figure}
<<hormone.expression.plot,echo=FALSE>>=
print(ggplot(combined.long[combined.long$type=="Hormone",], aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{IGF Related}
\begin{figure}
<<igfrelated.expression.plot,echo=FALSE>>=
print(ggplot(combined.long[combined.long$type=="IGF Related",], aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\end{figure}
\end{frame}

% \begin{frame}{Prolactin}
% \begin{itemize}
% \item \elephant is only analyzed mammal with high prolactin expression
%   (\textit{PRL}) in the placenta
% \item However, \mouse, \carmeli, and \cow have duplications of
%   \textit{PRL} with expression}
% \end{itemize}
% 
% \begin{figure}
% <<prolactin.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
% 
% tree.id <-
%     genes.to.tree[combined.fpkm[gene_short_name=="PRL" & species=="homo sapiens",gene_id],tree]
% gene.tree.table <- trees[[tree.id]]$table
% gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
% gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
% plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
%                  gene.tree.table,gene.tree.expression,
%                  axis.text.size=5,
%                  fontsize=5)
% @ 
% \end{figure}
% 
% \end{frame}

\subsubsection{GH family}

\begin{frame}{Growth Hormone family}

\begin{columns}
\column{0.4\textwidth}
The \ac{GH} family has expanded in anthropoid primates, resulting in
approximately 5 genes in \human (\textit{GH1}, \textit{GH2},
\textit{CSH1}, \textit{CSH2}, \textit{CSHL1}) and at least three in
\bonobo (two \textit{GH2} and \textit{GH1}; owing to the similarity
between the two \textit{GH1} in \textit{Pan troglodytes} we cannot
reliably distinguish transcripts between the two).

\column{0.6\textwidth}
<<chorionic.somatomammotropin.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CSH1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=4)
@ 
\end{columns}

\end{frame}

\begin{frame}{CGA}

  The human gene \textit{CGA} codes for the alpha subunit of multiple
  hormones including pituitary hormones (luteinizing hormone,
  follicle-stimulating hormone, thyroid-stimulating hormone), as well
  as numerous chorionic gonadotropins with variable beta subunits
  \cite{Maston.ea2002:Chorionicgonadotropinhasrecent}. The function of
  the chorionic gonadotropins is to prevent the regression of the
  corpus luteum \cite{Hanson.ea1971:EffectsofHCGand}. Among the six
  placentas we analyzed here, \textit{CGA} is only expressed in the
  placentas of \human and \bonobo
  (\cref{fig:hormone_genes,fig:cga_cgb_evolution_tree}), which
  corresponds with the emergence of the chorionic gonadotropins in
  anthropoid primates
  \cite{Maston.ea2002:Chorionicgonadotropinhasrecent}. However, oddly,
  \textit{CGA} is expressed at much higher levels in both species than
  the beta subunits with which its product dimerizes \textit{CGA} is
  one of several genes noted by
  \citet{Uddin.ea2008:Distinctgenomicsignaturesof} to have undergone
  adaptive evolution of the \human{}-\textit{P. troglodytes} stem
  lineage.

\begin{figure}
<<chorionic.somatomammotropin.alpha.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

plot.label.size <- 4.5
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CGA" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 offset.ratio=0.7,
                 fontsize=plot.label.size,
                 axis.text.size=3.5
                 )
@ 
\end{figure}

\end{frame}

\begin{frame}{CGB}
\begin{figure}
<<chorionic.somatomammotropin.beta.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CGB1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSMODG00000013768","ENSG00000104818"),
                 offset.ratio=1,
                 fontsize=plot.label.size,
                 axis.text.size=3.5
                 )
@ 
\end{figure}
\end{frame}

\subsection{Relaxin}

\begin{frame}
Relaxin (\textit{RLN}) is a hormone that has multiple effects during
pregnancy including increasing maternal blood flow, widening the pubic
bone and relaxing uterine musculature in preparation for labor. In
\dog, \elephant, \mouse, \carmeli, and \horse there is high expression
of relaxin in the term placenta, with little evidence in other species
examined here (Supplemental \cref{fig:rln2_evolution_tree}). \textit{RLN} is known
to be produced in the placenta of the pregnant \dog and is used in the
diagnosis of canine pregnancy
\cite{Steinetz.ea1989:Serumrelaxinandprogesterone,Tsutsui.ea1991:Determinationofsourceof},
the luteal ovary and placenta of \pig
\cite{Hisaw.ea1948:Relaxininovaryof}, and the placenta of \horse
\cite{Klonisch.ea1997:Placentallocalizationofrelaxin}. However, the
source of relaxin produced by pregnant \elephant remains unclear
\cite{Meyer.ea2004:Speciesandfetalgender}. The high rise in relaxin in
\elephant has been suggested to be a major signal for the initiation
of labor \cite{Niemuller.ea1998:Plasmaconcentrationsofimmunoreactive}.
Here we show that the placenta of elephants show high expression of
relaxin and is likely a major source of \textit{RLN} at term. In
contrast to the results of
\cite{Evans.ea1993:mouserelaxingenenucleotide}, we also see expression
of \textit{RLN} in the placenta of \mouse and \carmeli.

\end{frame}

\subsection{The Evolution of Placental Expression of Galectins}

\begin{frame}{Galectins}
Galectins are recognized as major regulators of pregnancy in both
fetal and maternal tissues
\cite{Than.ea2012:Galectinsguardiansofeutherian}. Indeed, the
evolution of new lineage-specific galectins has occurred in both
primates and ruminants, with the genes \textit{LGALS13},
\textit{LGALS14}, and \textit{LGALS16} recognized as being
placenta-specific in \human and regulating immune responses by
inducing apoptosis of maternal T-cells
\cite{Than.ea2009:primatesubfamilyofgalectins} (Supplemental
\cref{fig:galectins_lgals13_evolution_tree}). Indeed, \textit{LGALS14}
has the highest expression of any galectin in the \human and \smonkey
placenta (\cref{fig:hormone_genes}, Supplemental
\cref{fig:galectins_lgals13_evolution_tree}), and other primate
galectins such as \textit{LGALS13} and \textit{LGALS16} are also
expressed in \human and \smonkey placenta. However, in the closely
related \bonobo, \textit{LGALS14} is not found, and \textit{LGALS1} is
the most abundant galectin. \textit{LGALS1} shows generally high
expression in all placentas analyzed. \textit{LGALS1} is also involved
in mediating maternal-fetal immune tolerance, and is expressed in both
maternal and fetal tissue in primates
\cite{Than.ea2008:Emergenceofhormonaland}. \textit{LGALS1} placental
expression was thought to be confined to eutherian mammals, as there
is evidence of adaptive evolution at this node, as well as the
appearance of estrogen response elements in eutherian cis-regulatory
regions \cite{Than.ea2008:Emergenceofhormonaland}. It was unclear as
to whether galectins such as \textit{LGALS1} were important in the
regulation of pregnancy in marsupials, as well as eutherians. Here we
show substantial expression of galectins in the opossum, especially
\textit{LGALS1} (Supplemental \cref{fig:galectins_evolution_tree}). In
addition, \textit{LGALS2} is also expressed. The presence of galectins
at the maternal-fetal interface in \opossum implies that it was also
present in the ancestral placenta of therian mammals. Indeed, adaptive
evolution of \textit{LGALS1} on the eutherian stem is likely due to
the increased invasive nature of eutehrian placentation rather than
the origination of maternal-fetal tolerance. Although the function of
\textit{LGALS1} in the marsupial placenta needs to be confirmed, it is
likely that the mediation of maternal-fetal immune tolerance by
galectins is an ancient feature of mammalian viviparity.

\end{frame}

\subsection{The Opossum Placenta and the Evolution of Viviparity}

\begin{frame}
The \opossum placenta shows expression of genes that either do not
exist in eutherians or are not found in the term eutherian placenta
(\cref{tab:top50_opossum_placenta}). The existence of embryonic forms
of hemoglobin in fetal tissues at term, as evidenced by the expression
of \textit{HBE} and \textit{HBZ} is a feature of marsupials that is
generally due to their very short gestation periods. Embryonic
hemoglobins in marsupials have a lower affinity for oxygen and allow
for a lower rate of oxygen uptake from maternal tissues
\cite{Holland.ea1997:Characterizationofembryonicglobin}. One novel
uncharacterized gene that shows high expression (\textit{LOC10001195}) is not
found in eutherians; Ensembl gene trees reveal that homologs of this
gene exist in other amniotes such as birds and turtles, but are not
present in eutherian mammals. The homolog of this gene in chicken is
found to expressed in reproductive tract
\cite{Froman.ea2006:expressedsequencetaganalysis}, but its function is
unknown. \textit{WFDC2} is a member of the whey acidic protein (WAP)
family and involved in innate immunity. It is expressed in the mammary
gland of the tammar wallaby during pregnancy and early lactation
\cite{Sharp.ea2007:Molecularevolutionofmonotreme}, and present in
relative abundance in \opossum placenta.

\end{frame}

\subsection{Novel Gene Expansions in the Elephant Placenta}

\begin{frame}{PRSS16}

11 genes that are related
to the \textit{PRSS16} gene, a thymus-specific serine protease that
may be involved regulating the presentation of self-peptides in CD4+
T-lymphocytes \cite{Fornari.ea2012:TranscriptionprofilingofPrss16}.
All 11 of these genes are expressed in the elephant placenta, three of
which have very abundant mRNAs
(\cref{tab:top50_elephant_placenta} and \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}C). It is
possible that this elephant-specific gene family may be involved in
immunity in the elephant placenta and the mediation of maternal-fetal
immune tolerance; however, future work will need to test these
hypotheses. 
\begin{figure}
<<prss16.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRSS16" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 vp=viewport(layout.pos.col=1,layout.pos.row=2,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=axis.text.size,
                 offset.ratio=1.2
                 )

@ 
\end{figure}
\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}    
\begin{itemize}
\item Sequenced placentas of \Sexpr{star.logs[,length(unique(species))]} species

\item lineage-specific expansions of gene families and differential
  gene expression

\item Identified common expression patterns of

\item Still much more to learn about the changes in placenta
  expression during placentation

\item Many clades are only represented by a single species; need to
  have more information to make stronger inferences about evolution of
  placenta expression

\end{itemize}

\end{frame}

% Acronyms
\input{acronyms}

\begin{frame}[plain]{References}
  \begin{center}
    \mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]
  \end{center}
\end{frame}

\end{document}