### 2016-02-22 presentation/placenta_viviparous_presentation.Rnw
fix gene trees for most of the plots

\documentclass[ignorenonframetext]{beamer}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
\usepackage{pdfpcnotes}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage{qrcode}
\usepackage{cleveref}
\usepackage[absolute,overlay]{textpos}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}

\usepackage{texshade}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usepackage{xspace}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\renewcommand*{\bibfont}{\tiny}
\usepackage{mathtools,cancel}

\renewcommand{\CancelColor}{\color{red}} %change cancel color to red

\makeatletter
\let\my@cancelto\cancelto %copy over the original cancelto command
\newcommand<>{\cancelto}[2]{\alt#3{\my@cancelto{#1}{#2}}{\mathrlap{#2}\phantom{\my@cancelto{#1}{#2}}}}
% redefine the cancelto command, using \phantom to assure that the
% result doesn't wiggle up and down with and without the arrow
\makeatother

\usepackage{multirow}
\usepackage{array}

\mode<presentation>{ 
  \usetheme{CambridgeUS}
  \usecolortheme{crane}
%   % http://identitystandards.illinois.edu/graphicstandardsmanual/generalguidelines/colors.html
  \definecolor{ilboldblue}{HTML}{002058}
  \definecolor{ilboldorange}{HTML}{E87722}
  \definecolor{ilblue}{HTML}{606EB2}
  \definecolor{ilorange}{HTML}{D45D00}
%   \setbeamercolor{alerted text}{fg=ilboldorange}
%   \setbeamercolor*{palette primary}{fg=ilboldblue,bg=ilorange!70!white}
%   \setbeamercolor*{palette secondary}{fg=ilboldblue,bg=ilorange!60!white}
%   \setbeamercolor*{palette tertiary}{bg=ilblue!60!white,fg=ilboldorange}
%   \setbeamercolor*{palette quaternary}{fg=ilboldblue,bg=ilorange!80!white}
%   
%   \setbeamercolor*{sidebar}{fg=ilorange,bg=ilboldblue}
%   
%   \setbeamercolor*{palette sidebar primary}{fg=ilblue!10!white,bg=ilorange}
%   \setbeamercolor*{palette sidebar secondary}{fg=ilorange}
%   \setbeamercolor*{palette sidebar tertiary}{fg=ilblue}
%   \setbeamercolor*{palette sidebar quaternary}{fg=ilorange}
%   
%   % \setbeamercolor*{titlelike}{parent=palette primary}
%   \setbeamercolor{titlelike}{parent=palette primary,fg=ilboldblue,bg=ilorange}
%   \setbeamercolor{frametitle}{fg=ilboldorange,bg=ilblue!20!white}
%   \setbeamercolor{frametitle right}{fg=ilboldblue,bg=ilorange}
%   
%   \setbeamercolor*{separation line}{}
%   \setbeamercolor*{fine separation line}{}
%   \setbeamercovered{transparent}  
  \logo{\begin{tikzpicture}% Pale figure
      {\node[opacity=0.5]{\IfFileExists{figures/igb_wordmark.pdf}{\includegraphics[width=2cm,height=1cm,keepaspectratio]{figures/igb_wordmark}}{}%
        };}%
    \end{tikzpicture}}
}

\title[Placenta Gene Expression]{The evolution of gene expression in the term placenta of viviparous mammals}
\author[Don L Armstrong \textit{et al.}]{Don L Armstrong, Michael R.
  McGowen, Amy Weckle, Jason Caravas, Dalen Agnew, Kurt Benirschke,
  Sue Savage-Rumbaugh, Eviatar Nevo, Chong J Kim, Günter P. Wagner,
  Roberto Romero, and Derek E. Wildman}
\institute[IGB]{Institute for Genomic Biology, Computing Genomes 
  for Reproductive Health, University of Illinois, Urbana-Champaign}
\subject{Placenta Expression}

\date{\today}
\renewcommand\cite[1]{\autocite{#1}}
% \pagestyle{fancy}

% species names
\newcommand{\smonkey}[0]{\textit{A. fusciceps}\xspace}
\newcommand{\cow}[0]{\textit{B. taurus}\xspace}
\newcommand{\dog}[0]{\textit{C. familiaris}\xspace}
\newcommand{\armadillo}[0]{\textit{D. novemcinctus}\xspace}
\newcommand{\horse}[0]{\textit{E. caballus}\xspace}
\newcommand{\elephant}[0]{\textit{L. africana}\xspace}
\newcommand{\human}[0]{\textit{H. sapiens}\xspace}
\newcommand{\opossum}[0]{\textit{M. domestica}\xspace}
\newcommand{\mouse}[0]{\textit{M. musculus}\xspace}
\newcommand{\galili}[0]{\textit{N. galili}\xspace}
\newcommand{\sheep}[0]{\textit{O. aries}\xspace}
\newcommand{\bonobo}[0]{\textit{P. paniscus}\xspace}
\newcommand{\carmeli}[0]{\textit{S. carmeli}\xspace}
\newcommand{\pig}[0]{\textit{S. scrofa}\xspace}

% top 50 gene table explanation. Note that if you modify this, you
% should probably also modify the human table explanation, which is
% slightly different.
\newcommand{\tableexplanation}[2]{The top 50 genes by \ac{FPKM} in #1
  which are not annotated as Mt\_rRNA, Mt\_tRNA, rRNA or whose \human
  ortholog is a housekeeping gene. Gene symbol is the gene symbol or
  Ensembl gene id. Human Symbol is the 1:1 human ortholog or the best
  human alignment when enclosed in parenthesis. Human Name is the HUGO
  name corresponding to the human symbol. A complete list of all genes
  is in \cref{#2}.}

\newcommand{\genetreeexplanation}[1]{Gene tree and heatmap of
  expression of genes related to #1. The gene tree was obtained from
  Ensembl~\cite{Cunningham.ea2015:Ensembl2015} and genes from species
  not included in this study or having expression by \ac{FPKM} less
  than 1 were excluded from the plotted tree. Heatmap cells in white
  are from species where there was no expression of the relevant
  transcript (or a species where that transcript does not apply). The
  species names are abbreviated with the first letter of the genus and
  first two letters of the species (for example, \textit{Homo sapiens}
  is \textit{H.sa.}). Genes with duplicate gene names in the same
  species (or no gene named defined) are listed by the Ensemble gene
  ID.}

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.8\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 0)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")
library("scales")
library("magrittr")
library("tidyr")

@

<<source.to.latex,echo=FALSE,message=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R")))>>=
for (a in paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R"))) {
    source(a)
}
@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
setkey(hugo.nomenclature,"Approved Symbol")
@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

@

<<plot.defaults,echo=FALSE,results="hide">>=
grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)

@

\IfFileExists{figures/relevant_xkcd.png}{\frame[plain]{\centering \includegraphics[width=\textwidth,height=0.8\textheight,keepaspectratio]{figures/relevant_xkcd.png}

    \url{https://xkcd.com/1605/}}}

\frame[plain]{\titlepage}

\section{Background}

\begin{frame}{Function and Evolution of Placenta}
\begin{itemize}
\item Placenta functions in the transfer of nutrients, oxygen, and
  waste between mother and offspring
\item Placenta is highly variable structure and form in mammals, even though major
  function is conserved
\item Placenta likely arose before the origin of therian mammals
\item Form relatively well characterized, but diversity of molecular
  environment of the placenta is only recently being understood
\item Possible role of evolutionary interplay between maternal and
  fetal strategies
\end{itemize}
\end{frame}

\begin{frame}{Impacts on Human Health?}
\begin{itemize}
\item Humans possess a deeply invasive trophoblast layer that can
  penetrate the myometrium of the uterus
  \cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
\item Defects in placental growth can lead to pre-eclampsia, which can
  cause hypertension, proteinuria, and mortality of mother and/or
  infant \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
\item Pre-eclampsia has been seen in \textit{Pan troglodytes}
\cite{Stout.ea1969:Glomerularcapillaryendothelialswelling}, and
\textit{Gorilla gorilla}
\cite{Thornton.ea1992:Convulsionsinpregnancyin}
\item Far higher incidence in \textit{Homo sapiens}, with as many as 
  4\% of pregnancies United States affected
  \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
\end{itemize}
\end{frame}

\begin{frame}{Overall Hypotheses}
\begin{itemize}
\item Gene expression patterns correlate with placenta morphology
\item Altered expression of genes correlates with invasiveness, and
  therefore pre-eclampsia
\item Maternal/Fetal evolutionary strategies will be reflected in gene
  expression patterns
\end{itemize}
\end{frame}

\begin{frame}{Species Studied}
\begin{figure}
<<species.tree,fig.width=8,fig.height=6,echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Nannospalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.5,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.5,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.5,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.5,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.5,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.5,node=6,width=0.12)
      ## sheep
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.5,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.5,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.5,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.5,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.5,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.5,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.5,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.5,node=13,width=0.18)
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{Variation in Gestational Periods and Litter Size}
\footnotesize
%\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Gestation (days) & Litter Size  & Source \\
    \midrule
    \textit{Ateles fusciceps}     & 224 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Bos taurus}           & 278 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Canis familiaris}     & 61 & 4 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Dasypus novemcinctus} & 150 & 4 & \citet{Smith1984:amazingarmadillo} \\
    \textit{Equus caballus}       & 336 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation}\\
    \textit{Loxodonta africana}   & 644 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Homo sapiens}         & 270 & 1.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Monodelphis domestica}& 15 & 7.1 & \citet{Harder.ea1993:Gestationandplacentationin} \\
    \textit{Mus musculus}         & 20 & 6.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Nannospalax galili}   & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals} \\
    \textit{Ovis aries}           & 151 & 1.35 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Pan paniscus}         & 223 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental} \\
    \textit{Spalax carmeli}       & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals}\\
    \textit{Sus scrofa}           & 114 & 5.8 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \bottomrule
  \end{tabular}
%  \caption{Gestational periods in days and average liter sizes for studied species.}
%  \label{tab:gestational_periods}
%\end{table}
\end{frame}

\begin{frame}{Variations in Placenta Morphology}
%\begin{table}
\footnotesize
\begin{tabular}{c c c}
  \toprule
  Species & Interhemal Membrane & Pattern \\
  \midrule
  \smonkey & hemochorial & trabecular \\
  \cow     & epitheliochorial & villous \\
  \dog     & endotheliochorial & labyrinthine \\
  \armadillo & hemochorial & trabecular \\
  \horse & epitheliochorial & villous \\
  \elephant & endotheliochorial & labyrinthine \\
  \human & hemochorial & villous \\
% http://onlinelibrary.wiley.com/doi/10.1002/jez.1402660511/pdf
  \opossum & epitheliochorial & choriovitelline \\
  \mouse & hemotrichorial & labyrinthine \\
  \galili & hemotrichorial & labyrinthine \\
  \sheep & epitheliochorial & villous \\
  \bonobo & hemochorial & villous \\
% http://www.biolreprod.org/content/early/2014/10/17/biolreprod.114.124818.short
  \carmeli & hemotrichorial & labyrinthine \\
  \pig & epitheliochorial & diffuse \\
  \bottomrule
\end{tabular}
%\end{table}
\end{frame}

\section{Methods}

\subsection{Tissue/Extraction/Sequencing}

\begin{frame}{Collection of Placental Tissue}
\begin{itemize}
\item Collected from 8 mammalian species
\item Only fetal portion was sampled
\item Sequenced using GAII with an insert size of 230 bp
\item Four additional species and additional \mouse and \human samples
  were obtained from SRA
\end{itemize}
\end{frame}

\subsection{Transcriptome Assembly, Alignment, and Quantification}

\begin{frame}{Assembly, Alignment, Annotation and Quantification}
  \begin{itemize}
  \item QC Using FASTQC v0.11.2
  \item Transcriptomes without suitable reference (\smonkey, \galili,
    \carmeli)
    \begin{itemize}
    \item assembled using Trinity
    \item annotated using Diamond against \human or \mouse
    \end{itemize}
  \item Aligned to the reference (or assembled genome) using STAR
  \item Expression values were quantified using Cufflinks
  \end{itemize}
\end{frame}

\subsection{Additional Methods}
\begin{frame}{Additional Methods}
  All of the code used to produce every analysis presented in this
  poster is available at
  \url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of
  the sequences used are or will be deposited in SRA.
\end{frame}

\section{Results}

\subsection{Transcriptome Statistics}

\begin{frame}{Transcriptome Mapping Statistics: Reads}
\begin{figure}
<<transcriptome.mapping.figure,fig.width=10,fig.height=8,echo=FALSE>>=
print(ggplot(star.logs[grepl("input reads$",field),list(reads=as.numeric(value),species=capfirst(species))],
             aes(x=species,y=reads))
      + coord_flip()
      + geom_boxplot()
      + scale_y_log10(breaks=c(1e5,1e6,1e7,1e8,1e9,1e10),labels=trans_format("log10", math_format(10^.x)))
      + ylab("Reads per Sample")
      + xlab("Species")
      + theme_bw()
      + theme(legend.position="top",
              axis.text.x = element_text(size=10,face="italic"))
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{Transcriptome Mapping Statistics: Unmapped}
\begin{figure}
<<transcriptome.mapping.figure.unmapped,fig.width=8,fig.height=8,echo=FALSE>>=
star.logs.unmapped <- 
    star.logs[grepl("%",field) & 
                  grepl("unmapped",field) & 
                      !grepl("too many mismatches",field),]
star.logs.unmapped <- 
    star.logs.unmapped[,field:=capfirst(gsub("\\% of reads unmapped: ","",field))]
print(ggplot(star.logs.unmapped,
             aes(x=capfirst(species),y=as.numeric(gsub("%","",value))))
      + geom_boxplot()
      + facet_grid(field~.)
      + ylab("Percent of Reads")
      + xlab("Species")
      + theme_bw()
      + coord_flip()
      + theme(legend.position="top",
              axis.text.x = element_text(face="italic"))
      )
@ 
\end{figure}
\end{frame}

<<top.genes.function,echo=FALSE,results="hide">>=

gene.expression <- function(species.name,gene,complete=FALSE,units=TRUE) {
    species.name <- gsub("_"," ",species.name)
    fpkm <- combined.fpkm[species==species.name &
                              (gene_short_name==gene | 
                               gene_id==gene),
                          mean_fpkm][1]
    if (units || complete)
        fpkm <- formatC(fpkm,digits=0,format="f")
    if (units) {
        fpkm <- paste0(fpkm,
                       "\\ \\text{FPKM}")
    }
    if (complete) {
        if (any(grepl("^ENS.+G\\d+$",gene))) {
            fpkm <- paste0("\\text{",gene,"} = ",
                           fpkm)
        } else {
            fpkm <- paste0("\\text{\\textit{",gene,"}} = ",
                           fpkm)
        }
    }
    if (units || complete) {
        fpkm <- paste0("$",fpkm,"$")
    }
    return(fpkm)
}

@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE) {
    temp.table <- 
        head(placenta.lineage.specific.analysis[analysis.name==analysis &
                                                p.value <= 0.05][order(p.value),
                                                                 list(human_name,p.value,fdr,fc),],
             n=5)
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[order(`p (t-test)`),]
        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     align=c("c","c","p{1.5in}","c","c","c")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}

@

\subsection{Core Placenta Transcriptome}

\begin{frame}{Core Placenta Transcriptome}
\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species))))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(),
              axis.text.y = element_text(face="italic"),
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\end{figure}
\end{frame}

\begin{frame}{Core placenta transcriptome}
\begin{itemize}

\item Gene is core to the placenta transcriptome if a 1:1 ortholog is
  expressed with $\mathrm{FPKM} \ge 10$ in all species studied and was
  not a human housekeeping gene according to
  \citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}.

\item Multiple components of annexin complexes are included in the core
transcriptome, including \textit{ANXA2}, \textit{ANXA1},
\textit{S100A11}, \textit{S100A10}.

\item ANXA1 is involved in resolution of inflammation by inhibiting
  phospholipase A2 activity and signaling via formyl peptide receptor
  family
\item hypothesized to be important for the maintenance of the
  anti-inflammatory during pregnancy
  \cite{Hutchinson.ea2011:Molecularregulatorsofresolution}.

\item ANXA2 interacts with S100A10, S100A6, S100A11, and
  S100A4 \cite{Liu.ea2015:AnnexinA2complexeswith}, and there is
  evidence for its interaction with S100P as well
\item ANXA2 involved in cell-cell interaction as well as vesicle
  trafficking and von Willibrand Factor secretion.

\item \textit{TIMP3} has been implicated in pre-eclampsia
  \cite{Ma.ea2014:Down-regulationofTIMP3leads,Xiang.ea2013:PromoterhypomethylationofTIMP3}.
\end{itemize}

\end{frame}

\subsection{Primate Placenta Transcriptomes}

\begin{frame}{Primate Placenta Transcriptome}
\begin{table}
  \footnotesize
<<primate.lineage.specific,echo=FALSE,results="asis">>=
pri.all.lin.spec <- placenta.lineage.specific.table("pri.all")
setkey(pri.all.lin.spec,"Human Symbol")
@ 
\end{table}
\end{frame}

\begin{frame}{Pregnancy-Specific Glycoprotein Expansion}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item 11 highly expressed \textit{PSG} genes in \human and some
  expression in \smonkey and \bonobo
\item \mouse specific expansion and high expression of \textit{PSG} as
  well
\item No other species have high expression of \textit{PSG} or
  \textit{CEACAM}
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<psg.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
plot.label.size <- 3
axis.text.size <- 3.5

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PSG1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,min.fpkm=1,
                 vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"),
                 axis.text.size=axis.text.size,
                 fontsize=plot.label.size,
                 )
@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}{Cathepsin-L expansion}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item Cathepsins are highly expressed in many placentas
\item Expanded in \mouse and \galili and possibly also \carmeli to a
  lesser degree
\item Intracellular protease which degrades collagen, elastin, and in
  IGF1R
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<ctsl.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
plot.label.size <- 3
axis.text.size <- 3.5
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CTSL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(gene.tree,gene.tree.table,gene.tree.expression,min.fpkm=1,
                 subtree=c("ENSMODG00000002844","ENSMUSG00000055679"),
                 vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=axis.text.size,
                 offset.ratio=1
                 )

@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}{Pregnancy associated glycoprotein}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item Massive expansion in ruminants (\cow and \sheep)
\item 28 members of this family have $\mathrm{FPKM} \ge 1$ in \cow
\item \Sexpr{gene.expression("bos taurus","PAG2",complete=TRUE)}
\item \Sexpr{gene.expression("bos taurus","PAG10",complete=TRUE)}
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<pag2.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PAG2" & species=="bos taurus",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,min.fpkm=1,
                 subtree=c("ENSSSCG00000006792",
                           "ENSBTAG00000046085"),
                 vp=viewport(layout.pos.col=2,layout.pos.row=2,clip="off"),
                 axis.text.size=axis.text.size,
                 fontsize=plot.label.size,
                 offset.ratio=0.6
                 )

@ 
\end{figure}
\end{columns}
\end{frame}

\subsection{Evolution of Placental Hormones}

Reproductive hormones are central to the regulation of pregnancy and
their source, timing, and quantity can vary widely from species to
species \cite{Carter2012:Evolutionofplacentalfunction}. Expression and synthesis of hormones is
perhaps the most well-documented physiological process in the placenta
in many species. Here we quantified the extreme disparity in genes
with hormone products or enzymatic products that directly produce
hormones in multiple mammalian species (\cref{fig:hormone_genes}).

\begin{frame}{Galectin Expression}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item Galectins are major regulators of pregnancy in fetal and
  maternal tissues \cite{Than.ea2012:Galectinsguardiansofeutherian}
\item \textit{LGALS1} expressed everywhere
\item Primate expansion of \textit{LGALS13} and \textit{LGALS14}
\item Likely present in ancestral placenta of therian mammals as it is
  expressed even in \opossum
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<hormone.genes.plot,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

gene.expression.nozero <- function(genes) {
    temp <-
        data.matrix.dt(dcast(combined.fpkm[human_name %in% genes,],
                             human_name~species,
                             value.var="FPKM",
                             fun.aggregate=function(x){x[1]}))
    temp <-
        temp[apply(temp,1,function(x){sum(x,na.rm=TRUE)})>5,]
    return(temp)
}

### hormone genes
hormone.genes <- c("CGA","CGB","CGB1","CGB2","CGB5","CGB8","CSH1","CSH2","CSHL1","CYP19A1",
                   "GH2","GHRH","HSD3B1","HSD3B2","HSD17B2","HSD17B1","NPPB","NPPC","PRL","RLN1")
hormone.expression.nozero <-
    gene.expression.nozero(hormone.genes)

hor.exp.long <- melt(log10(hormone.expression.nozero+1)[rev(1:nrow(hormone.expression.nozero)),])
colnames(hor.exp.long)[1:3] <- c("gene","species","fpkm")
hor.exp.long <- hor.exp.long[hor.exp.long$gene!="IGF2",]

### galectin genes
galectin.genes <- unique(grep("^LGALS",combined.fpkm[,human_name],value=TRUE))

galectin.expression.nozero <-
    gene.expression.nozero(galectin.genes)

gal.exp.long <- melt(log10(galectin.expression.nozero+1)[rev(1:nrow(galectin.expression.nozero)),])
colnames(gal.exp.long)[1:3] <- c("gene","species","fpkm")

### IGF related genes

igfrelated.genes <- unique(c(grep("^IGF",combined.fpkm[,human_name],value=TRUE),"H19"))

igfrelated.expression.nozero <-
    gene.expression.nozero(igfrelated.genes)

igf.exp.long <- melt(log10(igfrelated.expression.nozero+1)[rev(1:nrow(igfrelated.expression.nozero)),])
colnames(igf.exp.long)[1:3] <- c("gene","species","fpkm")

igf.exp.long$type <- "IGF Related"
gal.exp.long$type <- "Galectin"
hor.exp.long$type <- "Hormone"

### scalerange <- range(c(igf.exp.long$fpkm,gal.exp.long$fpkm,hor.exp.long$fpkm),na.rm=TRUE)
### gradientends <- scalerange + rep(c(0,100,200), each=2)
### colorends <- c("white", "red", "white", "green", "white", "blue")
### 
### gal.exp.long$fpkm <- gal.exp.long$fpkm+100
### hor.exp.long$fpkm <- hor.exp.long$fpkm+200

combined.long <-
    rbind(igf.exp.long,
          gal.exp.long,
          hor.exp.long)
combined.long$species <-
    factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(combined.long$species)),
           levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(names(species.names))))

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
##         scale_fill_gradientn(colours = colorends, values = rescale(gradientends)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "right",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(size=15,face="italic"),
               axis.text.x = element_text(size=16,angle = 300, hjust = 0, vjust=1,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long[combined.long$type=="Galectin",], aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}{Hormone Expression}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item \human specific high expression of \textit{CGA}
\item \human expansion of \textit{CGB}
\item Vasodilation-involved proteins (\textit{NPPB} \& \textit{NPPC})
  highly-expressed in \elephant and \dog, but not expressed elsewhere
  \pnote{NPPB and NPPC have been seen in decidua of mouse
    \cite{Cameron.ea1996:sitesofgeneexpression}.}
\item \textit{PRL} highly expressed in \elephant
\pnote{more about PRL later}
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<hormone.expression.plot,echo=FALSE>>=
print(ggplot(combined.long[combined.long$type=="Hormone",], aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}{IGF Related}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item \textit{IGF2} is highly expressed in most clades
\item \textit{IGF2} promotes the growth and division of cells; highly active during fetal development
\item \textit{IGF2} binding partners are also highly expressed
  (\textit{IGFBP3}, \textit{IGFBP5})
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<igfrelated.expression.plot,echo=FALSE>>=
print(ggplot(combined.long[combined.long$type=="IGF Related",], aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}{Prolactin}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item \elephant is only analyzed mammal with high prolactin expression
  (\textit{PRL}) in the placenta
\item However, \mouse, \carmeli, and \cow have duplications of
  \textit{PRL} with expression
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<prolactin.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRL" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=4)
@ 
\end{figure}
\end{columns}
\end{frame}

\subsubsection{GH family}

\begin{frame}{Growth Hormone (GH) family}

\begin{columns}
\column{0.4\textwidth}
The GH family has expanded in anthropoid primates, resulting in
approximately 5 genes in \human (\textit{GH1}, \textit{GH2},
\textit{CSH1}, \textit{CSH2}, \textit{CSHL1}) and at least three in
\bonobo (two \textit{GH2} and \textit{GH1}; owing to the similarity
between the two \textit{GH1} in \textit{Pan troglodytes} we cannot
reliably distinguish transcripts between the two).

\column{0.6\textwidth}
<<chorionic.somatomammotropin.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CSH1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 offset=2,
                 axis.text.size=5,
                 fontsize=10)
@ 
\end{columns}

\end{frame}

\begin{frame}{CGA}

  \begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item \textit{CGA} codes for the alpha subunit of multiple hormones as
  well as numerous chorionic gonadotropins with variable beta subunits
  \cite{Maston.ea2002:Chorionicgonadotropinhasrecent}.

\item Chorionic gonadotropins prevent the regression of the corpus
  luteum \cite{Hanson.ea1971:EffectsofHCGand}.

\item \textit{CGA} only expressed in the placentas of \human and
  \bonobo

  \pnote{Expression corresponds with emergence of the chorionic gonadotropins in
    anthropoid primates
    \cite{Maston.ea2002:Chorionicgonadotropinhasrecent}.}

\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<chorionic.somatomammotropin.alpha.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

plot.label.size <- 4.5
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CGA" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 offset.ratio=1,
                 fontsize=8,
                 axis.text.size=5
                 )
@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}{CGB}
  \begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item \textit{CGB} is highly expressed in \human

\item Some expansion of \textit{CGB} in \bonobo, but not highly expressed
\item High expression of \textit{LHB} in \smonkey
\begin{itemize}
\item May be indicative of expansion
\item Would have to examine assembled transcripts to tell
\end{itemize}
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<chorionic.somatomammotropin.beta.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="CGB1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 subtree=c("ENSMODG00000013768","ENSG00000104818"),
                 offset.ratio=1,
                 fontsize=10,
                 axis.text.size=3.5
                 )
@ 
\end{figure}
\end{columns}
\end{frame}

\subsection{Relaxin}

\begin{frame}{Relaxin}
  \begin{columns}
    \column{0.4\textwidth}
    \begin{itemize}
    \item \textit{RLN} increasing maternal blood flow, widening the pubic
      bone and relaxing uterine musculature in preparation for labor
    \item We show that the placenta is a source of \textit{RLN} in \dog,
      \elephant, \mouse, \carmeli, and \horse.
    \end{itemize}
    \column{0.6\textwidth}
    \begin{figure}
<<rln2.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="RLN2" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 axis.text.size=5,
                 fontsize=5)
@ 
\end{figure}
\end{columns}
\end{frame}

\subsection{The Evolution of Placental Expression of Galectins}

\begin{frame}{Galectins}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item Galectins are major regulators of pregnancy in fetal and
  maternal tissues
\item \textit{LGALS1} mediates maternal-fetal
  immune tolerance
\item substantial expression of galectins in the \opossum
\item consistent with ancestral placenta may have maternal-fetal
  immune tolerance mediated by galectins
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<galectins.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="LGALS1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 offset=0.8,
                 axis.text.size=7,
                 fontsize=7,
                 subtree=c("ENSMODG00000000457","ENSCAFG00000016577")
                 )
@ 
\end{figure}
\end{columns}
\end{frame}

\begin{frame}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item Evolution of new lineage-specific galectins has occurred in both
  primates and ruminants, with the genes \textit{LGALS13},
  \textit{LGALS14}, and \textit{LGALS16} recognized as being
  placenta-specific in \human and regulating immune responses by
  inducing apoptosis of maternal T-cells
  \cite{Than.ea2009:primatesubfamilyofgalectins}
\end{itemize}

\column{0.6\textwidth}
\begin{figure}
<<lgals13.galectins.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="LGALS13" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),
                 gene.tree.table,gene.tree.expression,
                 offset=0.8,
                 axis.text.size=7,
                 fontsize=10
                 )
@ 
\end{figure}
\end{columns}
\end{frame}

\subsection{The Opossum Placenta and the Evolution of Viviparity}

\begin{frame}
\begin{itemize}
\item The \opossum placenta shows expression of genes that either do
  not exist in eutherians or are not found in the term eutherian
  placenta

\item The existence of embryonic forms of hemoglobin in fetal tissues
  at term, as evidenced by the expression of \textit{HBE} and
  \textit{HBZ} is a feature of marsupials that is generally due to
  their very short gestation periods.

\item Embryonic hemoglobins in marsupials have a lower affinity for
  oxygen and allow for a lower rate of oxygen uptake from maternal
  tissues \cite{Holland.ea1997:Characterizationofembryonicglobin}.

\item One novel uncharacterized gene that shows high expression
  (\textit{LOC10001195}) is not found in eutherians; Ensembl gene
  trees reveal that homologs of this gene exist in other amniotes such
  as birds and turtles, but are not present in eutherian mammals.

\item \textit{WFDC2} is a member of the whey acidic protein (WAP)
  family and involved in innate immunity.
\end{itemize}
\end{frame}

\subsection{Novel Gene Expansions in the Elephant Placenta}

\begin{frame}{PRSS16}
\begin{columns}
\column{0.4\textwidth}
\begin{itemize}
\item \textit{PRSS16}, a thymus-specific serine protease that
  regulates the presentation of self-peptides in CD4+ T-lymphocytes
  \cite{Fornari.ea2012:TranscriptionprofilingofPrss16}.
\item Possibly involved in the mediation of maternal-fetal immune
  tolerance in \elephant
\end{itemize}
\column{0.6\textwidth}
\begin{figure}
<<prss16.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRSS16" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 vp=viewport(layout.pos.col=1,layout.pos.row=2,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=axis.text.size,
                 offset.ratio=1.2
                 )

@ 
\end{figure}
\end{columns}
\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}    
\begin{itemize}
\item Sequenced placentas of \Sexpr{star.logs[,length(unique(species))]} species

\item lineage-specific expansions of gene families and differential
  gene expression

\item Identified common expression patterns of the mammalian
  expression

\item Still much more to learn about the changes in placenta
  expression during placentation

\item Many clades are only represented by a single species; need to
  have more information to make stronger inferences about evolution of
  placenta expression

\item Lots to learn about what is actually driving the changes in
  expression

\end{itemize}

\end{frame}

% Acronyms
\input{acronyms}

\begin{frame}[plain,allowframebreaks]{References}
  \begin{center}
    \mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]
  \end{center}
\end{frame}

\end{document}