### 2016-03-25 manuscript/common_latex.tex
add lineage table explanation to common_latex

% species names
\newcommand{\smonkey}[0]{\textit{A. fusciceps}\xspace}
\newcommand{\cow}[0]{\textit{B. taurus}\xspace}
\newcommand{\dog}[0]{\textit{C. familiaris}\xspace}
\newcommand{\armadillo}[0]{\textit{D. novemcinctus}\xspace}
\newcommand{\horse}[0]{\textit{E. caballus}\xspace}
\newcommand{\elephant}[0]{\textit{L. africana}\xspace}
\newcommand{\human}[0]{\textit{H. sapiens}\xspace}
\newcommand{\opossum}[0]{\textit{M. domestica}\xspace}
\newcommand{\mouse}[0]{\textit{M. musculus}\xspace}
\newcommand{\galili}[0]{\textit{S. galili}\xspace}
\newcommand{\sheep}[0]{\textit{O. aries}\xspace}
\newcommand{\bonobo}[0]{\textit{P. paniscus}\xspace}
\newcommand{\carmeli}[0]{\textit{S. carmeli}\xspace}
\newcommand{\pig}[0]{\textit{S. scrofa}\xspace}

% top 50 gene table explanation. Note that if you modify this, you
% should probably also modify the human table explanation, which is
% slightly different.
\newcommand{\tableexplanation}[2]{The top 50 genes by FPKM in #1 which
  are not annotated as Mt\_rRNA, Mt\_tRNA, rRNA or whose \human
  ortholog is a housekeeping gene. Gene symbol is the gene symbol or
  Ensembl gene id. Human Symbol is the 1:1 human ortholog or the best
  human alignment when enclosed in parenthesis. Human Name is the HUGO
  name corresponding to the human symbol. A complete list of all genes
  is in \cref{#2}.}

\newcommand{\genetreeexplanation}[2][1]{Gene tree and heatmap of
  expression of genes related to #2. The gene tree was obtained from
  Ensembl~\cite{Cunningham.ea2015:Ensembl2015} and genes from species
  not included in this study or having expression by FPKM less than #1
  were excluded from the plotted tree. Heatmap cells in white are from
  species where there was no expression of the relevant transcript (or
  a species where that transcript does not apply). The species names
  are abbreviated with the first letter of the genus and first two
  letters of the species (for example, \textit{Homo sapiens} is
  \textit{H.sa.}). Genes with duplicate gene names in the same species
  (or no gene named defined) are listed by the Ensemble gene ID.}

\newcommand{\lineagetableexplanation}[3]%
[A two-sided t test was used to calculate p values.]{%
  Genes whose expression changed in #2 as compared to #3. #1 FDR
  is the False Discovery Rate as estimated by the Benjamini and
  Hochberg procedure \cite{Benjamini1995:bh_procedure}.
  $\mathrm{log}_2$ FC is the base-2 logarithm of the fold change from
  #2 to #3 .}

\newcommand{\lineagetableexplanationtypeiii}[2]{%
  \lineagetableexplanation[A nested type-III anova was used to
  calculate p values.]{#1}{#2}}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manuscript.Rnw"
%%% End: