### 2016-04-25 poster/placenta_viviparous_poster.Rnw
add start of poster

\documentclass[a0paper,landscape]{tikzposter}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{booktabs}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage{qrcode}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}
% \usepackage[nomargin,inline,draft]{fixme}
% \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\renewcommand*{\bibfont}{\small}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.  Use showboxes option to show outlines of textboxes.
% \usepackage[absolute]{textpos}
\usepackage[absolute,showboxes]{textpos}

\usepackage{multirow}
\usepackage{array}

\title{The Evolution of Gene Expression in the Term Placenta of
  Viviparous Mammals}

\author{Don L. Armstrong, Michael R. McGowen, Amy Weckle, Jason
  Caravas, Dalen Agnew, Sue Savage-Rumbaugh, Eviatar Nevo, Chong J.
  Kim, Günter P. Wagner, Roberto Romero, and Derek Wildman}
\institute{Institute for Genomic Biology, Computing Genomes for
  Reproductive Health, University of Illinois, Urbana-Champaign}
\usetheme{Default}

% use this
% http://www.identitystandards.illinois.edu/graphicstandardsmanual/othermedia/posterpresentation.html
\definecolorstyle{UofI}{
  \definecolor{boldBlue}{HTML}{002058}
  \definecolor{boldOrange}{HTML}{E87722}
  \definecolor{professionalBlue}{HTML}{606EB2}
  \definecolor{professionalOrange}{HTML}{D45D00}
  }{
    \colorlet{backgroundcolor}{professionalBlue}
    \colorlet{framecolor}{black}
    \colorlet{titlefgcolor}{boldBlue}
    \colorlet{titlebgcolor}{white}
    \colorlet{blockbodybgcolor}{white}
    \colorlet{blocktitlebgcolor}{professionalOrange}
    \colorlet{blocktitlefgcolor}{boldBlue}
    }
\usecolorstyle{UofI}
\defineblockstyle{UofIMinimal}{
    titlewidthscale=1, bodywidthscale=1, titleleft,
    titleoffsetx=0pt, titleoffsety=0pt, bodyoffsetx=0pt, bodyoffsety=0pt,
    bodyverticalshift=0pt, roundedcorners=0, linewidth=0.2cm,
    titleinnersep=1cm, bodyinnersep=1cm
}{
    \begin{scope}[line width=\blocklinewidth, rounded corners=\blockroundedcorners]
       \ifBlockHasTitle %
           \draw[draw=none,fill=blockbodybgcolor]
               (blockbody.south west) rectangle (blocktitle.north east);
           \draw[color=blocktitlebgcolor, loosely dashed]
               (blocktitle.south west) -- (blocktitle.south east);%
       \else
             \draw[draw=none,fill=blockbodybgcolor]
                 (blockbody.south west) rectangle (blockbody.north east);
        \fi
    \end{scope}
}

\useblockstyle{UofIMinimal}
\definebackgroundstyle{UofIVertical}{
    \draw[line width=0pt, bottom color=backgroundcolor, 
    top color=white] (bottomleft) rectangle (topright);
}
\usebackgroundstyle{UofIVertical}
\definetitlestyle{UofITitle}{
    width=\paperwidth, roundedcorners=0, linewidth=0pt, innersep=12mm,
    titletotopverticalspace=5mm, titletoblockverticalspace=10mm,
    titlegraphictotitledistance=10pt
}{}
\usetitlestyle{UofITitle}

\makeatletter
\renewcommand\TP@maketitle{%
   \begin{minipage}{0.8\textwidth}
        \color{titlefgcolor}
        {\bfseries \Huge \sc \@title \par}
        \vspace*{1em}
        {\huge \@author \par}
        \vspace*{1em}
        {\LARGE \@institute}
    \end{minipage}%
    \hfill
    \begin{minipage}{0.2\linewidth}
      \includegraphics[width=0.8\linewidth,height=6cm,keepaspectratio]{figures/igb_wordmark}
    \end{minipage}
}
\makeatother

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 0)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")
library("scales")
library("magrittr")
library("tidyr")

@

<<source.to.latex,echo=FALSE,message=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R")))>>=
for (a in paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R"))) {
    source(a)
}
@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
star.logs.wide[,input.reads:=as.numeric(input.reads)]
star.logs.wide[,input.read.length:=as.numeric(input.read.length)]
setkey(hugo.nomenclature,"Approved Symbol")
@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

@

<<plot.defaults,echo=FALSE,results="hide">>=
grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)

@

\input{acronyms}

\maketitle

\begin{columns}

  \column{0.25}

  \block{Introduction}{

%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
%     uterus specific. CGB, LGALS13, and members of the PSG family were
%     identified as placenta specific. The accuracy of the classifier
%     was verified using RNAseq transcriptome data not in the training
%     set. We show the applicability of the classifier to simulated EVs
%     by down sampling reads to $10^5$ reads and biasing the reads to
%     account for preferential association of RNA motifs with EV
%     packaging machinery. When coupled with single-EV sequencing, our
%     classifier enables clinicians to monitor disease-specific markers
%     over the course of treatment without biopsies, increasing the
%     ability of clinicians to adapt treatment modalities to the
%     individual patient. 
  }

  \block{Methods}{     
}

\column{0.25}

\begin{figure}
<<species.tree,fig.width=8,fig.height=6,echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Spalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.5,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.5,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.5,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.5,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.5,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.5,node=6,width=0.12)
      ## sheep -- this is a male, the sample is female, but there aren't any females in phylopic
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.5,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.5,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.5,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.5,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.5,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.5,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.5,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.5,node=13,width=0.18)
      )
@ 
\caption{Species tree for organisms studied showing the relationship
  between species. Silhouettes of species were obtained from
  \url{http://phylopic.org}. }
\label{fig:species_tree}
\end{figure}

\column{0.25}

\column{0.25}

\block{Conclusions}{
}

\block{References}{

  \begin{center}
    \mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]
  \end{center}
}

\block{Code and Datasets}{
The code and underlying data for this poster can be found at
\begin{center}
\qrcode[padding]{http://dla2.us/p/pv2016}
\url{http://dla2.us/p/pv2016}
\end{center}
% 
}
\block{}{
  \includegraphics[width=\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
}

\end{columns}

\end{document}