### 2016-04-27 poster/placenta_viviparous_poster.Rnw
add conclusions and switch to KISS1 instead of CGA

\documentclass[a0paper,landscape]{tikzposter}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{fancybox}
\usepackage{xspace}
\usepackage{booktabs}
% \usepackage{multirow}
\usepackage{acronym}
\usepackage{qrcode}
\usepackage[backend=biber,maxnames=1,natbib=true,hyperref=true,style=nature]{biblatex}
\addbibresource{references.bib}
% \usepackage[nomargin,inline,draft]{fixme}
% \newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\renewcommand*{\bibfont}{\small}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.  Use showboxes option to show outlines of textboxes.
% \usepackage[absolute]{textpos}
\usepackage[absolute,showboxes]{textpos}

\usepackage{multirow}
\usepackage{array}

\title{The Evolution of Gene Expression in the Term Placenta of
  Viviparous Mammals}

\author{Don L. Armstrong, Michael R. McGowen, Amy Weckle, Jason
  Caravas, Dalen Agnew, Sue Savage-Rumbaugh, Eviatar Nevo, Chong J.
  Kim, Günter P. Wagner, Roberto Romero, and Derek Wildman}
\institute{Institute for Genomic Biology, Computing Genomes for
  Reproductive Health, University of Illinois, Urbana-Champaign}
\usetheme{Default}

% species names
\newcommand{\smonkey}[0]{\textit{A. fusciceps}\xspace}
\newcommand{\cow}[0]{\textit{B. taurus}\xspace}
\newcommand{\dog}[0]{\textit{C. familiaris}\xspace}
\newcommand{\armadillo}[0]{\textit{D. novemcinctus}\xspace}
\newcommand{\horse}[0]{\textit{E. caballus}\xspace}
\newcommand{\elephant}[0]{\textit{L. africana}\xspace}
\newcommand{\human}[0]{\textit{H. sapiens}\xspace}
\newcommand{\opossum}[0]{\textit{M. domestica}\xspace}
\newcommand{\mouse}[0]{\textit{M. musculus}\xspace}
\newcommand{\galili}[0]{\textit{N. galili}\xspace}
\newcommand{\sheep}[0]{\textit{O. aries}\xspace}
\newcommand{\bonobo}[0]{\textit{P. paniscus}\xspace}
\newcommand{\carmeli}[0]{\textit{S. carmeli}\xspace}
\newcommand{\pig}[0]{\textit{S. scrofa}\xspace}

% use this
% http://www.identitystandards.illinois.edu/graphicstandardsmanual/othermedia/posterpresentation.html
\definecolorstyle{UofI}{
  \definecolor{boldBlue}{HTML}{002058}
  \definecolor{boldOrange}{HTML}{E87722}
  \definecolor{professionalBlue}{HTML}{606EB2}
  \definecolor{professionalOrange}{HTML}{D45D00}
  }{
    \colorlet{backgroundcolor}{professionalBlue}
    \colorlet{framecolor}{black}
    \colorlet{titlefgcolor}{boldBlue}
    \colorlet{titlebgcolor}{white}
    \colorlet{blockbodybgcolor}{white}
    \colorlet{blocktitlebgcolor}{professionalOrange}
    \colorlet{blocktitlefgcolor}{boldBlue}
    }
\usecolorstyle{UofI}
\defineblockstyle{UofIMinimal}{
    titlewidthscale=1, bodywidthscale=1, titleleft,
    titleoffsetx=0pt, titleoffsety=0pt, bodyoffsetx=0pt, bodyoffsety=0pt,
    bodyverticalshift=0pt, roundedcorners=0, linewidth=0.2cm,
    titleinnersep=1cm, bodyinnersep=1cm
}{
    \begin{scope}[line width=\blocklinewidth, rounded corners=\blockroundedcorners]
       \ifBlockHasTitle %
           \draw[draw=none,fill=blockbodybgcolor]
               (blockbody.south west) rectangle (blocktitle.north east);
           \draw[color=blocktitlebgcolor, loosely dashed]
               (blocktitle.south west) -- (blocktitle.south east);%
       \else
             \draw[draw=none,fill=blockbodybgcolor]
                 (blockbody.south west) rectangle (blockbody.north east);
        \fi
    \end{scope}
}

\useblockstyle{UofIMinimal}
\definebackgroundstyle{UofIVertical}{
    \draw[line width=0pt, bottom color=backgroundcolor, 
    top color=white] (bottomleft) rectangle (topright);
}
\usebackgroundstyle{UofIVertical}
\definetitlestyle{UofITitle}{
    width=\paperwidth, roundedcorners=0, linewidth=0pt, innersep=12mm,
    titletotopverticalspace=5mm, titletoblockverticalspace=10mm,
    titlegraphictotitledistance=10pt
}{}
\usetitlestyle{UofITitle}

\makeatletter
\renewcommand\TP@maketitle{%
   \begin{minipage}{0.8\textwidth}
        \color{titlefgcolor}
        {\bfseries \Huge \sc \@title \par}
        \vspace*{1em}
        {\huge \@author \par}
        \vspace*{1em}
        {\LARGE \@institute}
    \end{minipage}%
    \hfill
    \begin{minipage}{0.13\linewidth}
      \begin{flushright}
        \includegraphics[width=\linewidth,height=6cm,keepaspectratio]{figures/igb_wordmark}
      \end{flushright}
    \end{minipage}
}
\makeatother

\begin{document}

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="0.9\\colwidth",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE)
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 0)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")
library("scales")
library("magrittr")
library("tidyr")

@

<<source.to.latex,echo=FALSE,message=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R")))>>=
for (a in paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R"))) {
    source(a)
}
@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
star.logs.wide[,input.reads:=as.numeric(input.reads)]
star.logs.wide[,input.read.length:=as.numeric(input.read.length)]
setkey(hugo.nomenclature,"Approved Symbol")
@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

@

<<plot.defaults,echo=FALSE,results="hide">>=
grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)

@

\input{acronyms}

\maketitle

\begin{columns}

  \column{0.2}

  \block{Introduction}{
\begin{itemize}
\item Function and Evolution of Placenta
    \begin{itemize}
    \item Major function is the transfer of nutrients, oxygen, and
      waste between mother and offspring
    \item Highly variable structure and form in mammals, even though
      major function is conserved
    \item Likely arose before the origin of therian mammals
    \item Form relatively well characterized, but diversity of
      molecular environment understudied
    \item Role in the evolutionary interplay between maternal and fetal
      strategies
    \end{itemize}

  \item Impacts on Human Health
    \begin{itemize}
    \item Humans possess a deeply invasive trophoblast layer that can
      penetrate the myometrium of the uterus
      \cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
    \item Defects in placental growth can lead to pre-eclampsia, which can
      cause hypertension, proteinuria, and mortality of mother and/or
      infant \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
    \item Pre-eclampsia has been seen in \textit{Pan troglodytes}
      \cite{Stout.ea1969:Glomerularcapillaryendothelialswelling}, and
      \textit{Gorilla gorilla}
      \cite{Thornton.ea1992:Convulsionsinpregnancyin}
    \item Far higher incidence in \textit{Homo sapiens}, with as many as 
      4\% of pregnancies United States affected
      \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
    \end{itemize}

  \item Hypotheses
    \begin{itemize}
    \item Gene expression patterns correlate with placenta morphology
    \item Altered expression of genes correlates with invasiveness, and
      therefore pre-eclampsia
    \item Maternal/Fetal evolutionary strategies will be reflected in gene
      expression patterns
    \end{itemize}
  \end{itemize}

%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
%     uterus specific. CGB, LGALS13, and members of the PSG family were
%     identified as placenta specific. The accuracy of the classifier
%     was verified using RNAseq transcriptome data not in the training
%     set. We show the applicability of the classifier to simulated EVs
%     by down sampling reads to $10^5$ reads and biasing the reads to
%     account for preferential association of RNA motifs with EV
%     packaging machinery. When coupled with single-EV sequencing, our
%     classifier enables clinicians to monitor disease-specific markers
%     over the course of treatment without biopsies, increasing the
%     ability of clinicians to adapt treatment modalities to the
%     individual patient. 
  }

  \block{Methods}{
    \small
    \begin{itemize}
    \item Fetal placenta tissue from 9 mammalian species collected;
      total RNA isolated and sequenced using Illumina GA II with an
      average read size of 150nt
    \item Additional mammalian placenta sequences obtained from SRA
    \item QCed using FASTQC, aligned using STAR, and quantified using
      Cufflinks. Species missing suitable reference (Spider monkey,
      \textit{S. galili}, \textit{S. carmeli}) were assembled using
      Trinity and annotated using Diamond against the Human or Mouse
      transcriptome.
    \item Ensembl build 80 was the source of annotation.
    \item Enrichment in human and mouse were calculated using a
      type-III anova, all other enrichments used Student's t test.
    \end{itemize}
  }

\column{0.2}

<<top.genes.function,echo=FALSE,results="hide">>=
top.genes.table <- function(species.name,n=50) {
    alignment.species <- species.name
    if (alignment.species=="pan paniscus") {
        alignment.species <- "pan troglodytes"
    } else if (alignment.species=="ateles fusciceps") {
        alignment.species <- "homo sapiens"
    } else if (grepl("spalax",alignment.species)) {
        alignment.species <- "mus musculus"
    }
    temp <- 
        combined.fpkm[species==species.name &
                      !(!is.na(human_name) & 
                        human_name %in% housekeeping.genes.superset[,gene_short_name]) &
                      ! (gene_biotype %in% c("Mt_tRNA","Mt_rRNA","rRNA")) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?[L]?$)",human_name) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?[L]?$)",human_alignment_symbol) &
                      !grepl("(^[Mm][Tt]-|^R[pP][lsLS]\\d+[abcdABCD]?[lL]?(?:-.+)?$|rRNA$)",
                             gene_short_name)
                      ][order(-mean_fpkm)][,list(name_or_id,
                                                 human_alignment_symbol,
                                                 human_name,mean_fpkm,sd_fpkm,n_fpkm)]
    ## remove any genes which have a null name_or_id; they are probably from unassembled contigs
    temp <- temp[!(is.na(name_or_id)|name_or_id==""),]
    temp <- temp[!duplicated(name_or_id)][1:n]
    temp[,used_alignment:=FALSE]
    temp[is.na(human_name) & !is.na(human_alignment_symbol),
                    used_alignment:=TRUE]
    temp[is.na(human_name),human_name:=human_alignment_symbol]
    setkey(temp,"human_name")
    temp <- 
        hugo.nomenclature[temp][order(-mean_fpkm),
                                list(name_or_id,used_alignment,`Approved Symbol`,`Approved Name`,
                                     mean_fpkm,sd_fpkm,n_fpkm)]
    temp[!is.na(name_or_id),
         name_or_id:=paste0("\\href{http://www.ensembl.org/",
                             capfirst(gsub(" ","_",alignment.species)),
                             "/Gene/Summary?g=",name_or_id,"}{\\textit{",
                             xtable.sanitize.text(name_or_id),"}}")]
    temp[!is.na(`Approved Name`),
         `Approved Name`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Name`),"}")]
    temp[!is.na(`Approved Symbol`),
         `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{\\textit{",xtable.sanitize.text(`Approved Symbol`),"}}")]
    temp[used_alignment==TRUE,`Approved Symbol`:=paste0("(",`Approved Symbol`,")")]
    temp[,used_alignment:=NULL]
    temp[,sem_fpkm:=as.numeric(NA)]
    temp[n_fpkm>1,sem_fpkm:=sd_fpkm/sqrt(n_fpkm)]
    temp[,sd_fpkm:=NULL]
    temp[,n_fpkm:=NULL]
    setnames(temp,c("Gene symbol","Human Symbol","Human Name","FPKM","SEM"))
    if (temp[,all(is.na(`SEM`))]) {
        temp[,`SEM`:=NULL]
    }
    if (species.name=="homo sapiens" || species.name=="ateles fusciceps") {
        temp[,`Human Symbol`:=NULL]
    }
    temp[,`FPKM`:=round(`FPKM`)]
    return(temp)
}
gene.expression <- function(species.name,gene,complete=FALSE,units=TRUE) {
    species.name <- gsub("_"," ",species.name)
    fpkm <- combined.fpkm[species==species.name &
                              (gene_short_name==gene | 
                               gene_id==gene),
                          mean_fpkm][1]
    if (is.na(fpkm)) {
        stop("FPKM is NA!!!!")
    }
    if (units || complete)
        fpkm <- formatC(fpkm,digits=0,format="f")
    if (units) {
        fpkm <- paste0(fpkm,
                       "\\ \\text{FPKM}")
    }
    if (complete) {
        if (any(grepl("^ENS.+G\\d+$",gene))) {
            fpkm <- paste0("\\text{",gene,"} = ",
                           fpkm)
        } else {
            fpkm <- paste0("\\text{\\textit{",gene,"}} = ",
                           fpkm)
        }
    }
    if (units || complete) {
        fpkm <- paste0("$",fpkm,"$")
    }
    return(fpkm)
}

@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific","human_mouse_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
load("../data/human_mouse_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE,lineage.table=placenta.lineage.specific.analysis,members=5,orderbyfc=FALSE,fdr.cutoff=1) {
    temp.table <- 
        lineage.table[analysis.name==analysis &
                      p.value <= 0.05 &
                      fdr <= fdr.cutoff][order(p.value),
                                       list(human_name,p.value,fdr,fc),]
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[`FDR`<=fdr.cutoff,][order(`p (t-test)`),]
        if (orderbyfc) {
            display.table <- temp.table[order(-abs(`Fold Change`)),]
        }
        display.table <- display.table[1:members,]
        display.table[,`p (t-test)`:=NULL]
        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        setnames(display.table,"Fold Change","$\\mathrm{log}_2$ FC")
        print(xtable(display.table,
                     display=c("s","s","s","g","f"),
                     align=c("c","c","p{3.3in}","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}
output.top.genes.table <- function(data,output=TRUE) {
    if (output) {
        table.display <- c("s","s","s","d")
        table.align <- c("c","c","p{3in}","r")
        if (any(colnames(data) %in% "SEM"))  {
            table.display <- c(table.display,"d")
            table.align <- c(table.align,"r")
        }
        if (any(colnames(data) %in% "Human Symbol"))  {
            table.align <- c("c",table.align)
            table.display <- c("s",table.display)
        }
        print(xtable(data,
                     display=table.display,
                     align=table.align
                     ),
              booktabs=TRUE,
              floating=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)
    }
}

@

\block{Phylogeny of Species}{
\centering
<<species.tree,fig.width=10,fig.height=8,echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Spalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.5,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.5,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.5,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.5,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.5,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.5,node=6,width=0.12)
      ## sheep -- this is a male, the sample is female, but there aren't any females in phylopic
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.5,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.5,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.5,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.5,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.5,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.5,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.5,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.5,node=13,width=0.18)
      )
@ 
\captionof{figure}{Species tree for organisms studied showing the relationship
  between species. Silhouettes of species were obtained from
  \url{http://phylopic.org}. }
\label{fig:species_tree}
}

% \block{Variation in Gestational Period and Litter Size}{
%   \small
%   \begin{tabular}{>{\raggedright \centering}p{2in} >{\raggedright \centering}p{2.5in} c c }
%     \toprule
%     Species                       & Common Name & \multicolumn{1}{p{1in}}{\centering Gestation (days)} & \multicolumn{1}{p{1in}}{\centering Average Litter Size} \\
%     \midrule
%     \textit{Ateles fusciceps}     & Spider monkey & 224 & 1 \\
%     \textit{Bos taurus}           & Cow & 278 & 1 \\
%     \textit{Canis familiaris}     & Dog & 61 & 4 \\
%     \textit{Dasypus novemcinctus} & Nine-banded Armadillo & 150 & 4 \\
%     \textit{Equus caballus}       & Horse & 336 & 1 \\
%     \textit{Loxodonta africana}   & African savanna elephant & 644 & 1 \\
%     \textit{Homo sapiens}         & Human & 270 & 1.1 \\
%     \textit{Monodelphis domestica}& Gray short-tailed opossum & 15 & 7.1 \\
%     \textit{Mus musculus}         & Mouse & 20 & 6.1 \\
%     \textit{Spalax galili}   & Blind mole rat from Upper Galilee & 28-42 & 4 \\
%     \textit{Ovis aries}           & Domestic sheep & 151 & 1.35 \\
%     \textit{Pan paniscus}         & Bonobo & 223 & 1 \\
%     \textit{Spalax carmeli}       & Blind mole rat from northern Israel & 28-42 & 4 \\
%     \textit{Sus scrofa}           & Domestic pig (Duroc) & 114 & 5.8 \\
%     \bottomrule
%   \end{tabular}
%   \captionof{table}{Gestational periods in days and average liter sizes for studied species.}
% }

% \block{Variations in Placenta Morphology}{
% \begin{tabular}{c c c}
%   \toprule
%   Species & Interhemal Membrane & Pattern \\
%   \midrule
%   \smonkey & hemochorial & trabecular \\
%   \cow     & epitheliochorial & villous \\
%   \dog     & endotheliochorial & labyrinthine \\
%   \armadillo & hemochorial & trabecular \\
%   \horse & epitheliochorial & villous \\
%   \elephant & endotheliochorial & labyrinthine \\
%   \human & hemochorial & villous \\
% % http://onlinelibrary.wiley.com/doi/10.1002/jez.1402660511/pdf
%   \opossum & epitheliochorial & choriovitelline \\
%   \mouse & hemotrichorial & labyrinthine \\
%   \galili & hemotrichorial & labyrinthine \\
%   \sheep & epitheliochorial & villous \\
%   \bonobo & hemochorial & villous \\
% % http://www.biolreprod.org/content/early/2014/10/17/biolreprod.114.124818.short
%   \carmeli & hemotrichorial & labyrinthine \\
%   \pig & epitheliochorial & diffuse \\
%   \bottomrule
% \end{tabular}
% }

\block{Human-specific Genes}{
{\small
<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=10,
                                    orderbyfc=TRUE,
                                    fdr.cutoff=0.001
                                    )
setkey(hum.all.lin.spec,"Human Symbol")
@
}
\captionof{table}{1:1 Human orthologous genes whose expression
  changed on the human lineage as opposed to all other analyzed
  species.}
}

\block{Primate-specific Genes}{
{\small
<<primate.lineage.specific,echo=FALSE,results="asis">>=
pri.all.lin.spec <- placenta.lineage.specific.table("pri.all")
setkey(pri.all.lin.spec,"Human Symbol")
@ 
}
}

\column{0.2}

\block{Core Placenta Transcriptome}{
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species),]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]
print(ggplot(cptl,
             aes(y=human_name,x=Species))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(),
              axis.text.y = element_text(face="italic"),
              axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic"))
      )
@ 
\captionof{figure}{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  1:1 human ortholog genes expressed (\ac{FPKM}$ \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied with housekeeping genes
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} removed
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_transcriptome}
}

\block{\textit{PRSS16} Expression}{
<<prss16.evolution.tree,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="PRSS16" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 vp=viewport(layout.pos.col=1,layout.pos.row=2,clip="off"),
                 fontsize=4,
                 width=3,
                 axis.text.size=3.5,
                 offset.ratio=1.2
                 )

@ 
\captionof{figure}{\textit{PRSS16} is highly expressed and massively
  expanded in \elephant, but not in \armadillo. It is currently
  unknown whether this high expression and expansion is a shared
  feature afrotherians.}
}

\column{0.2}

\block{\textit{CGA} Expression}{

<<kiss1.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

plot.label.size <- 4.5
tree.id <-
    genes.to.tree[combined.fpkm[gene_short_name=="KISS1" & species=="homo sapiens",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression,
                 offset.ratio=1,
                 fontsize=8,
                 axis.text.size=5
                 )
@ 
\captionof{figure}{\textit{KISS1}, which is implicated in
  pre-eclampsia, is significantly more expressed on the human lineage
  in comparison to all other lineages eamined. }
}

\block{Hormones, Galectins, \textit{IGF} and Pre-eclampsia}{
<<hormone.genes.plot,echo=FALSE,fig.width=8,fig.height=10,warning=FALSE,message=FALSE>>=

gene.expression.nozero <- function(genes) {
    temp <-
        data.matrix.dt(dcast(combined.fpkm[human_name %in% genes,],
                             human_name~species,
                             value.var="FPKM",
                             fun.aggregate=function(x){x[1]}))
    temp <-
        temp[apply(temp,1,function(x){sum(x,na.rm=TRUE)})>5,]
    return(temp)
}

### hormone genes
hormone.genes <- c("CGA","CGB","CGB1","CGB2","CGB5","CGB8","CSH1","CSH2","CSHL1","CYP19A1",
                   "GH2","GHRH","HSD3B1","HSD3B2","HSD17B2","HSD17B1","NPPB","NPPC","PRL","RLN1")
hormone.expression.nozero <-
    gene.expression.nozero(hormone.genes)

hor.exp.long <- melt(log10(hormone.expression.nozero+1)[rev(1:nrow(hormone.expression.nozero)),])
colnames(hor.exp.long)[1:3] <- c("gene","species","fpkm")
hor.exp.long <- hor.exp.long[hor.exp.long$gene!="IGF2",]

### galectin genes
galectin.genes <- unique(grep("^LGALS",combined.fpkm[,human_name],value=TRUE))

galectin.expression.nozero <-
    gene.expression.nozero(galectin.genes)

gal.exp.long <- melt(log10(galectin.expression.nozero+1)[rev(1:nrow(galectin.expression.nozero)),])
colnames(gal.exp.long)[1:3] <- c("gene","species","fpkm")

### IGF related genes

igfrelated.genes <- unique(grep("^IGF",combined.fpkm[,human_name],value=TRUE))

igfrelated.expression.nozero <-
    gene.expression.nozero(igfrelated.genes)

igf.exp.long <- melt(log10(igfrelated.expression.nozero+1)[rev(1:nrow(igfrelated.expression.nozero)),])
colnames(igf.exp.long)[1:3] <- c("gene","species","fpkm")

### Pre-eclampsia genes ADAM12 CRH KISS1 PAPPA

pe.expression.nozero <-
    gene.expression.nozero(c("ADAM12","CRH","KISS1","PAPPA"))

pe.exp.long <- melt(log10(pe.expression.nozero+1)[rev(1:nrow(pe.expression.nozero)),])
colnames(pe.exp.long)[1:3] <- c("gene","species","fpkm")

igf.exp.long$type <- "IGF Related"
gal.exp.long$type <- "Galectin"
hor.exp.long$type <- "Hormone"
pe.exp.long$type <- "PE"

combined.long <-
    rbind(igf.exp.long,
          gal.exp.long,
          hor.exp.long,
          pe.exp.long
          )
combined.long$species <-
    factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(combined.long$species)),
           levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(names(species.names))))

common.plot.options <- 
    list(geom_raster(aes(fill = fpkm)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         ylab("Human Ortholog Name"),
         xlab("Species"),
         theme_grey(),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\captionof{figure}{Expression in $\mathrm{log}_{10}(\mathrm{FPKM})$ of 1:1 \human
  orthologs of Hormone Genes, Galectins, genes implicated in
  pre-eclampsia (PE), and \textit{IGF}-related genes. Empty locations
  indicate no detectable expression and/or no 1:1 \human\ ortholog in
  that species.}
}

\column{0.2}

\block{Conclusions}{
\begin{itemize}
\item Expression resource of
  \Sexpr{star.logs[,length(unique(species))]} species representing the
  breadth of mammalia for future research.

\item Lineage-specific expansions of gene families and differential
  gene expression

\item Components of annexin complexes, including \textit{ANXA2},
  \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10} are expressed
  in all placentas examined, and argue for the evolution importance of
  annexins in placenta function, where they likely establish
  maternal-fetal
  tolerance\cite{Liu.ea2015:AnnexinA2complexeswith}.

\item We identified of multiple genes (\textit{CRH},
  \textit{ADAM12}, \textit{KISS1}, \textit{PAPPA}, \textit{IL1RL1},
  and \textit{PLAC1}) which are associated with pre-eclampsia which
  have increased expression on the \human lineage and may explain
  larger proportion of pre-eclampsia observed in humans.

\item Placenta samples from additional species and time points are
  needed to more completely understand the complex molecular milieu
  of the developing placenta.

\end{itemize}
}

\block{References}{

  \begin{center}
    \mbox{}\vspace{-\baselineskip}
    \printbibliography[heading=none]
  \end{center}
}

\block{}{
The code and underlying data for this poster can be found at
\begin{center}
\qrcode[padding]{http://dla2.us/p/pv2016}
\url{http://dla2.us/p/pv2016}
\end{center}

  \includegraphics[width=\linewidth,height=6cm,keepaspectratio]{figures/uofi_mark}
}

\end{columns}

\end{document}