### 2016-10-04 manuscript/extended_discussion.Rnw
use different figure path; remove author list

\documentclass[english,12pt]{article}
%\documentclass[5p]{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage{acronym}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
% \newenvironment{narrow}[2]{%
%   \begin{list}{}{%
%       \setlength{\topsep}{0pt}%
%       \setlength{\leftmargin}{#1}%
%       \setlength{\rightmargin}{#2}%
%       \setlength{\listparindent}{\parindent}%
%       \setlength{\itemindent}{\parindent}%
%       \setlength{\parsep}{\parskip}}%
%   \item[]}{\end{list}}
% \def\newblock{\hskip}
% \newenvironment{paperquote}{%
%   \begin{quote}%
%      \it
%   }%
%   {\end{quote}}
% \renewcommand{\textfraction}{0.15}
% \renewcommand{\topfraction}{0.85}
% \renewcommand{\bottomfraction}{0.65}
% \renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
% \newenvironment{enumerate*}%
%   {\begin{enumerate}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{enumerate}}
% \newenvironment{itemize*}%
%   {\begin{itemize}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{itemize}}
% \setcounter{tocdepth}{3}
% \oddsidemargin 0.0in 
% \textwidth 6in
% \raggedbottom
% \clubpenalty = 10000
% \widowpenalty = 10000
% set a large emergency stretch to avoid issues with overfull boxes
% \setlength{\emergencystretch}{20em}
% make the afill smaller
% \renewcommand\Affilfont{\itshape\small}
%\renewcommand\cite[1]{\autocite{#1}}
%\renewcommand\citet[1]{\textcite{#1}}
% \pagestyle{fancy}

\input{common_latex}

\begin{document}
% \raggedright

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(dev="cairo_pdf",out.width="\\columnwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(cache=TRUE, autodep=TRUE,cache.path="extended_discussion_cache/")
opts_chunk$set(fig.path="extended_discussion_figure/")
options(device = function(file, width = 8, height = 7, ...) {
  cairo_pdf(tempfile(), width = width, height = height, ...)
})
options(scipen = -2, digits = 0)
library("data.table")
library("ggplot2")
library("reshape2")
library("xtable")
library("grid")
library("gridExtra")
library("ape")
library("ggtree")
library("topGO")
library("org.Hs.eg.db")
library("scales")
library("magrittr")
library("tidyr")
library("phytools")
@

<<source.to.latex,echo=FALSE,message=FALSE,results="hide",cache.extra=file.info(paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R")))[,"mtime"]>>=
for (a in paste0("../common_r_code/",c("to_latex.R","plot_tree_matrix.R","utility_functions.R"))) {
    source(a)
}
@

<<load.data,echo=FALSE,results="hide",cache.extra=file.info(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))[,"mtime"]>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
star.logs.wide[,input.reads:=as.numeric(input.reads)]
star.logs.wide[,input.read.length:=as.numeric(input.read.length)]
setkey(hugo.nomenclature,"Approved Symbol")
@

<<load.go.results,echo=FALSE,results="hide",cache.extra=tools::md5sum("../go_annotation/go_tissue_analysis")>>=
load("../go_annotation/go_tissue_analysis")

@

<<load_plot_ancestral_trees,message=FALSE,echo=FALSE,results="hide",cache.extra=file.info(paste0("../common_r_code/",c("plot_ancestral_state_tree.R")))[,"mtime"]>>=
for (a in paste0("../common_r_code/",c("plot_ancestral_state_tree.R"))) {
    source(a)
}
@

<<plot.defaults,echo=FALSE,results="hide">>=
#grad.def <- list(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = NA)
grad.def <- list(low = "gray90", high = "blue", space = "Lab", na.value = NA)

@

% Acronyms
\input{acronyms}

\title{The evolution of the transcriptome of mammalian placentas}%

\section{Background}

The main role of the placenta in viviparous mammals is the transfer of
nutrients, oxygen, and waste between mother and offspring. Although
these functions are conserved among mammals, the placenta is perhaps
the most variable organ in terms of structure and form
\cite{Mossman.1937:comparativemorphogenesisofthe,Moffett.ea2006:Immunologyofplacentationin}. Species differ in
the shape of attachment (i.e., discoid, diffuse, cotyledonary,
zonary), degree of intimacy between fetal tissues and maternal tissues
(i.e. epitheliochorial, endotheliochorial, hemochorial), and degree of
interdigitation of maternal-fetal tissue (i.e., villous, folded,
trabecular, lamellar, labrynthine)
\cite{Mossman.1987:vertebratefetalmembranes}. In addition, therian
mammals differ drastically in the number and size of their offspring,
as well as the length of gestation, traits often associated with
specific aspects of placental morphology
\cite{Capellini.ea2011:Placentationandmaternalinvestment}
(\cref{tab:gestational_periods}).

The placenta arose at some point before the origin of extant mammals,
a group that includes marsupials, monotremes, and eutherian or
“placental mammals”. The term “placental mammals” is misleading, as
extant monotremes (such as \textit{Ornithorhynchus anatinus}, the
platypus) possess a rudimentary placenta, as do marsupials, although
the direct apposition to maternal tissues lasts for a proportionally
shorter time during gestation
\cite{Selwood.ea2006:Trophoblastandhypoblastin,Renfree2010:ReviewMarsupialsplacentalmammals}.
Marsupials also differ from eutherian mammals in that most possess
only a choriovitelline placenta, produced from the fusion of the
chorion and yolk sac, during most of gestation. Although some
eutherian mammals may possess a choriovitelline placenta at some point
in development, the major site of nutrient exchange in placentals
occurs via the chorioallantoic placenta, produced from the fusion of
the chorion and allantois
\cite{Mossman.1987:vertebratefetalmembranes}. It is unclear whether
the first placenta was more like that of marsupials or eutherians;
however, using phylogenetic methods, the ancestral eutherian was
reconstructed as possessing a relatively invasive hemochorial
placenta, as in humans and mouse
\cite{Elliot.ea2009:Phylogeneticevidenceforearly,Wildman.ea2006:Evolutionofmammalianplacenta}.

Although the morphology of the mammalian placenta has been well
characterized in a number of species
\cite{Mossman.1987:vertebratefetalmembranes}, the diversity of the
molecular environment is only beginning to be understood from a
genetic perspective \cite{Carter2012:Evolutionofplacentalfunction}.
Studies of certain model species (i.e., \textit{Homo sapiens},
\textit{Mus musculus}, \textit{Bos taurus}) have revealed a stunning
divergence in the genetics of the placenta, especially in later stages
of pregnancy
\cite{Carter2012:Evolutionofplacentalfunction,Hou.ea2012:Elephanttranscriptomeprovidesinsights,Knox.ea2008:Genomicevolutionofplacenta,Su.ea2002:Large-scaleanalysisofhuman}.
For example, each lineage has evolved expansive placenta-specific gene
families with expression patterns that can vary in space and time
\cite{Green.ea2000:Pregnancy-associatedbovineandovine,Simmons.ea2008:Spatialandtemporalexpression}.
These include independent expansions of the prolactins in \mouse,
\textit{Spalax carmeli}, \textit{Spalax galili}, and \cow, \acp{PSG} in \human and \mouse, \acp{PAG}
in \cow, genes related to the thymus-specific serine protease
\textit{PRSS16} in \textit{Loxodona africana} (in this work),
 and the expansion of the growth hormone
and chorionic gonadotropin beta genes in anthropoid primates
\cite{Haig2008:Placentalgrowthhormone-relatedproteins,Rawn.ea2008:evolutionregulationandfunction}.
Expression patterns of specific genes also can vary widely between
species. For example, the expression of orthologous genes in the
placenta have been found to vary substantially between \human and
\mouse \cite{Su.ea2002:Large-scaleanalysisofhuman}, and the
recruitment of \textit{SIGLEC6} to placental expression in \human
differs from both \textit{Pan troglodytes} and \mouse
\cite{Linden.ea2007:Human-specificexpressionofSiglec-6}.

One explanation for the divergent paths of the placental interface is
the distinct evolutionary interplay between maternal and fetal
strategies during pregnancy \cite{Haig1993:Geneticconflictsinhuman}.
The maternal strategy is to requisition an adequate degree of
nutrients to the fetus without risking the health of the mother, while
the fetal strategy via paternal genes is to extract as many resources
from the mother as possible. Therefore, an interplay between these
extremes can set up an intricate “arms race”, involving the imprinting
of parental alleles, expansion of gene families, and enlistment of
existing genes to expression at the maternal-fetal interface
\cite{Haig1993:Geneticconflictsinhuman}.

One symptom of this maternal-fetal conflict is its potential affect on
health
\cite{Hollegaard.ea2013:Parent-offspringconflictandpersistence}. In
addition to having an invasive hemochorial placenta, \human possess a
deeply invasive trophoblast layer that can penetrate the myometrium of
the uterus \cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
Defects in placental growth are known to be associated with the
devastating condition of pre-eclampsia, which can cause hypertension,
proteinuria, and has the potential to lead to death of the mother
and/or infant \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
Although cases have been documented in \textit{P. troglodytes}
\cite{Stout.ea1969:Glomerularcapillaryendothelialswelling}, and
\textit{Gorilla gorilla}
\cite{Thornton.ea1992:Convulsionsinpregnancyin}, pre-eclampsia has a
far higher incidence in \human pregnancies with as much as 4\% of
pregnancies being affected in the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. \textit{P.
  troglodytes} and \textit{G. gorilla} also have deeply invasive
trophoblasts, although the progression of invasion in great apes may
be slower than \human
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}. However,
large-scale expression studies of great ape placentas have not been
conducted. The increasing affordability of next-generation sequencing
technology has enabled the large-scale study of specific tissues from
a wide array of species to address broader phylogenetic questions in
the evolution of gene expression
\cite{Brawand.ea2011:evolutionofgeneexpression}.

Here we use RNA-Seq
technology to sequence and quantify the mRNA environment of mammalian
placentas at term, including that of the \human, \textit{Pan paniscus}
(the sister species to \textit{P. troglodytes}), \textit{Ateles
  fusciceps}, \mouse, \galili, \carmeli, \textit{Canis familiaris},
\cow, \elephant, \textit{Dasypus novemcinctus}, and
\textit{Monodelphis domesticus}. These species represent four major
branches of eutherian mammals (Euarchontoglires, Laurasiatheria,
Afrotheria, and Xenartha) as well as marsupials. These placentas also represent a
wide array of placental morphologies as well as traits such as number
and size of offspring and length of gestation
(\cref{tab:gestational_periods}). We use this data, along with
existing RNA-Seq data from \mouse
\cite{Wang.ea2011:surveyfornovelimprinted}, \human
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual}, \textit{Equus caballus}
\cite{Coleman.ea2010:Structuralannotationofequine}, \textit{Ovis aries}
\cite{Archibald.ea2010:sheepgenomereferencesequence}, and \textit{Sus scrofa}
\cite{Ahanda.ea2012:Predictionofaltered3-} to address the major
differences between expression in marsupial and eutherian mammals at
term, identify the evolution of expression in important placental
genes such as hormones and galectins, identify novel lineage-specific
expression patterns, and quantify \human specific differences in gene
expression that may be related to obstetrical disorders such as
pre-eclampsia.

\section{Methods}

\subsection{Collection of Placental Tissue and RNA Extraction}

Fetal tissue was collected from the placenta of nine mammalian
species (\human [human], \bonobo [bonobo], \smonkey [spider monkey], \dog [domestic dog], \cow
[domestic cow], \elephant [African savanna elephant], \opossum [gray
short-tailed opossum], \galili [blind mole rat from Upper Galilee],
\carmeli [blind mole rat from northern Israel]
\cite{Nevo2001:adaptiveradiationofblindsubterraneanmole}). Information
on specimen ID, and collector is shown for each sample in Supplemental
\cref{tab:srr_numbers}. In the case of \bonobo, \cow, and \elephant,
the placenta was collected at term upon expulsion from the mother
following delivery. Placental tissue from \human was collected upon
delivery via a cesarean section. Near-term placental tissue from \dog
was collected upon the death of the mother from a car accident.
Placental tissue from \opossum was sampled at fetal stage 32, roughly
one day prior to birth. Only the fetal portion of the placenta was
sampled at the maternal-fetal interface (i.e. villous portion for
\human, \bonobo, \dog, and \elephant; cotyledon for \cow; diffuse
placenta for \opossum). All tissues were collected and stored in
RNAlater (Life Technologies, Carlsbad, CA) until RNA isolation could
be performed. Tissues were homogenized in Trizol, and total RNA was
isolated following the manufacturer's instructions. RNA was further
purified using RNeasy Midi Kit (Qiagen, Valencia, CA), in conjunction
with the RNase-Free DNase Set (Qiagen, Valencia, CA), following the
manufacturer’s protocol to remove any contaminating DNA from the
sample. Total RNA was then evaluated on a Nanodrop 1000 (Thermo
Scientific, Wilmington, DE) and an Agilent Bioanalyzer 2100 (Santa
Clara, CA) to confirm concentration and quality of the isolation. RNA
quality was assessed using A260/A280 and A260/A230 absorbance values.

\subsection{Sequencing}

All RNA-Seq mRNA isolation, library construction, and Illumina
sequencing was performed by the Applied Genomics Technology Center
(AGTC) at Wayne State University School of Medicine using 2.5-5 µg of
total RNA and the standard Illumina protocol. Paired-end sequencing
was performed using the Illumina Genome Analyzer II with an insert
size of 230 base pairs (bp). Read lengths for species are shown in
Supplemental \cref{tab:ref_genome_aligned_species}.

\subsubsection{Additional Sequences}

For more complete evolutionary comparisons additional sequences of
placenta in \mouse
(n=\Sexpr{star.logs[grepl("musculus",species),length(unique(file))]})
\cite{Wang.ea2011:surveyfornovelimprinted}, \human
(n=\Sexpr{star.logs[grepl("sapiens",species),length(unique(file))]})
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual} \horse
\cite{Coleman.ea2010:Structuralannotationofequine}, \sheep
\cite{Archibald.ea2010:sheepgenomereferencesequence}, and \pig
\cite{Ahanda.ea2012:Predictionofaltered3-} were obtained from the
\ac{SRA} (Supplemental \cref{tab:srr_numbers}) and analyzed using the
same methodology as RNAs sequenced specifically for this study.

\subsection{Transcriptome Assembly, Alignment, and Quantification}

The quality of reads was evaluated using FASTQC v0.11.2
(\url{http://www.bioinformatics.babraham.ac.uk/projects/fastqc/}).
Transcriptomes without a suitable reference genome (\smonkey, \galili,
\carmeli; italic reference in Supplemental
\cref{tab:ref_genome_aligned_species}) were assembled using Trinity
v2014-07-17 \cite{Haas.ea2013:Denovotranscriptsequence} and annotated
using Diamond \cite{Buchfink.ea2015:Fastandsensitiveprotein} against
either the \human or \mouse transcriptome (Supplemental
\cref{tab:ref_genome_aligned_species}). For annotation using Diamond,
the match with the highest bitscore was chosen as the annotation;
transcripts where the highest scoring match had a bitscore below 40
were left unannotated.
% 
All transcriptomes were then aligned to the appropriate reference (or
assembled) genome using STAR v2.4.2a
\cite{Dobin.ea2013:STARultrafastuniversalRNA-seq} and expression
values in \ac{FPKM}
\cite{Trapnell.ea2010:Transcriptassemblyandquantification} were
quantified using Cufflinks v2.2.1
\cite{Trapnell.ea2009:TopHatdiscoveringsplicejunctions} (Supplemental
\cref{tab:all_ateles_fusciceps_placenta,tab:all_homo_sapiens_placenta,tab:all_pan_paniscus_placenta,tab:all_mus_musculus_placenta,tab:all_spalax_galili_placenta,tab:all_spalax_carmeli_placenta,tab:all_bos_taurus_placenta,tab:all_ovis_aries_placenta,tab:all_sus_scrofa_placenta,tab:all_equus_caballus_placenta,tab:all_canis_familiaris_placenta,tab:all_loxodonta_africana_placenta,tab:all_dasypus_novemcinctus_placenta,tab:all_monodelphis_domestica_placenta}).
%
The genome of the closely-related \textit{P. troglodytes} was used as
a reference for \bonobo. The separation between the two recognized Pan
species is generally perceived to be $\approx 1$ million years
\cite{Prüfer.ea2012:bonobogenomecomparedwith}. %
%

\begin{figure}
<<species.tree,fig.width=8,fig.height=6,out.width="\\columnwidth",echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Spalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.3,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.3,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.3,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.3,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.3,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.3,node=6,width=0.12)
      ## sheep -- this is a male, the sample is female, but there aren't any females in phylopic
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.3,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.3,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.3,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.3,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.3,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.3,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.3,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.3,node=13,width=0.18)
      )
@ 
\caption{Species tree for organisms studied showing the relationship
  between species. Silhouettes of species were obtained from
  \url{http://phylopic.org}. }
\label{fig:species_tree}
\end{figure}

\subsection{Exploration of Individual Transcriptome Data}

To determine enrichment levels of functional categories of genes in
highly-expressed transcripts in each transcriptome, we performed
functional annotation cluster analyses per transcriptome using topGO
\cite{topGO:2.20.0}. Gene Ontology clusters were considered
significantly enriched if the \ac{FDR} as estimated by the Benjamini
and Hochberg procedure \cite{Benjamini1995:bh_procedure} on from the
$p$ values obtained from the Fisher Exact Test were less than or equal
to 0.05. Enrichment tests were considered with two gene lists per
transcriptome: annotated genes with expression levels of the top 5\%
of the frequency distribution of \ac{FPKM} (Supplemental
\cref{tab:go_terms_enriched_top_one_dasypus_novemcinctus,tab:go_terms_enriched_top_one_equus_caballus,tab:go_terms_enriched_top_one_homo_sapiens,tab:go_terms_enriched_top_one_loxodonta_africana,tab:go_terms_enriched_top_one_monodelphis_domestica,tab:go_terms_enriched_top_one_mus_musculus,tab:go_terms_enriched_top_one_ovis_aries,tab:go_terms_enriched_top_one_pan_paniscus,tab:go_terms_enriched_top_one_sus_scrofa,tab:go_terms_enriched_top_one_ateles_fusciceps,tab:go_terms_enriched_top_one_spalax_galili,tab:go_terms_enriched_top_one_spalax_carmeli,tab:go_terms_enriched_top_one_bos_taurus,tab:go_terms_enriched_top_one_canis_familiaris});
the top 100 genes with the highest \ac{FPKM} (Supplemental
\cref{tab:go_terms_enriched_dasypus_novemcinctus,tab:go_terms_enriched_equus_caballus,tab:go_terms_enriched_homo_sapiens,tab:go_terms_enriched_loxodonta_africana,tab:go_terms_enriched_monodelphis_domestica,tab:go_terms_enriched_mus_musculus,tab:go_terms_enriched_ovis_aries,tab:go_terms_enriched_pan_paniscus,tab:go_terms_enriched_sus_scrofa,tab:go_terms_enriched_ateles_fusciceps,tab:go_terms_enriched_spalax_galili,tab:go_terms_enriched_spalax_carmeli,tab:go_terms_enriched_bos_taurus,tab:go_terms_enriched_canis_familiaris}).
Due to the paucity of annotated gene ontology terms for some species,
gene lists were compared using the \human Gene Ontology using the 1:1
\human ortholog. Enrichment was considered out of the set of all 1:1
orthologs present in each species.

\subsection{Comparative Analysis of Gene Expression}

All 1:1 orthologs present among \opossum, \textit{Pan troglodytes},
\cow, \dog, \elephant, \mouse, and \human were determined using
Ensembl v.80 gene trees via Biomart
\cite{Kasprzyk2011:BioMartdrivingparadigmchange} using the biomaRt R
package
\cite{Durinck.ea2009:Mappingidentifiersforintegration,Durinck.ea2005:BioMartandBioconductorpowerful}.
A total of \Sexpr{combined.fpkm[species!="homo sapiens" & mean_fpkm >=
  1,length(unique(human_name))]} 1:1 orthologous genes were expressed
in at least one species. Of these, \Sexpr{combined.fpkm[species!="homo
  sapiens" & mean_fpkm >= 100,length(unique(human_name))]} 1:1
orthologous genes with \ac{FPKM} $\ge 100$ were retained for future
analyses of gene expression evolution described below.

Significant shifts in expression levels between groups of species were
calculated using Student's t test on 1:1 orthologs using the mean of
each species to avoid large biases towards samples with $n \gg 1$ and
considered significant if the \ac{FDR} as estimated by the Benjamini
and Hochberg procedure \cite{Benjamini1995:bh_procedure} was
$\le 0.05$. For \human and \mouse lineage specific calculations, a
type-III anova was calculated using
lme4~\cite{Bates.ea2015:FittingLinearMixed-EffectsModels} with each
species having separate random effects.

\subsection{Sequences, FPKMs, and Code}

All of the code used to produce every analysis presented in this paper
is available at
\url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of the
sequences used are deposited in SRA, and new sequences for this work
can be obtained under the series
\href{http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE79121}{GSE79121}
(Supplemental \cref{tab:srr_numbers}). The mean FPKM for gene in each
species with ortholog annotations are located in the supplemental data
file all\_species\_mean\_fpkm.txt.xz and the per-sample FPKM values are in
all\_species\_per\_sample\_fpkm.txt.xz.

\section{Results}

\subsection{Transcriptome Statistics}

Supplemental \cref{fig:transcriptome_statistics} shows the RNAseq read and
alignment statistics for all species. For transcriptomes sequenced
with 75 bp %
%
(\Sexpr{species(star.logs.wide[input.read.length > 70 & input.read.length < 80,unique(species)])}),%
%
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,max(as.numeric(input.reads))]}
with an average of \Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,mean(`% good reads`)]}\% of reads 
mapping to the respective reference genome (\cref{tab:ref_genome_aligned_species}). 
%
For transcriptomes sequenced with 150 bp %
% The mismatch here is because we're reporting trimmed reads here, not
% full length reads
(\Sexpr{species(star.logs.wide[input.read.length > 120 & input.read.length < 160,unique(species)])}),
%
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length > 120 & input.read.length < 160,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length  > 120 & input.read.length < 160,max(as.numeric(input.reads))]}
%
with an average of \Sexpr{star.logs.wide[input.read.length > 120 & input.read.length < 160,mean(`% good reads`)]}\% %
of reads mapping to the
respective reference genome (\cref{tab:ref_genome_aligned_species}).
%
%
For transcriptomes sequenced with 35 bp %
% The mismatch here is because we're reporting trimmed reads here, not
% full length reads
(\Sexpr{species(star.logs.wide[input.read.length < 40 ,unique(species)])}),
%
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length < 40 ,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length  < 40  ,max(as.numeric(input.reads))]}
%
with an average of \Sexpr{star.logs.wide[input.read.length < 40, mean(`% good reads`)]}\%
of reads mapping to the
respective reference genome (\cref{tab:ref_genome_aligned_species}).
The expression of all genes with \ac{FPKM} $\ge 50$ is shown in Supplemental
\cref{tab:all_ateles_fusciceps_placenta,tab:all_homo_sapiens_placenta,tab:all_pan_paniscus_placenta,tab:all_mus_musculus_placenta,tab:all_spalax_galili_placenta,tab:all_spalax_carmeli_placenta,tab:all_bos_taurus_placenta,tab:all_ovis_aries_placenta,tab:all_sus_scrofa_placenta,tab:all_equus_caballus_placenta,tab:all_canis_familiaris_placenta,tab:all_loxodonta_africana_placenta,tab:all_dasypus_novemcinctus_placenta,tab:all_monodelphis_domestica_placenta}.

<<top.genes.function,echo=FALSE,results="hide">>=
top.genes.table <- function(species.name,n=50) {
    alignment.species <- species.name
    if (alignment.species=="pan paniscus") {
        alignment.species <- "pan troglodytes"
    } else if (alignment.species=="ateles fusciceps") {
        alignment.species <- "homo sapiens"
    } else if (grepl("spalax",alignment.species)) {
        alignment.species <- "mus musculus"
    }
    temp <- 
        combined.fpkm[species==species.name &
                      !(!is.na(human_name) & 
                        human_name %in% housekeeping.genes.superset[,gene_short_name]) &
                      ! (gene_biotype %in% c("Mt_tRNA","Mt_rRNA","rRNA")) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?[L]?$)",human_name) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?[L]?$)",human_alignment_symbol) &
                      !grepl("(^[Mm][Tt]-|^R[pP][lsLS]\\d+[abcdABCD]?[lL]?(?:-.+)?$|rRNA$)",
                             gene_short_name)
                      ][order(-mean_fpkm)][,list(name_or_id,
                                                 human_alignment_symbol,
                                                 human_name,mean_fpkm,sd_fpkm,n_fpkm)]
    ## remove any genes which have a null name_or_id; they are probably from unassembled contigs
    temp <- temp[!(is.na(name_or_id)|name_or_id==""),]
    temp <- temp[!duplicated(name_or_id)][1:n]
    temp[,used_alignment:=FALSE]
    temp[is.na(human_name) & !is.na(human_alignment_symbol),
                    used_alignment:=TRUE]
    temp[is.na(human_name),human_name:=human_alignment_symbol]
    setkey(temp,"human_name")
    temp <- 
        hugo.nomenclature[temp][order(-mean_fpkm),
                                list(name_or_id,used_alignment,`Approved Symbol`,`Approved Name`,
                                     mean_fpkm,sd_fpkm,n_fpkm)]
    temp[!is.na(name_or_id),
         name_or_id:=paste0("\\href{http://www.ensembl.org/",
                             capfirst(gsub(" ","_",alignment.species)),
                             "/Gene/Summary?g=",name_or_id,"}{\\textit{",
                             xtable.sanitize.text(name_or_id),"}}")]
    temp[!is.na(`Approved Name`),
         `Approved Name`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Name`),"}")]
    temp[!is.na(`Approved Symbol`),
         `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{\\textit{",xtable.sanitize.text(`Approved Symbol`),"}}")]
    temp[used_alignment==TRUE,`Approved Symbol`:=paste0("(",`Approved Symbol`,")")]
    temp[,used_alignment:=NULL]
    temp[,sem_fpkm:=as.numeric(NA)]
    temp[n_fpkm>1,sem_fpkm:=sd_fpkm/sqrt(n_fpkm)]
    temp[,sd_fpkm:=NULL]
    temp[,n_fpkm:=NULL]
    setnames(temp,c("Gene symbol","Human Symbol","Human Name","FPKM","SEM"))
    if (temp[,all(is.na(`SEM`))]) {
        temp[,`SEM`:=NULL]
    }
    if (species.name=="homo sapiens" || species.name=="ateles fusciceps") {
        temp[,`Human Symbol`:=NULL]
    }
    temp[,`FPKM`:=round(`FPKM`)]
    return(temp)
}
@

<<gene.expression,results="hide",echo=FALSE>>=
mentioned.genes <- load_cache(label="save_mentioned_genes",object="saved_mentioned_genes",list())
gene.expression <- function(species.name,gene,complete=FALSE,units=TRUE,mention.only=FALSE) {
    species.name <- gsub("_"," ",species.name)
    mentioned.genes[[paste0(species.name,"_",gene)]] <<-
        list("species.name"=species.name,
             "gene"=gene)
    if (mention.only) {
        return("")
    }
    fpkm <- combined.fpkm[species==species.name &
                              (gene_short_name==gene | 
                               gene_id==gene),
                          mean_fpkm][1]
    if (is.na(fpkm)) {
        stop("FPKM is NA!!!!")
    }
    if (units || complete)
        fpkm <- formatC(fpkm,digits=0,format="f")
    if (units) {
        fpkm <- paste0(fpkm,
                       "\\ \\text{FPKM}")
    }
    if (complete) {
        if (any(grepl("^ENS.+G\\d+$",gene))) {
            fpkm <- paste0(gsub("(ENS)(.+)(G)0*(\\d+)","\\1\\\\linebreak[0]\\2\\\\linebreak[0]\\3...\\4",gene)," = ",
                           fpkm)
        } else {
            fpkm <- paste0("\\textit{",gene,"} = ",
                           fpkm)
        }
    }
    ## if (units || complete) {
    ##     fpkm <- paste0("$",fpkm,"$")
    ## }
    return(fpkm)
}

@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific","human_mouse_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
load("../data/human_mouse_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE,lineage.table=placenta.lineage.specific.analysis,members=5,orderbyfc=FALSE,fdr.cutoff=1) {
    temp.table <- 
        lineage.table[analysis.name==analysis &
                      p.value <= 0.05 &
                      fdr <= fdr.cutoff][order(p.value),
                                       list(human_name,p.value,fdr,fc),]
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[`FDR`<=fdr.cutoff,][order(`p (t-test)`),]
        if (orderbyfc) {
            display.table <- temp.table[order(-abs(`Fold Change`)),]
        }
        display.table <- display.table[1:members,]

        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        display.table[,`Human Name`:=NULL]
        # setnames(display.table,"Human Symbol","\\multicolumn{1}{p{0.5in}}{Human Symbol}")6
        setnames(display.table,"Fold Change","$\\text{log}_2$ FC")
        print(xtable(display.table,
                     display=c("s","s","e","g","f"),
                     align=c("c","p{0.5in}","r","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}
output.top.genes.table <- function(data,output=TRUE) {
    if (output) {
        table.display <- c("s","s","s","d")
        table.align <- c("c","c","p{3in}","r")
        if (any(colnames(data) %in% "SEM"))  {
            table.display <- c(table.display,"d")
            table.align <- c(table.align,"r")
        }
        if (any(colnames(data) %in% "Human Symbol"))  {
            table.align <- c("c",table.align)
            table.display <- c("s",table.display)
        }
        print(xtable(data,
                     display=table.display,
                     align=table.align
                     ),
              booktabs=TRUE,
              floating=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)
    }
}

@

<<mentioned.genes.table.functions,echo=FALSE,results="asis">>=
mentioned.genes.table <- list()
if (length(mentioned.genes)==0) {
    mentioned.genes.table[["No genes"]] <-
        data.table(species="No genes",
                   name_or_id="No genes",
                   human_alignment_symbol="NOGENE",
                   human_name="NONAME",
                   mean_fpkm = 0,
                   sd_fpkm = 1,
                   n_fpkm = 2)
}
for (entry in names(mentioned.genes)) {
    mentioned.genes.table[[entry]] <-
        combined.fpkm[species==mentioned.genes[[entry]][["species.name"]] &
                      (gene_short_name==mentioned.genes[[entry]][["gene"]] | 
                       gene_id==mentioned.genes[[entry]][["gene"]]),
                      list(species,
                           name_or_id,
                           human_alignment_symbol,
                           human_name,
                           mean_fpkm,
                           sd_fpkm,
                           n_fpkm)]
}
mentioned.genes.table <-
    rbindlist(mentioned.genes.table)
mentioned.genes.table[,used_alignment:=FALSE]
mentioned.genes.table[is.na(human_name) & !is.na(human_alignment_symbol),
                      used_alignment:=TRUE]
mentioned.genes.table[is.na(human_name),human_name:=human_alignment_symbol]
setkey(mentioned.genes.table,"human_name")
mentioned.genes.table <- 
    hugo.nomenclature[mentioned.genes.table][,
                                             list(species,
                                                  name_or_id,
                                                  used_alignment,
                                                  `Approved Symbol`,
                                                  `Approved Name`,
                                                  mean_fpkm,sd_fpkm,n_fpkm)]
mentioned.genes.table[,alignment.species:=species]
mentioned.genes.table[alignment.species=="pan paniscus",alignment.species:="pan troglodytes"]
mentioned.genes.table[alignment.species=="ateles fusciceps",alignment.species:="homo sapiens"]
mentioned.genes.table[grepl("spalax",alignment.species),alignment.species:="mus musculus"]
mentioned.genes.table[!is.na(name_or_id),
                      name_or_id:=paste0("\\href{http://www.ensembl.org/",
                                         capfirst(gsub(" ","_",alignment.species)),
                                         "/Gene/Summary?g=",name_or_id,"}{\\textit{",
                                         gsub("(ENS[^0-9]+)","\\1\\\\linebreak[0]",
                                              xtable.sanitize.text(name_or_id)),"}}")]
mentioned.genes.table[!is.na(`Approved Name`),
                      `Approved Name`:=paste0("\\href{http://www.ensembl.org/",
                                              "Homo_sapiens/Gene/Summary?g=",
                                              `Approved Symbol`,"}{",
                                              xtable.sanitize.text(`Approved Name`),
                                              "}")]
mentioned.genes.table[!is.na(`Approved Symbol`),
                      `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/",
                                                "Homo_sapiens/Gene/Summary?g=",
                                                `Approved Symbol`,
                                                "}{\\textit{",
                                                xtable.sanitize.text(`Approved Symbol`),
                                                "}}")]
mentioned.genes.table[used_alignment==TRUE,`Approved Symbol`:=paste0("(",`Approved Symbol`,")")]
mentioned.genes.table[,sem_fpkm:=as.numeric(NA)]
mentioned.genes.table[n_fpkm>1,sem_fpkm:=sd_fpkm/sqrt(n_fpkm)]
mentioned.genes.table[,used_alignment:=NULL]
mentioned.genes.table[,alignment.species:=NULL]
mentioned.genes.table[,sd_fpkm:=NULL]
mentioned.genes.table[,n_fpkm:=NULL]
mentioned.genes.table[,`Approved Name`:=NULL]
### order by species and then by FPKM
setorder(mentioned.genes.table,`Approved Symbol`,species,-mean_fpkm)
### italicize species names
mentioned.genes.table[,species:=paste0("\\textit{",gsub("^(.)\\S+","\\U\\1.",species,perl=TRUE),"}")]
setnames(mentioned.genes.table,c("Species","Gene Symbol",
                                 "Human Symbol",
                                 ## "Human Name",
                                 "FPKM","SEM"))
mentioned.genes.table[,`FPKM`:=round(`FPKM`)]

output.mentioned.genes.table <- function(subset.table=quote(TRUE),
                                         human.symbol=TRUE) {
    mgt.subset <- mentioned.genes.table[eval(subset.table)]
    mgt.display <- c("s","s","s","s","d","d")
    mgt.align <- c("c","c",
                   ">{\\raggedright \\centering}p{0.5in}",
                   ">{\\raggedright \\centering}p{0.5in}",
                   ##">{\\raggedright}p{4in}",
                   "r","r")
    if (!human.symbol) {
        mgt.subset[,`Human Symbol`:=NULL]
        mgt.display <- mgt.display[-4]
        mgt.align <- mgt.align[-4]
        mgt.align[3] <- ">{\\raggedright \\centering}p{1.2in}"
    }

    print(xtable(mgt.subset,
                 display=mgt.display,
                 align=mgt.align
                 ),
          math.style.exponents=TRUE,
          booktabs=TRUE,
          floating=FALSE,
          sanitize.text.function=genes.table.sanitize,
          include.rownames=FALSE)
}
@

\subsection{Primate Placenta Transcriptomes}

\begin{table}[htb]
\footnotesize
<<primate.genes,echo=FALSE,results="asis">>=
output.mentioned.genes.table(subset.table=quote(grepl("sapiens|fusciceps|paniscus",Species)),
                             human.symbol=TRUE
                             )
@ 
\caption{Notable genes ordered by species and \ac{FPKM}. Gene
  symbol is the gene symbol or Ensembl gene id. Human Name is the HUGO
  name corresponding to the human symbol. A complete table of 
  genes with \ac{FPKM} $\ge 50$ is located in the supplemental material }
\label{tab:notable_primate_genes}
\end{table}

\begin{table}[htb]
  \footnotesize
<<primate.lineage.specific,echo=FALSE,results="asis">>=
pri.all.lin.spec <- placenta.lineage.specific.table("pri.all")
setkey(pri.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanation{primates}{all other analyzed species}}
\label{tab:primate_lineage_specific}
\end{table}

We examined the expression of three primates, \human, \bonobo, and
\smonkey. The genes whose expressions are different on the primate
lineage in comparison to all other analyzed species are listed in
\cref{tab:primate_lineage_specific}. \textit{FSTL1} has significantly
higher expression in the primate lineage
(Supplemental \cref{fig:fstl1_evolution_tree},
$\text{FDR}=\Sexpr{pri.all.lin.spec["FSTL1",`FDR`]}$), is implicated
in BMP signaling, and has previously been implicated in inflammation
and invasiveness
\cite{Mouillet.ea2015:expressionandpost-transcriptionalregulation}. Of
non-primates, \horse shows the highest expression
%
\Sexpr{gene.expression(species.name="equus caballus",gene="FSTL1",complete=TRUE)}, 
%
which is still lower than the lowest primate, \smonkey
%
(\Sexpr{gene.expression(species.name="ateles fusciceps",gene="FSTL1",complete=TRUE)}). 
%
No other genes were significantly differentially expressed on this
lineage by t-test.

\subsubsection{\human Placenta Transcriptome}
\label{sec:human_placenta}

% \begin{table}
%   \footnotesize
% <<top.50.human.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="homo sapiens"))
% @ 
% \caption{The top 50 genes by \ac{FPKM} in \human which are not
%   annotated as Mt\_rRNA, Mt\_tRNA, rRNA or a housekeeping gene. Gene
%   symbol is the gene symbol or Ensembl gene id. Human Name is the HUGO
%   name corresponding to the human symbol. A complete list of all genes
%   is in \cref{tab:all_homo_sapiens_placenta}.}
% \label{tab:top50_human_placenta}
% \end{table}

<<cgb.min.expression,results="hide",echo=FALSE,warnings=FALSE>>=
cgb.min.expression <- 
    min(sapply(paste0("CGB",c("","1","2","5","7","8")),
               function(x){gene.expression(species.name="homo sapiens",gene=x,units=FALSE)}))
cgb.sum.expression <- 
    sum(sapply(paste0("CGB",c("","1","2","5","7","8")),
               function(x){gene.expression(species.name="homo sapiens",gene=x,units=FALSE)}))
@

The gene with by far the highest \ac{FPKM} value
%
(\Sexpr{gene.expression(species.name="homo sapiens",gene="CSH1")}) %
%
in \human is \textit{CSH1}, followed by \textit{CSH2}, and \textit{CSHL1}, all
three of which are members of a distinct expansion of the growth
hormone gene family within catarrhines (Supplemental \cref{fig:galectins_evolution_tree}). Another member of the growth
hormone gene family, \ac{GH2} (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="GH2")}), %
is also expressed to a high degree in the \human placenta
(Supplemental \cref{tab:all_homo_sapiens_placenta}).
% 
Eleven \acp{PSG} are highly-expressed
(\ac{FPKM}$\ge 100$) in the term placenta of \human, with \textit{PSG1} showing the
greatest degree of expression %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG1")}) % 
and the pseudogene \textit{PSG10P} showing the least %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG10P")}, \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~A) . %
These
genes are members of a hominoid-specific expansion of the \ac{CEA} family
and known to be expressed by the syncytiotrophoblast
\cite{Rawn.ea2008:evolutionregulationandfunction,Lei.ea1992:Cloningandexpressionof}. %
%
Genes involved in the production of estrogens (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP19A1",complete=TRUE)},
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD17B1",complete=TRUE)}, and
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)}%
% 
) and progesterone (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)},
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP11A1",complete=TRUE)}, and
\Sexpr{gene.expression(species.name="homo sapiens",gene="ADM",complete=TRUE)})
%
show exceptionally high expression in the human placenta, as well as %
%
\textit{CGA} (FPKM=\Sexpr{gene.expression(species.name="homo sapiens",gene="CGA")}, %
%
which generates the alpha subnit of multiple human chorionic
gonadotropin hormones (Supplemental \cref{tab:all_homo_sapiens_placenta}). The genes
producing the beta subunits of the chorionic gonadotropins
(\textit{CGB}, \textit{CGB1}, \textit{CGB2}, \textit{CGB5},
\textit{CGB7}, \textit{CGB8}) have reasonable expression
($\text{FPKM} \ge \Sexpr{cgb.min.expression}$), and the sum of their
expression is on the order of the expression of
\textit{CGA} ($\sum \mathit{CGBs}=\Sexpr{cgb.sum.expression}$ FPKM) .
%
%
Other highly expressed genes in the top 50 include members of the
hemoglobin family (e.g. \textit{HBA1}, \textit{HBA2}, \textit{HBB},
and \textit{HBG2}), the maternally imprinted long noncoding RNA gene
\textit{H19}, genes coding for proteins required for growth and/or
cell proliferation (\textit{ADAM12}, \textit{KISS1}, \textit{IGF2},
\textit{GPC3}, \textit{CSF3R}), anticoagulation (\textit{TFPI2}), and
glucose transport (\textit{SLC2A1}) among others
(Supplemental \cref{tab:all_homo_sapiens_placenta}, Supplemental
\cref{tab:all_homo_sapiens_placenta}).

% \paragraph{Pre-eclampsia \human Lineage Specific Genes}

\begin{table}
  \footnotesize
<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=40,
                                    orderbyfc=TRUE,
                                    fdr.cutoff=0.001
                                    )
setkey(hum.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanationtypeiii{\human}{all other analyzed species}}
\label{tab:human_lineage_specific}
\end{table}

Using a mixed-effects linear model to account for repeated testing of
\human and \mouse, we identified multiple genes which are
differentially expressed in the \human lineage in comparison to all
other lineages (\cref{tab:human_lineage_specific}). These genes
include multiple genes which are thought to be involved in
pre-eclampsia. Notably, \textit{CRH}
(FDR=\Sexpr{hum.all.lin.spec["CRH",`FDR`]},
FC=\Sexpr{hum.all.lin.spec["CRH",`Fold Change`]}, Supplemental
\cref{fig:crh_evolution_tree}), has been shown to be more highly
expressed in placentas from pre-eclamptic pregnancies
\cite{Petsas.ea2012:Aberrantexpressionofcorticotropin-releasing,Paiva.ea2011:MeasurementofmRNAtranscripts},
where it is hypothesized to activate decidual macrophages. Likewise,
\textit{KISS1} (\Sexpr{hum.all.lin.spec["KISS1",`FDR`]},
FC=\Sexpr{hum.all.lin.spec["KISS1",`Fold Change`]}, Supplemental
\cref{fig:kiss1_evolution_tree}) also shows increased expression in
pre-eclamptic placentas
\cite{Vazquez-Alaniz.ea2011:Comparativeexpressionprofilesfor}, and is
involved in modulating trophoblast invasion. Two other genes,
\textit{PAPPA} (\Sexpr{hum.all.lin.spec["PAPPA",`FDR`]},
FC=\Sexpr{hum.all.lin.spec["PAPPA",`Fold Change`]} Supplemental
\cref{fig:pappa_evolution_tree}) and \textit{ADAM12}
(\Sexpr{hum.all.lin.spec["ADAM12",`FDR`]},
FC=\Sexpr{hum.all.lin.spec["ADAM12",`Fold Change`]}, Supplemental
\cref{fig:adam12_evolution_tree}) are also differentially expressed
but are negatively correlated with pre-eclampsia
\cite{Spencer.ea2008:ADAM12sinmaternalserum}.

\begin{figure*}[htb]
<<psg.ctsl.evolution.tree,echo=FALSE,out.width="\\textwidth",fig.width=8,fig.height=8>>=
pushViewport(viewport(layout=grid.layout(nrow=2,ncol=2),clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="PSG1" & species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                subtree=c("ENSG00000221826","ENSG00000124469"),
                                trees=trees
                                )+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,0.2)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="CTSL" & species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                subtree=c("ENSMODG00000002844","ENSMUSG00000055679"),
                                trees=trees)+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,0.8)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="PRSS16" &
                                                    species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                subtree=c(   "ENSG00000112812",
                                          "ENSLAFG00000032211"),
                                combined.fpkm=combined.fpkm,
                                trees=trees)+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,1.2),
      vp=viewport(layout.pos.col=1,layout.pos.row=2,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="PAG2" &
                                                    species=="bos taurus"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                subtree=c("ENSSSCG00000013097",
                                          "ENSBTAG00000046085"),
                                trees=trees) +
      geom_tiplab(size=2,color="black",fontface="italic") +
      xlim(0,0.8)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=2,layout.pos.row=2,clip="off"))
popViewport(1)

grid.text(c("A","B","C","D"),
          x=unit(c(0,0.5,0,0.5),"npc"),y=unit(c(1,1,0.5,0.5),"npc"),
          just=c("left","top"),
          gp=gpar(fontsize="24")
          )
@ 
\caption{\genetreeexplanation{(A) \textit{PSG1}, (B) \textit{CTSL}, (C) \textit{PRSS16}, and (D) \textit{PAG2}}}
\label{fig:pag2_prss16_ctsl_psg_evolution_tree}
\end{figure*}

\subsubsection{\bonobo Placenta Transcriptome}

% \begin{table}
%   \footnotesize
% <<top.50.bonobo.expressed.genes,echo=FALSE,results="asis">>=
% top.bonobo.genes <- top.genes.table(species.name="pan paniscus")
% output.top.genes.table(top.bonobo.genes)
% setkey(top.bonobo.genes,"Gene symbol")
% @ 
% \caption{\tableexplanation{\bonobo}{tab:all_pan_paniscus_placenta}}
% \label{tab:top50_bonobo_placenta}
% \end{table}

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \bonobo are shown in Supplemental \cref{tab:all_pan_paniscus_placenta}
with a list of all genes in Supplemental
\cref{tab:all_pan_paniscus_placenta}. Expression in the \bonobo
placenta differs somewhat from that of \human, as evidenced by the
makeup of the top 50 genes (Supplemental \cref{tab:all_pan_paniscus_placenta}). The highest expression
 in \bonobo is %
\Sexpr{gene.expression("pan paniscus","PRG2",complete=TRUE)}, %
%
which is involved in insulin growth factor signaling and angiogenesis
during pregnancy \cite{Weyer.ea2011:Placentalregulationofpeptide}.
\textit{PRG2} is also expressed at relatively high levels (though
comparatively much lower) in the human placenta
(\Sexpr{gene.expression("homo sapiens","PRG2")};
Supplemental \cref{tab:all_homo_sapiens_placenta}). Four \bonobo genes
related to the chorionic somatomammotropins
are also expressed at relatively high levels (\textit{GH1} 
\Sexpr{gene.expression("pan paniscus","ENSPTRG00000034560",complete=TRUE)}, 
\textit{GH1} 
\Sexpr{gene.expression("pan paniscus","ENSPTRG00000034212",complete=TRUE)}, 
\textit{GH2}
\Sexpr{gene.expression("pan paniscus","ENSPTRG00000030805",complete=TRUE)}, 
and \textit{GH2}
\Sexpr{gene.expression("pan paniscus","ENSPTRG00000034209",complete=TRUE)}, 
%
\cref{tab:all_pan_paniscus_placenta}, \cref{fig:csh1_evolution_tree}).
Placental expression of the genes involved in estrogen and
progesterone production (\textit{CYP19A1}, \textit{HSD3B1}
respectively) show lower expression in \bonobo as compared to \human %
(\Sexpr{gene.expression("pan paniscus","CYP19A1",complete=TRUE)},
\cref{fig:estrogen_evolution_tree}; %
\Sexpr{gene.expression("pan paniscus","HSD3B1",complete=TRUE)},
\cref{fig:progesterone_evolution_tree} and
Supplemental \cref{tab:all_homo_sapiens_placenta}) although the expression of
\textit{HSD3B1} is fairly similar to that of \smonkey %
(\Sexpr{gene.expression("ateles fusciceps","HSD3B1",complete=TRUE)}. %
In addition, \textit{CGA} is expressed in the bonobo placenta %
(\Sexpr{gene.expression("pan paniscus","CGA")}), but at much
lower levels than \human as well (Supplemental \cref{tab:all_homo_sapiens_placenta}).
%
Two Ensembl-annotated beta subunits of the chorionic gonadotropin
genes are expressed in \bonobo, one related to \textit{CGB} and
another related to \textit{CGB7}; both are not highly expressed
(\cref{fig:cga_cgb_evolution_tree}). Among the other top 50 expressed
genes, there is high expression of keratins (e.g., \textit{KRT19},
\textit{KRT18}), heat-shock proteins (e.g., \textit{HSP90B1},
\textit{HSPA1A}), serpins (\textit{SERPINE1}, \textit{SERPINE2},
Supplemental \cref{fig:serpine_evolution_tree}), insulin growth factor
2 (\textit{IGF2}) and one of its binding proteins \textit{IGFBP3}, and
the hormone adrenomedullin (\textit{ADM}).

\subsubsection{\smonkey Transcriptome}

% \begin{table}
%   \footnotesize
% <<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
% top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
% output.top.genes.table(top.spider.monkey.genes)
% setkey(top.spider.monkey.genes,"Gene symbol")
% @ 
% \caption{\tableexplanation{\smonkey}{tab:all_ateles_fusciceps_placenta}}
% \label{tab:top50_spider_monkey_placenta}
% \end{table}

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(Supplemental \cref{tab:all_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles_fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles_fusciceps","PSG8")}), %
only one member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles_fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
Supplemental \cref{tab:all_spider_monkey_placenta}), and multiple components of
annexin complexes
(\Sexpr{gene.expression("ateles_fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}.
\Sexpr{gene.expression("ateles_fusciceps","AL365202.1",complete=TRUE)}
is also highly expressed, has no known function, and is duplicated
multiple times in \textit{Macaca mulatta} (Ensembl gene tree
\href{http://useast.ensembl.org/Multi/GeneTree/Image?gt=ENSGT00580000081348}{ENSGT00580000081348}).
All genes with expression greater than 50 \ac{FPKM} are listed in
Supplementary \cref{tab:all_ateles_fusciceps_placenta}.

\subsection{Rodentia Placenta Transcriptomes}

\begin{table}
  \footnotesize
<<rodentia.genes,echo=FALSE,results="asis">>=
output.mentioned.genes.table(subset.table=quote(grepl("musculus|galili|carmeli",Species)),
                             human.symbol=TRUE
                             )
@ 
\caption{Notable genes ordered by species and \ac{FPKM}. Gene
  symbol is the gene symbol or Ensembl gene id. Human Name is the HUGO
  name corresponding to the human symbol. A complete table of 
  genes with \ac{FPKM} $\ge 50$ is located in the supplemental material }
\label{tab:notable_non_primate_genes}
\end{table}

We examined the expression of three rodents, \mouse, \carmeli, and
\galili{}. We did not detect any significant changes on the Rodentia
lineage in comparison to all other analyzed species (Supplemental
\cref{tab:rodentia_lineage_specific}), likely because of a lack power
to overcome the multiple testing correction. Additional species and/or
samples in each of these lineages will be necessary to detect lineage
specific shifts.

\subsubsection{\mouse Placenta Transcriptome}

% \begin{table}[h]
%   \footnotesize
% <<top.50.mouse.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="mus musculus"))
% @ 
% \caption{\tableexplanation{\mouse}{tab:all_mus_musculus_placenta}}
% \label{tab:top50_mouse_placenta}
% \end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \mouse are shown in
Supplemental \cref{tab:all_mus_musculus_placenta} with a list of all genes with
$\text{FPKM} \ge 50$ in Supplemental
\cref{tab:all_mus_musculus_placenta}. The top expressed genes include
\Sexpr{gene.expression("mus_musculus","Mir6236",complete=TRUE)}, whose
target \textit{in silico} is the orphan G-coupled receptor
\textit{Mrgprg}
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget},
\Sexpr{gene.expression("mus_musculus","H19",complete=TRUE)} and
\Sexpr{gene.expression("mus_musculus","Gm27786",complete=TRUE)}, both
of which are maternally imprinted (the later of which is a homolog of
human \textit{H19\_2}),
\Sexpr{gene.expression("mus_musculus","Igf2",complete=TRUE)}, numerous
cathepsins and a rodentia specific expansion of prolactin related
genes. There are also two miRNA which overlap with the of the 7S
signal recognition particle
(\Sexpr{gene.expression("mus_musculus","AC157822.1",complete=TRUE)}
and
\Sexpr{gene.expression("mus_musculus","AC099934.1",complete=TRUE)})
and numerous pseudogenes
(\Sexpr{gene.expression("mus_musculus","Gm11966",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10736",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10180",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm29216",complete=TRUE)}) which
appear to be highly expressed, however, the short read length (34--44)
in many samples is likely preventing unambiguous assignment of all
reads to the more lengthy complete transcript, so these shorter
transcripts should be treated with suspicion, and we will ignore them
in further discussion. The prolactin-related genes
(\Sexpr{gene.expression("mus_musculus","Prl3b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl2b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl8a9",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Prl8a6",complete=TRUE)}) are
very highly expressed in \mouse in, with %
\Sexpr{combined.fpkm[species=="mus musculus" &
  grepl("Prl(\\d.\\d)?",gene_short_name), sum(mean_fpkm >= 1)]}
prolactin related genes having $\text{FPKM} \ge 1$
(\cref{fig:prolactin_evolution_tree}, Supplemental
\cref{fig:prolactin_receptor_evolution_tree}). The cathespins have
also expanded in rodents, with six showing high expression
(\Sexpr{gene.expression("mus_musculus","Ctsj",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts6",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsr",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsq",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts3",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Ctsm",complete=TRUE)},
Supplemental \cref{tab:all_mus_musculus_placenta,fig:pag2_prss16_ctsl_psg_evolution_tree}).
Finally, \textit{Peg10}, an imprinted gene required for proper
formation of the placenta in mice is also highly expressed
(\Sexpr{gene.expression("mus_musculus","Peg10")}), as previously
reported \cite{Ono.ea2006:DeletionofPeg10imprinted}.

% \paragraph{\mouse Lineage Specific Genes}

\begin{table}
  \footnotesize
<<mouse.lineage.specific,echo=FALSE,results="asis">>=
mou.all.lin.spec <-
    placenta.lineage.specific.table("mou.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=40,
                                    orderbyfc=TRUE,
                                    fdr.cutoff=0.001
                                    )
setkey(mou.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanationtypeiii{\mouse}{all other analyzed species}}
\label{tab:mouse_lineage_specific}
\end{table}

Using a mixed-effects linear model to account for repeated testing of
\human and \mouse, we identified multiple genes which are
differentially expressed in the \mouse lineage in comparison to all
other lineages (\cref{tab:mouse_lineage_specific}). \mouse
\textit{Gkn2} (FDR=\Sexpr{mou.all.lin.spec["GKN2",`FDR`]},
FC=\Sexpr{mou.all.lin.spec["GKN2",`Fold Change`]}) has been shown to
have a pattern of expression distinct from the \human pattern, and has
been suggested to be important for placental development
\cite{Fahlbusch.ea2015:Trophoblastexpressiondynamicsof}. Notable
mouse-specific increases in expression include include genes like
\textit{Slc38a4} (FDR=\Sexpr{mou.all.lin.spec["SLC38A4",`FDR`]},
FC=\Sexpr{mou.all.lin.spec["SLC38A4",`Fold Change`]}) and \textit{Afp}
(FDR=\Sexpr{mou.all.lin.spec["AFP",`FDR`]},
FC=\Sexpr{mou.all.lin.spec["AFP",`Fold Change`]}) which are implicated
in fetal growth restriction
\cite{Sharma.ea2016:Tpbpa-Cre-mediateddeletionofTFAP2C} and placenta
abnormalities in humans
\cite{Waller.ea1996:associationbetweenmaternalserum}, and
\textit{Cnx2} (FDR=\Sexpr{mou.all.lin.spec["CNX2",`FDR`]},
FC=\Sexpr{mou.all.lin.spec["CNX2",`Fold Change`]}) which is critical
for the exchange of metabolites in the placenta labyrinth
\cite{Gabriel.ea1998:Transplacentaluptakeofglucose}.

\subsubsection{\galili Placenta Transcriptome}

% \begin{table}[h]
%   \footnotesize
% <<top.50.n.galili.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="spalax galili"))
% @ 
% \caption{\tableexplanation{\galili}{tab:all_spalax_galili_placenta}}
% \label{tab:top50_n_galili_placenta}
% \end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \galili are shown in
Supplemental \cref{tab:all_n_galili_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_spalax_galili_placenta}.
Components of annexin complexes are highly expressed
(\Sexpr{gene.expression("spalax_galili","S100a6",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","S100a11",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","S100a10",complete=TRUE)},
and
\Sexpr{gene.expression("spalax_galili","Anxa2",complete=TRUE)}),
as are members of the thymosin family
(\Sexpr{gene.expression("spalax_galili","Tmsb4x",complete=TRUE)}
and
\Sexpr{gene.expression("spalax_galili","Tmsb10",complete=TRUE)}),
various cytoskeleton related proteins
(\Sexpr{gene.expression("spalax_galili","Actc1",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Actg2",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Acta2",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Myl9",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Vim",complete=TRUE)}), and
transmembrane protein of currently unknown function, 
\Sexpr{gene.expression("spalax_galili","Gm10134",complete=TRUE)}.

\subsubsection{\carmeli Placenta Transcriptome}

% \begin{table}[h]
%   \footnotesize
% <<top.50.s.carmeli.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="spalax carmeli"))
% @ 
% \caption{\tableexplanation{\carmeli}{tab:all_spalax_carmeli_placenta}}
% \label{tab:top50_s_carmeli_placenta}
% \end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \carmeli are shown in
Supplemental \cref{tab:all_s_carmeli_placenta} with a list of all highly
expressed genes in Supplemental
\cref{tab:all_spalax_carmeli_placenta}. Numerous hemoglobin subunits
ware highly expressed
(\Sexpr{gene.expression("spalax_carmeli","Hbb-bh1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Hba-a1",complete=TRUE)}, and
\Sexpr{gene.expression("spalax_carmeli","Hbb-bs",complete=TRUE)}), as
well as the relaxin
\Sexpr{gene.expression("spalax_carmeli","Rln1",complete=TRUE)},
multiple cathespins
(\Sexpr{gene.expression("spalax_carmeli","Ctsm",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","BC051665",complete=TRUE)},
and \Sexpr{gene.expression("spalax_carmeli","Ctsl",complete=TRUE)},
\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~B), and many prolactin-related genes
(\Sexpr{gene.expression("spalax_carmeli","Prl4a1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Prl6a1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Prl3b1",complete=TRUE)}, and
\Sexpr{gene.expression("spalax_carmeli","Prl7c1",complete=TRUE)},
\cref{fig:prolactin_evolution_tree}) which shows a similar pattern of
expansion seen in \mouse.

\subsection{Laurasiatheria}

We examined the expression of five laurasiatherians: \cow, \sheep,
\pig, \horse, and \dog. We were unable to detect any significant
differences after correcting for multiple testing between
laurasiatherians and all other analyzed species
(Supplemental \cref{tab:laurasiatheria_lineage_specific}) or between
laurasiatherians and Euarchontoglires
(Supplemental \cref{tab:laurasiatheria_euarchontoglires_lineage_specific}).

\subsubsection{\dog Placenta Transcriptome}

% \begin{table}
%   \footnotesize
% <<top.50.dog.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="canis familiaris"))
% @ 
% \caption{\tableexplanation{\dog}{tab:all_canis_familiaris_placenta}}
% \label{tab:top50_dog_placenta}
% \end{table}

The top 50 non-housekeeping expressed genes as measured by \ac{FPKM}
in the term placenta of \dog are shown in
Supplemental \cref{tab:all_canis_familiaris_placenta} with a list of all genes in Supplemental
\cref{tab:all_canis_familiaris_placenta}. The gene with the highest
expression %
(\Sexpr{gene.expression("canis familiaris","TMSB10")}) is the thymosin 
\textit{TMSB10} %
(\Sexpr{gene.expression("canis familiaris","ENSCAFG00000002115")}) %
is the hormone relaxin (\textit{RLN2}). Other highly expressed genes
among the top 50 (Supplemental \cref{tab:all_canis_familiaris_placenta}) include genes
involved in the cytoskeleton (\textit{TMSB10}, \textit{TMSB4X}),
anticoagulation (\textit{TFPI2}, \textit{PLA2G7}), innate immunity
(\textit{LCN2}, \textit{GM2A}), lipid metabolism (\textit{APOE}),
inhibition of matrix metalloproteinases (\textit{TIMP2},
\textit{TIMP3}), and growth factors (\textit{PGF}, \textit{EGFR},
\textit{HTRA1}). The gene \textit{LGALS1}, which shows evidence of
immunosupression during pregnancy
\cite{Than.ea2008:Emergenceofhormonaland}, is highly expressed %
(\Sexpr{gene.expression("canis familiaris","LGALS1")}). %
In addition, an ortholog of \textit{SSLP1} (ENSCAFG00000010638) a gene
expressed in the placentas of \cow (%
\Sexpr{gene.expression("bos taurus","SSLP1",complete=TRUE)}, %
Supplemental \cref{tab:all_bos_taurus_placenta}) and \sheep
(\Sexpr{gene.expression("ovis aries","SOLD1",complete=TRUE)},
Supplemental \cref{tab:all_sheep_placenta}) but is not present in primates
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}
(\cref{fig:sslp1_evolution_tree}), is also expressed at high levels
%
(\Sexpr{gene.expression("canis familiaris","ENSCAFG00000010638")}). %
There is no placental expression of the estrogen synthase gene
(\textit{CYP19A1}) in \dog 
(\Sexpr{gene.expression("canis familiaris","CYP19A1")}
\cref{fig:estrogen_evolution_tree}), 
and expression of \textit{HSD3B7} is also low
(\Sexpr{gene.expression("canis familiaris","HSD3B7")}, 
\cref{fig:progesterone_evolution_tree}).

\subsubsection{\cow Placenta Transcriptome}

% \begin{table}
%   \footnotesize
% <<top.50.cow.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="bos taurus"))
% @ 
% \caption{\tableexplanation{\cow}{tab:all_bos_taurus_placenta}}
% \label{tab:top50_cow_placenta}
% \end{table}

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the \cow are shown in Supplemental \cref{tab:all_bos_taurus_placenta} with a
list of all genes in Supplemental \cref{tab:all_bos_taurus_placenta}.
The gene with the highest expression was \textit{TMSB10} %
(\Sexpr{gene.expression("bos taurus","TMSB10")}), and
\textit{TMSB4} was also highly expressed %
(\Sexpr{gene.expression("bos taurus","TMSB4")}). As in
\dog{}, members of annexin complexes were also highly expressed
(\Sexpr{gene.expression("bos taurus","S100G",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","S100A2",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","S100A11",complete=TRUE)}), as well as %
at least one of a set of four miRNA with 100\% identity
(\Sexpr{gene.expression("bos taurus","ENSBTAG00000047046",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000047154",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000045710",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000046351",complete=TRUE)}). 
%
As these are novel miRNA, there is a possibility that the assignment
of the short reads to these four miRNA is due to miss-annotation of
the \cow genome.
% 
Several members of the ruminant-specific pregnancy-associated
glycoprotein family \cite{Green.ea2000:Pregnancy-associatedbovineandovine} 
show especially high expression in \cow placenta
(Supplemental \cref{tab:all_bos_taurus_placenta}, \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~D, and
Supplemental \cref{tab:all_bos_taurus_placenta}) including
%
\Sexpr{gene.expression("bos taurus","PAG2",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PAG10",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000022759",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000013214",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000019226",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","ENSBTAG00000024611",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PAG12",complete=TRUE)}. %
%
28 members of this family have $\text{FPKM} \ge 1$, and this family of
genes has greatly expanded in ruminants (\cow and \sheep{}) although
the expression levels of this family are higher in \cow than in \sheep
(\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~D). Other genes with high expression
include genes involved with regulation of growth (\textit{IGF2},
\textit{PHLDA2}), imprinted genes (\textit{DLK1}, \textit{PHLDA2}),
and \textit{CSH1}, a gene that is a member of a ruminant specific
expansion of placental prolactins and not directly related to the gene
of the same name in hominids \ref{fig:prolactin_evolution_tree}. Other
ruminant-specific placental prolactins also show high expression in
the cow placenta including %
\Sexpr{gene.expression("bos taurus","PRP1",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP5",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP6",complete=TRUE)}, %
\Sexpr{gene.expression("bos taurus","PRP8",complete=TRUE)}, and %
\Sexpr{gene.expression("bos taurus","PRP7",complete=TRUE)}, %
with a total of 14 with $\text{FPKM} \ge 1$.
%
\textit{SSLP1} (\textit{SOLD1} in \sheep{}), a gene expressed in the
cotyledonary villi of the ruminiant placenta and involved in cell
proliferation
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}, also
shows evidence of high expression %
(\Sexpr{gene.expression("bos taurus","SSLP1")}). %
There is some placental expression of the estrogen synthase gene
(\textit{CYP19A1}) in cows %
(\Sexpr{gene.expression("bos taurus","CYP19A1")}), and
comparatively greater expression of \textit{HSD3B} %
(\Sexpr{gene.expression("bos taurus","HSD3B")}).

\subsubsection{\pig Placenta Transcriptome}

% \begin{table}[h]
%   \footnotesize
% <<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="sus scrofa"))
% @ 
% \caption{\tableexplanation{\pig}{tab:all_sus_scrofa_placenta}}
% \label{tab:top50_pig_placenta}
% \end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
Supplemental \cref{tab:all_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}. The top two
expressed transcripts,
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000025162",complete=TRUE)}
and
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000021356",complete=TRUE)},
are miRNA which correspond to the \cow miRNA
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MI0013080}{bta-mir-2904-1}
and
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MIMAT0013845}{bta-mir-2887-1},
respectively, both of which are also expressed in \cow ovary
\cite{Hossain.ea2009:Identificationandcharacterizationof}, although
neither of which have an obvious target gene on the basis of
\textit{in silico} analysis
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}. Other highly
expressed genes include members of the insulin growth factor family
(\Sexpr{gene.expression("sus_scrofa","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("sus_scrofa","IGFBP3",complete=TRUE)}), members
of annexin complexes
(\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","ANXA2",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A12",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","S100A10",complete=TRUE)}), the
galectin \Sexpr{gene.expression("sus_scrofa","LGALS3",complete=TRUE)}
and multiple cytoskeleton and extracellular matrix proteins
(\Sexpr{gene.expression("sus_scrofa","KRT19",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","COL1A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","BETA-ACTIN",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","CLDN4",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)}).

\subsubsection{\horse Placenta Transcriptome}

% \begin{table}[h]
%   \footnotesize
% <<top.50.horse.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="equus caballus"))
% @ 
% \caption{\tableexplanation{\horse}{tab:all_equus_caballus_placenta}}
% \label{tab:top50_horse_placenta}
% \end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \horse are shown in
Supplemental \cref{tab:all_equus_caballus_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of annexin complexes
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.

\subsubsection{\sheep Placenta Transcriptome}

% \begin{table}[h]
%   \footnotesize
% <<top.50.sheep.expressed.genes,echo=FALSE,results="asis">>=
% top.sheep.genes <- top.genes.table(species.name="ovis aries",n=200)
% top.sheep.genes <-
%     top.sheep.genes[!grepl("\\{(7SK|.+_([SM]RP|nuc)|SNOR[AD]\\d+|U\\d+|HIST.*)\\}",
%                            `Gene symbol`) &
%                     !grepl("\\((7SK|.+_([SM]RP|nuc)|SNOR[AD]\\d+|U\\d+|HIST.*)\\)",
%                            `Human Symbol`)
%                    ,][1:50,]
% output.top.genes.table(top.sheep.genes)
% @ 
% \caption{\tableexplanation{\sheep}{tab:all_ovis_aries_placenta}}
% \label{tab:top50_sheep_placenta}
% \end{table}

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \sheep are shown in
Supplemental \cref{tab:all_sheep_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_ovis_aries_placenta}. The complete
list of highly expressed genes includes many snRNAs and rRNAs likely
due to the RNA extraction and library preparation procedure utilized
for this sample. Ignoring these, notable highly expressed genes
include members of annexin complexes
(\Sexpr{gene.expression("ovis_aries","S100G",complete=TRUE)},
\Sexpr{gene.expression("ovis_aries","ANXA2",complete=TRUE)}),
hemoglobin subunits
(\Sexpr{gene.expression("ovis_aries","ENSOARG00000010944",complete=TRUE)}),
an ortholog of \textit{SSLP1},
\Sexpr{gene.expression("ovis_aries","SOLD1",complete=TRUE)}, two
cathespins (\Sexpr{gene.expression("ovis_aries","CTSL",complete=TRUE)}
and \Sexpr{gene.expression("ovis_aries","CTSB",complete=TRUE)}), and
placentally expressed transcript 1
\Sexpr{gene.expression("ovis_aries","PLET1",complete=TRUE)} which may
play a role in cellular migration
\cite{Raymond.ea2010:Expressionoforphanprotein}.

\subsection{Atlantogenata (Afrotheria + Xenartha)}
\subsubsection{\elephant Placenta Transcriptome}

% \begin{table}
% \footnotesize
% <<top.50.elephant.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="loxodonta africana"))
% @ 
% \caption{\tableexplanation{\elephant}{tab:top50_elephant_placenta}}
% \label{tab:top50_elephant_placenta}
% \end{table}

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \elephant are shown in Supplemental \cref{tab:all_loxodonta_africana_placenta}
with a list of all genes in
\cref{tab:all_loxodonta_africana_placenta}. The gene with the highest
expression was the natriuretic peptide precursor B
(\Sexpr{gene.expression("loxodonta_africana","NPPB",complete=TRUE)}).
Two thymosins,
\Sexpr{gene.expression("loxodonta_africana","TMSB10",complete=TRUE)}
and
\Sexpr{gene.expression("loxodonta_africana","TMSB4Y",complete=TRUE)}
were also also was highly expressed. The miRNA mir-675
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000027406",complete=TRUE)}).
which is imbedded in the first exon of the maternally imprinted long
non-coding RNA gene \textit{H19}
\cite{Keniry.ea2012:H19lincRNAisdevelopmental}, a gene which is
currently un-annotated in \elephant{}. %
% \DLA{Probably consider fixing the unannotation of H19; it seems pretty
%   obvious to me, though it would require fixing the GTF file for
%   elephant with the right
%   \href{http://useast.ensembl.org/Loxodonta_africana/Transcript/Summary?db=core;pt=GENSCAN00000034833;r=scaffold_145:855987-899506}{genscan
%     prediction} for this region} %
Five uncharacterized genes are expressed to a high degree in the
\elephant placenta (Supplemental \cref{tab:all_loxodonta_africana_placenta}), three of
which
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000028493",complete=TRUE)},
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000032011",complete=TRUE)},
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000031933",complete=TRUE)})
are members of an \elephant{}-specific expansion of 11 genes closely
related to the thymus-specific serine protease gene \textit{PRSS16}
(\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~C). The other 8
\textit{PRSS16}-like genes are also expressed to a lesser degree
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000026505",complete=TRUE)}).
Two other novel genes with high expression are related to the gene
\textit{SSLP1}
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000032659",complete=TRUE)}
and
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000029551",complete=TRUE)},
\cref{fig:sslp1_evolution_tree}), which is also expressed in \cow and
\dog placentas (Supplemental \cref{tab:all_canis_familiaris_placenta,tab:all_bos_taurus_placenta}).
Placental expression of the hormones relaxin (\textit{RLN}, %
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000017411")})
and prolactin (\textit{PRL}) are exceptionally high in the elephant
when compared to other mammals
(\cref{fig:rln2_evolution_tree,fig:prolactin_evolution_tree}). There
is little to no placental expression of the estrogen synthase gene,
\textit{CYP19A1}
(\Sexpr{gene.expression("loxodonta_africana","CYP19A1")}) and the
enzyme that produces progesterone, \textit{HSD3B1}
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000027896")}). Other genes
with high expression in the elephant placenta include genes involved
in oxidative stress
(\Sexpr{gene.expression("loxodonta_africana","TXN",complete=TRUE)}),
growth
(\Sexpr{gene.expression("loxodonta_africana","IGF2",complete=TRUE)})
and
\Sexpr{gene.expression("loxodonta_africana","PAEP",complete=TRUE)}, a
gene predominantly expressed in the endometrium of the uterus in humans
\cite{Rutanen.ea1987:Progesterone-associatedproteinsPP12and}.

\subsubsection{\armadillo Placenta Transcriptome}

% \begin{table}
% \footnotesize
% <<top.50.armadillo.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="dasypus novemcinctus"))
% @ 
% \caption{\tableexplanation{\armadillo}{tab:all_dasypus_novemcinctus_placenta}}
% \label{tab:top50_armadillo_placenta}
% \end{table}

The top 50 expressed genes in the term placenta of \armadillo are
shown in Supplemental \cref{tab:all_dasypus_novemcinctus_placenta} with a list of all genes
with $\text{FPKM}\ge 50$ in Supplemental
\cref{tab:all_dasypus_novemcinctus_placenta}. The gene with the
highest expression is
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000045029",complete=TRUE)}
which is most similar to the \mouse \textit{Pga5} is expanded in
\opossum into three members which are expressed in placenta (the other
two being
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000033885",complete=TRUE)}
and
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000040431",complete=TRUE)}),
though the expansion is less than that seen \cow and \sheep
(\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~D). There is also high expression of
hemoglobin subunit beta,
\Sexpr{gene.expression("dasypus_novemcinctus","HBB",complete=TRUE)},
an un-annotated gene with high similarity to \human and \cow
\textit{OVOS2}
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000033320",complete=TRUE)}
(Supplemental \cref{fig:ovos2_evolution_tree}), the growth factors
\Sexpr{gene.expression("dasypus_novemcinctus","IGF2",complete=TRUE)}
and
\Sexpr{gene.expression("dasypus_novemcinctus","IGFBP4",complete=TRUE)},
and the thymosin
\Sexpr{gene.expression("dasypus_novemcinctus","TMSB4X",complete=TRUE)}.
As has been seen previously, multiple components of the annexin
complex are also highly expressed, including
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000045634",complete=TRUE)}
(\textit{ANXA8L1}),
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000017399",complete=TRUE)}
(\textit{S100B}),
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000048504",complete=TRUE)}
(\textit{S100A10}), and
\Sexpr{gene.expression("dasypus_novemcinctus","ENSDNOG00000004237",complete=TRUE)}
(\textit{ANXA2}).

\subsection{Metatherians}

\subsubsection{\opossum Placenta Transcriptome}

% \begin{table}
%   \footnotesize
% <<top.50.opossum.expressed.genes,echo=FALSE,results="asis">>=
% output.top.genes.table(top.genes.table(species.name="monodelphis domestica"))
% @ 
% \caption{\tableexplanation{\opossum}{tab:all_monodelphis_domestica_placenta}}
% \label{tab:top50_opossum_placenta}
% \end{table}

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the \opossum are shown in
Supplemental \cref{tab:all_opossum_placenta} with a list of all genes in
Supplemental \cref{tab:all_monodelphis_domestica_placenta}. The gene with the
second highest expression is superoxide dismutase 3, extracellular
(\Sexpr{gene.expression("monodelphis_domestica","SOD3",complete=TRUE)}),
which functions as an antioxidant. The gene with the second highest
expression is the beta hemoglobin subunit epsilon
(\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000011325",complete=TRUE)}),
which is expressed at the embryonic stage of development in
marsupials, as in other mammals
\cite{Holland.ea1997:Characterizationofembryonicglobin}. \textit{HBE}
subunits link up with protein products of two hemoglobin, zeta
(\textit{HBZ}) genes, which are also expressed to a high degree in the
opossum placenta
(\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000016635",complete=TRUE)}
and
\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000016631",complete=TRUE)}).
In addition, there is high expression of the gene hemoglobin, omega
(\textit{HBO},
\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000016637",complete=TRUE)}),
% this is LOC_1000016060 which I figured out by tblastn'ing AAK11486.1
% against Monodelphis domestica
a hemoglobin found only in marsupials and monotremes
\cite{Opazo.ea2008:Genomicevidenceforindependent} (Supplemental
\cref{fig:hbe1_evolution_tree}). The hemoglobin gene \textit{HBA} is
also expressed in the placenta
(\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000016633",complete=TRUE)}).
Two cystatin genes
(\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000018295",complete=TRUE)}
and
\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000018296",complete=TRUE)})
are expressed to a high degree, along with one \textit{CTSL} related
gene
(\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000002844",complete=TRUE)},
\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~B), which are known to interact with
cystatins \cite{Turk.ea2001:Lysosomalcysteineproteasesfacts}. The
uncharacterized gene \textit{LOC10001195}
(\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000006547",complete=TRUE)})
also shows high expression in \opossum placenta. The gene
\textit{WFDC2}, a member of the whey acidic protein (WAP) family is
also expressed to a high degree in \opossum placenta
(\Sexpr{gene.expression("monodelphis_domestica","WFDC2",complete=TRUE)}).
Other genes with high expression include members of the annexin
complex
(\Sexpr{gene.expression("monodelphis_domestica","S100A6",complete=TRUE)},
\Sexpr{gene.expression("monodelphis_domestica","ANXA2",complete=TRUE)}).
In addition, there was little production of CYP19A1 and HSD3B1
(\Sexpr{gene.expression("monodelphis_domestica","CYP19A1",complete=TRUE)}
and
\Sexpr{gene.expression("monodelphis_domestica","ENSMODG00000005385",complete=TRUE)},
respectively).

\subsection{Core Placenta Transcriptome}

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species),]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]
print(ggplot(cptl,
             aes(y=human_name,x=Species))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(),
              axis.text.y = element_text(face="italic"),
              axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  1:1 \human ortholog genes expressed (\ac{FPKM}$ \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied with housekeeping genes
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} removed
  ordered (from top to bottom) by median expression. Supplemental
  \cref{fig:core_placenta_transcriptome_shape} shows the shape of the
  distribution of the number of genes in the core placenta
  transcriptome if different FPKM and species restriction criteria are
  utilized.}
\label{fig:core_placenta_transcriptome}
\end{figure}

The
\Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a 1:1 ortholog of that gene was
expressed with $\text{FPKM} \ge 10$ in all species studied and was
not a human housekeeping gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}.
Supplemental \cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently.
%
Multiple components of annexin complexes are included in the core
transcriptome, including \textit{ANXA2}, \textit{ANXA1},
\textit{S100A11}, \textit{S100A10}. ANXA1 is involved in the
resolution of the inflammation response through inhibition of
phospholipase A2 activity and signaling through the formyl peptide
receptor family, and has previously been hypothesized to be important
for the maintenance of the anti-inflammatory state during pregnancy
\cite{Hutchinson.ea2011:Molecularregulatorsofresolution}. ANXA2 is
known to interact with S100A10, S100A6, S100A11, and S100A4
\cite{Liu.ea2015:AnnexinA2complexeswith}, and there is evidence for
its interaction with S100P as well; it is involved in cell-cell
interaction as well as vesicle trafficking and von Willibrand Factor
secretion. The expression regulation of another core gene,
\textit{TIMP3}, has been implicated in pre-eclampsia
\cite{Ma.ea2014:Down-regulationofTIMP3leads,Xiang.ea2013:PromoterhypomethylationofTIMP3}.
Supplemental \cref{tab:core_placenta_go_terms} lists the significant
enrichment of \ac{GO} terms in the core placenta transcriptome;
notable terms include those involved in cytoskeletal structure,
protein transport, membrane protein structure, and inflammation
resolution. Increasing the permissiveness of the core placenta
transcriptome by only requiring 10 of 14 species to express 1:1
orthologs with FPKM $>10$ and results in the same terms also being
significantly enriched, along with additional terms (Supplemental
\cref{tab:core_placenta_go_terms}).

In contrast to the shape of the number of genes included in the core
placenta transcriptome, the set of \human housekeeping genes expressed
all placentas is flatter (Supplemental
\cref{fig:core_housekeeping_transcriptome_shape}), and includes more
of the total fraction of genes. However, there is still a large
difference between the set of housekeeping genes which are expressed
in \human and those which are expressed in all placentas.

% SPINT2 -- Serine Peptidase inhibitor

% SLC3A2 -- this is probably actually a housekeeping gene; it's
% involved in amino acid transport

% SQSTM1 -- NFκB signaling (immune supression?)

% KRTCAP2  N-glycosylation (probably housekeeping)

% NRDG1 -- interacts with HSP90

% SDC4 -- proteoglycan (syndecan) cell adhesion

% LAMC1 -- laminin

% LUM -- proteoglycan

% VDAC1 -- mitochondrial gene?

% F11R -- tight junctions

% SERPINH1? with TIMP3 
% http://www.proteinatlas.org/ENSG00000149257-SERPINH1/tissue/placenta

\subsubsection{\ac{GSE} Analyses of Individual Transcriptomes}

In order to identify major functional categories of genes expressed in
the individual transcriptomes, we examined \ac{GSE} in 1:1 human
orthologs which had \ac{FPKM} values in the top 1\% of expressed genes
(Supplemental %
\cref{tab:go_terms_enriched_top_one_dasypus_novemcinctus,tab:go_terms_enriched_top_one_equus_caballus,tab:go_terms_enriched_top_one_homo_sapiens,tab:go_terms_enriched_top_one_loxodonta_africana,tab:go_terms_enriched_top_one_monodelphis_domestica,tab:go_terms_enriched_top_one_mus_musculus,tab:go_terms_enriched_top_one_ovis_aries,tab:go_terms_enriched_top_one_pan_paniscus,tab:go_terms_enriched_top_one_sus_scrofa,tab:go_terms_enriched_top_one_ateles_fusciceps,tab:go_terms_enriched_top_one_spalax_galili,tab:go_terms_enriched_top_one_spalax_carmeli,tab:go_terms_enriched_top_one_bos_taurus,tab:go_terms_enriched_top_one_canis_familiaris}) using topGO \cite{topGO:2.20.0}. %
%
Significant enrichment in genes ($\text{FDR}\le0.05$) in the top 1\%
of \ac{FPKM} values in at least 12 of the
\Sexpr{length(species.names)} species included clusters such as
membrane organization , locomotion, tissue development, extracellular
matrix organization, regulation of peptidase activity, and various
cell death related pathways among others
(Supplemental \cref{tab:go_terms_all_species}). If we instead examine the top 100
genes by \ac{FPKM} in each of the \Sexpr{length(species.names)}
species (Supplemental %
\cref{tab:go_terms_enriched_dasypus_novemcinctus,tab:go_terms_enriched_equus_caballus,tab:go_terms_enriched_homo_sapiens,tab:go_terms_enriched_loxodonta_africana,tab:go_terms_enriched_monodelphis_domestica,tab:go_terms_enriched_mus_musculus,tab:go_terms_enriched_ovis_aries,tab:go_terms_enriched_pan_paniscus,tab:go_terms_enriched_sus_scrofa,tab:go_terms_enriched_ateles_fusciceps,tab:go_terms_enriched_spalax_galili,tab:go_terms_enriched_spalax_carmeli,tab:go_terms_enriched_bos_taurus,tab:go_terms_enriched_canis_familiaris})
notable enrichments include extracellular matrix, female pregnancy,
integrin and collagen-related processes, and various cell death
related processes.

\clearpage

\section{Discussion}

\subsection{Evolution of Placental Hormones}

Reproductive hormones are central to the regulation of pregnancy and
their source, timing, and quantity can vary widely from species to
species \cite{Carter2012:Evolutionofplacentalfunction}. Expression and synthesis of hormones is
perhaps the most well-documented physiological process in the placenta
in many species. Here we quantified the extreme disparity in genes
with hormone products or enzymatic products that directly produce
hormones in multiple mammalian species (\cref{fig:hormone_genes}).

\begin{figure}[htb]
<<hormone.genes.plot,echo=FALSE,fig.width=8,fig.height=8,warning=FALSE,message=FALSE>>=

gene.expression.nozero <- function(genes) {
    temp <-
        data.matrix.dt(dcast(combined.fpkm[human_name %in% genes,],
                             human_name~species,
                             value.var="FPKM",
                             fun.aggregate=function(x){x[1]}))
    temp <-
        temp[apply(temp,1,function(x){sum(x,na.rm=TRUE)})>5,]
    return(temp)
}

### hormone genes
hormone.genes <- c("CGA","CGB","CGB1","CGB2","CGB5","CGB8","CSH1","CSH2","CSHL1","CYP19A1",
                   "GH2","GHRH","HSD3B1","HSD3B2","HSD17B2","HSD17B1","NPPB","NPPC","PRL","RLN1")
hormone.expression.nozero <-
    gene.expression.nozero(hormone.genes)

hor.exp.long <- melt(log10(hormone.expression.nozero+1)[rev(1:nrow(hormone.expression.nozero)),])
colnames(hor.exp.long)[1:3] <- c("gene","species","fpkm")
hor.exp.long <- hor.exp.long[hor.exp.long$gene!="IGF2",]

### galectin genes
galectin.genes <- unique(grep("^LGALS",combined.fpkm[,human_name],value=TRUE))

galectin.expression.nozero <-
    gene.expression.nozero(galectin.genes)

gal.exp.long <- melt(log10(galectin.expression.nozero+1)[rev(1:nrow(galectin.expression.nozero)),])
colnames(gal.exp.long)[1:3] <- c("gene","species","fpkm")

### IGF related genes

igfrelated.genes <- unique(grep("^IGF",combined.fpkm[,human_name],value=TRUE))

igfrelated.expression.nozero <-
    gene.expression.nozero(igfrelated.genes)

igf.exp.long <- melt(log10(igfrelated.expression.nozero+1)[rev(1:nrow(igfrelated.expression.nozero)),])
colnames(igf.exp.long)[1:3] <- c("gene","species","fpkm")

### Pre-eclampsia genes ADAM12 CRH KISS1 PAPPA

pe.expression.nozero <-
    gene.expression.nozero(c("ADAM12","CRH","KISS1","PAPPA"))

pe.exp.long <- melt(log10(pe.expression.nozero+1)[rev(1:nrow(pe.expression.nozero)),])
colnames(pe.exp.long)[1:3] <- c("gene","species","fpkm")

igf.exp.long$type <- "IGF Related"
gal.exp.long$type <- "Galectin"
hor.exp.long$type <- "Hormone"
pe.exp.long$type <- "PE"

combined.long <-
    rbind(igf.exp.long,
          gal.exp.long,
          hor.exp.long,
          pe.exp.long
          )
combined.long$species <-
    factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(combined.long$species)),
           levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(names(species.names))))

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\caption{Expression in $\text{log}_{10}(\text{FPKM})$ of 1:1 \human
  orthologs of Hormone Genes, Galectins, genes implicated in
  pre-eclampsia (PE), and \textit{IGF}-related genes. Empty locations
  indicate no detectable expression and/or no 1:1 \human ortholog in
  that species.}
\label{fig:hormone_genes}
\end{figure}

\subsubsection{Estrogen and Progesterone}

In both ruminants and primates, the placenta is capable of producing
estrogen and progesterone
\cite{Strauss.ea1996:Placentalsteroidhormonesynthesis}. High relative
expression of \textit{CYP19A1} (Supplemental
\cref{fig:estrogen_evolution_tree}), the enzyme that synthesizes
estrogen from androstenedione is seen in \human, \bonobo, \cow, and
\horse, though not in \sheep, \smonkey. %
%
\textit{HSD3B1}, one of two enzymes that synthesizes progesterone from
pregnalone
(\cite{Simard.ea2005:Molecularbiologyof3beta-hydroxysteroid},
Supplemental \cref{fig:progesterone_evolution_tree}) is also highly
expressed in \human, \bonobo, \cow, \horse, and \dog, but is not
highly expressed in \smonkey, \sheep, \pig, \elephant, \armadillo, or
\opossum. This is consistent with the previously observed local
progesterone withdrawal in \sheep, and the leuteo-placental shift of
progesterone expression in boreoeutheria in contrast to afrotherians
which do not show such a shift
\cite{Swaggart.ea2015:Genomicsofpretermbirth}. The diversity of
\textit{CYP19A1} and \textit{HSD3B1} expression levels in in primates
and ungulates seen here points to the need for more complete sampling
of placentas at different timepoints and in different clades.

\subsubsection{Prolactin}

The elephant is strikingly the only mammal analyzed that shows high
expression of the primarily pituitary hormone prolactin (\textit{PRL})
in the term placenta, although we see high expression of genes which
resulted from duplications of \textit{PRL} in the \mouse, \carmeli,
and \cow (Supplemental \cref{fig:prolactin_evolution_tree}). The
results in \mouse are relatively consistent with the temporal
expression results of
\citet{Simmons.ea2008:Spatialandtemporalexpression}. For example, we
see very high expression of \textit{Prl3b1} and \textit{Prl8a9} in the
term placentas of \mouse, but low expression of \textit{Prl7a1} and
\textit{Prl8a2} which are expressed at earlier embryonic stages. %
All afrotherians yet analyzed contain only one prolactin gene, and
prolactin has undergone adaptive evolution on the \elephant lineage
\cite{Wallis2009:ProlactininAfrotheriacharacterization}.
\citet{Yamamoto.ea2011:PlacentationinAfricanelephant} noted the
presence of a prolactin in the trophoblast of the elephant placenta
using immunohistological staining and human derived antibodies. A
previous analysis of \elephant placenta also detected the expression
of prolactin \cite{Hou.ea2012:Elephanttranscriptomeprovidesinsights}.
Here we confirm that the sole prolactin in \elephant is not only
expressed but highly expressed in the placenta, and likely functions
both in the maintenance of the corpus luteum and as a placental
lactogen \cite{Yamamoto.ea2011:PlacentationinAfricanelephant}.
\textit{PRL} is not expressed in \opossum, even though there is
evidence that it is expressed in the placenta of another marsupial,
\textit{Macropus eugenii} (tammar wallaby)
\cite{Menzies.ea2011:Placentalexpressionofpituitary}. This suggests
that placental prolactin expression in marsupials may be as diverse as
in eutherian mammals. One possible explanation of the diversity of
expression is the presence of endogenous retroviruses such as MER39 in
primates, MER77 in rodents, and L1-2\_LA in \elephant
\cite{Emera.ea2012:Convergentevolutionofendometrial} although this
does not explain the diversity of expression of different isoforms of
\textit{PRL} in \mouse or their tissue specificity in primates. An
alternative, that \textit{HOXA11} expression is responsible for
\textit{PRL} expression
\cite{Lynch.ea2008:Adaptivechangesintranscription} seems not to be the
case in placenta, as the rodent with the highest \textit{HOXA11}
expression, \galili
(\Sexpr{gene.expression("spalax_galili","Hoxa11",complete=TRUE)}) is
the rodent with the lowest expression of \textit{PRL}
(\cref{fig:prolactin_evolution_tree}), and \elephant also does not
express \textit{HOXA11} highly either
(\Sexpr{gene.expression("loxodonta_africana","HOXA11",complete=TRUE)},
Supplemental \cref{fig:hoxa11_evolution_tree}).

\subsubsection{Growth hormone family}

The growth hormone family of genes has expanded in anthropoid
primates, resulting in approximately 5 genes in \human (\textit{GH1},
\textit{GH2}, \textit{CSH1}, \textit{CSH2}, \textit{CSHL1}) and at
least three in \bonobo (two \textit{GH2} and \textit{GH1}; owing to
the similarity between the two \textit{GH1} in \textit{P. troglodytes}
we cannot reliably distinguish transcripts between the two). All show
substantial expression in the \human placenta
(\cref{fig:csh1_evolution_tree}). The expression of the \ac{GH} family
in placenta appears to be primate specific, and likely existed at the
time of the last common ancestor between all extant anthropoids
\cite{Papper.ea2009:Ancientoriginofplacental,Soares2004:prolactinandgrowthhormone}.
There is some evidence that increased expression of these genes may be
correlated with overall birth weight of offspring in \human
\cite{Männik.ea2010:Differentialexpressionprofileof}, and it is
possible that the extreme expression levels of these genes may be due
to high birth weight. \human placental \ac{GH} genes are involved in
increasing blood flow, replacing maternal pituitary expression of
\ac{GH}, as well as trophoblast invasion
\cite{Alsat.ea1998:Physiologicalroleofhuman,Lacroix.ea2005:Stimulationofhumantrophoblast,Schiessl.ea2007:Roleofplacentalgrowth},
and it is likely that they have the same function in \bonobo.

\begin{figure}[htb]
<<chorionic.somatomammotropin.alpha.evolution.tree,echo=FALSE,fig.width=4,fig.height=4,warning=FALSE,message=FALSE>>=

pushViewport(viewport(layout=grid.layout(nrow=1,ncol=2,widths=c(2,3))))

print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="CGA" &
                                                    species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                combined.fpkm=combined.fpkm,
                                trees=trees
                                )+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,0.4)+
      theme(legend.position="none"),
      vp=viewport(layout.pos.col=1,layout.pos.row=1,clip="off"))
print(plot.ancestral.state.tree(gene_selector=quote(gene_short_name=="CGB1" &
                                                    species=="homo sapiens"),
                                genes.to.tree=genes.to.tree,
                                subtree=c("ENSMODG00000013768","ENSG00000104818"),
                                combined.fpkm=combined.fpkm,
                                trees=trees
                                )+
      geom_tiplab(size=2,color="black",fontface="italic")+xlim(0,0.5),
      vp=viewport(layout.pos.col=2,layout.pos.row=1,clip="off"))
popViewport(1)
grid.text(c("A","B"),
          x=unit(c(0,2/5),"npc"),y=unit(c(1,1),"npc"),
          just=c("left","top"),
          gp=gpar(fontsize="24")
          )
@ 
\caption{\genetreeexplanation{(A) \textit{CGA} (B) \textit{CGB}}}
\label{fig:cga_cgb_evolution_tree}
\end{figure}

The human gene \textit{CGA} codes for the alpha subunit of multiple
hormones including pituitary hormones (luteinizing hormone,
follicle-stimulating hormone, thyroid-stimulating hormone), as well as
numerous chorionic gonadotropins with variable beta subunits
\cite{Maston.ea2002:Chorionicgonadotropinhasrecent}. The function of
the chorionic gonadotropins is to prevent the regression of the corpus
luteum \cite{Hanson.ea1971:EffectsofHCGand}. Among the six placentas
we analyzed here, \textit{CGA} is only expressed in the placentas of
\human and \bonobo (\cref{fig:hormone_genes,fig:cga_cgb_evolution_tree}),
which corresponds with the emergence of the chorionic gonadotropins in
anthropoid primates
\cite{Maston.ea2002:Chorionicgonadotropinhasrecent}. However, oddly,
\textit{CGA} is expressed at much higher levels in both species than
the beta subunits with which its product dimerizes (\cref{fig:hormone_genes,fig:cga_cgb_evolution_tree}).
\textit{CGA} is one of several genes noted by
\citet{Uddin.ea2008:Distinctgenomicsignaturesof} to have undergone
adaptive evolution of the \human{}-\textit{P. troglodytes} stem lineage.

\subsubsection{Relaxin}

Relaxin (\textit{RLN}) is a hormone that has multiple effects during
pregnancy including increasing maternal blood flow, widening the pubic
bone and relaxing uterine musculature in preparation for labor. In
\dog, \elephant, \mouse, \carmeli, and \horse there is high expression
of relaxin in the term placenta, with little evidence in other species
examined here (Supplemental \cref{fig:rln2_evolution_tree}). \textit{RLN} is known
to be produced in the placenta of the pregnant \dog and is used in the
diagnosis of \dog pregnancy
\cite{Steinetz.ea1989:Serumrelaxinandprogesterone,Tsutsui.ea1991:Determinationofsourceof},
the luteal ovary and placenta of \pig
\cite{Hisaw.ea1948:Relaxininovaryof}, and the placenta of \horse
\cite{Klonisch.ea1997:Placentallocalizationofrelaxin}. However, the
source of relaxin produced by pregnant \elephant remains unclear
\cite{Meyer.ea2004:Speciesandfetalgender}. The high rise in \textit{RLN} in
\elephant has been suggested to be a major signal for the initiation
of labor \cite{Niemuller.ea1998:Plasmaconcentrationsofimmunoreactive}.
Here we show that the placenta of elephants show high expression of
\textit{RLN} and is likely a major source of \textit{RLN} at term. In
contrast to the results of
\cite{Evans.ea1993:mouserelaxingenenucleotide}, we also see expression
of \textit{RLN} in the placenta of \mouse and \carmeli.

\subsubsection{\textit{NPPB} and \textit{NPPC}}

Two genes, \textit{NPPB} and \textit{NPPC}, code for peptide hormones
that are important in the inducement of vasodilation, and may be
important in increasing maternal blood flow to the fetus
\cite{Cameron.ea1996:sitesofgeneexpression}. There is some indication
that these proteins are present in the decidua of mice
\cite{Cameron.ea1996:sitesofgeneexpression}, but we find very little
expression in the fetal portion of the mouse placenta
(\cref{fig:hormone_genes}, Supplemental
\cref{fig:nppc_evolution_tree,fig:nppb_evolution_tree}). While we do
find high expression of at least one of these genes in \dog, \cow, and
\elephant, the expression pattern does not follow the placental
organization (endotheliochorial for \dog and \elephant,
synepitheliochorial for \cow and \sheep), so it is unlikely that
expression of these genes are essential for specific placental
organizations.

\subsection{The Evolution of Placental Expression of Galectins}

Galectins are recognized as major regulators of pregnancy in both
fetal and maternal tissues
\cite{Than.ea2012:Galectinsguardiansofeutherian}. Indeed, the
evolution of new lineage-specific galectins has occurred in both
primates and ruminants, with the genes \textit{LGALS13},
\textit{LGALS14}, and \textit{LGALS16} recognized as being
placenta-specific in \human and regulating immune responses by
inducing apoptosis of maternal T-cells
\cite{Than.ea2009:primatesubfamilyofgalectins} (Supplemental
\cref{fig:galectins_lgals13_evolution_tree}). Indeed, \textit{LGALS14}
has the highest expression of any galectin in the \human and \smonkey
placenta (\cref{fig:hormone_genes}, Supplemental
\cref{fig:galectins_lgals13_evolution_tree}), and other primate
galectins such as \textit{LGALS13} and \textit{LGALS16} are also
expressed in \human and \smonkey placenta. However, in the closely
related \bonobo, \textit{LGALS14} is not found, and \textit{LGALS1} is
the most abundant galectin. \textit{LGALS1} shows generally high
expression in all placentas analyzed. \textit{LGALS1} is also involved
in mediating maternal-fetal immune tolerance, and is expressed in both
maternal and fetal tissue in primates
\cite{Than.ea2008:Emergenceofhormonaland}. \textit{LGALS1} placental
expression was thought to be confined to eutherian mammals, as there
is evidence of adaptive evolution at this node, as well as the
appearance of estrogen response elements in eutherian cis-regulatory
regions \cite{Than.ea2008:Emergenceofhormonaland}. It was unclear as
to whether galectins such as \textit{LGALS1} were important in the
regulation of pregnancy in marsupials, as well as eutherians. Here we
show substantial expression of galectins in the opossum, especially
\textit{LGALS1} (Supplemental \cref{fig:galectins_evolution_tree}). In
addition, \textit{LGALS2} is also expressed. The presence of galectins
at the maternal-fetal interface in \opossum implies that it was also
present in the ancestral placenta of therian mammals. Indeed, adaptive
evolution of \textit{LGALS1} on the eutherian stem is likely due to
the increased invasive nature of eutehrian placentation rather than
the origination of maternal-fetal tolerance. Although the function of
\textit{LGALS1} in the marsupial placenta needs to be confirmed, it is
likely that the mediation of maternal-fetal immune tolerance by
galectins is an ancient feature of mammalian viviparity.

\subsection{The Opossum Placenta and the Evolution of Viviparity}

The placenta of \opossum is distinctly different from that of the term
placenta of eutherian mammals in that it is derived from both the
chorion and yolk sac. Other differences between the placentas of
marsupials and eutherians include lack of invasive extravillous
trophoblasts in marsupials, the retention of a permeable shell coat
for 2/3 of marsupial gestation, and the greater degree of imprinting
in placenta-expressed genes in eutherians
\cite{Renfree2010:ReviewMarsupialsplacentalmammals,Renfree.ea2013:originandevolutionof}.
In addition, the placenta of \opossum is also distinct from other
marsupials, in that it is considered one of the least specialized, and
thought be closer to the original marsupial condition
\cite{Renfree2010:ReviewMarsupialsplacentalmammals}. In \opossum, the
duration of direct connection between maternal and fetal tissues is as
short as two days in a gestational period of 13-15 days
\cite{Harder.ea1993:Gestationandplacentationin}. During this period,
the placenta is directly engaged in the exchange of urea with maternal
tissues, owing to the lack of a vesicular allantois
\cite{Freyer.ea2007:Placentalfunctionintwo}. This differs drastically
with the situation in another well-studied marsupial, the tammar
wallaby \cite{Freyer.ea2007:Placentalfunctionintwo}. Therefore, we
must interpret differences between the opossum and eutherian placentas
with caution, as highly expressed genes and functional annotation
clusters may represent didelphid or \opossum specific adaptations.

The \opossum placenta shows expression of genes that either do not
exist in eutherians or are not expressed in the term eutherian placenta
(Supplemental \cref{tab:all_opossum_placenta}). The existence of embryonic forms
of hemoglobin in fetal tissues at term, as evidenced by the expression
of \textit{HBE} and \textit{HBZ} is a feature of marsupials that is
generally due to their very short gestation periods. Embryonic
hemoglobins in marsupials have a lower affinity for oxygen and allow
for a lower rate of oxygen uptake from maternal tissues
\cite{Holland.ea1997:Characterizationofembryonicglobin}. One novel
uncharacterized gene that shows high expression (\textit{LOC10001195}) is not
found in eutherians; Ensembl gene trees reveal that homologs of this
gene exist in other amniotes such as birds and turtles, but are not
present in eutherian mammals. The homolog of this gene in chicken is
found to expressed in reproductive tract
\cite{Froman.ea2006:expressedsequencetaganalysis}, but its function is
unknown. \textit{WFDC2} is a member of the whey acidic protein (WAP)
family and involved in innate immunity. It is expressed in the mammary
gland of the tammar wallaby during pregnancy and early lactation
\cite{Sharp.ea2007:Molecularevolutionofmonotreme}, and present in
relative abundance in \opossum placenta.

\subsection{Novel Gene Expansions in the Elephant Placenta}

We found evidence of high expression in two novel expansions of genes
in the \elephant genome. One family includes 11 genes that are related
to the \textit{PRSS16} gene, a thymus-specific serine protease that
may be involved regulating the presentation of self-peptides in CD4+
T-lymphocytes \cite{Fornari.ea2012:TranscriptionprofilingofPrss16}.
All 11 of these genes are expressed in the elephant placenta, three of
which have very abundant mRNAs
(Supplemental \cref{tab:all_loxodonta_africana_placenta} and \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}C). It is
possible that this elephant-specific gene family may be involved in
immunity in the elephant placenta and the mediation of maternal-fetal
immune tolerance; however, future work will need to test these
hypotheses. Three other genes abundantly expressed in the elephant
placenta are homologues of \textit{SSLP1}, a gene with predominant
expression in the seminal vesicle in mouse
\cite{Li.ea2006:SSLP-1secretedLy-6protein}. Homologues of
\textit{SSLP1} are also expressed in \cow, \dog, \horse, and \sheep.
\textit{SSLP1} also contains an SINE repeat and may be
involved in the development of villi via cell proliferation
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}. This
gene is not found in primates. Expansion of these genes with high
expression in \elephant may indicate an elaboration of this system.

% Bovine has their own alu-like sequence: PMID 7835712

\subsection{Core placenta transcriptome}

The transcriptomes from multiple different species enabled us to
identify the set of expressed genes which are common to all placentas.
Components of the annexin complexes, including
\textit{ANXA2}, \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10}
are expressed in all placentas examined, and argue for the evolution
importance of annexins in placenta function, where they likely
establish maternal-fetal
tolerance\cite{Liu.ea2015:AnnexinA2complexeswith,Hutchinson.ea2011:Molecularregulatorsofresolution,Aarli.ea1997:Suppressiveeffectonlymphoproliferation}.
We suggest that further research into molecular role of these annexin
complexes may illuminate how immune tolerance is maintained during
placentation. The core placenta transcriptome is also enriched in
genes involved in maintaining the structure of the placenta and
transporting proteins, as would be expected given the function of the
placenta.

\subsection{\human Specific Expression Shifts and the Evolution of Pre-eclampsia}

The \human placenta is distinctive from many other primates in that it
possesses a deeply invasive trophoblast that penetrates the inner
myometrium of the uterus
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}. It is
hypothesized that this feature is related to the increased demand for
resources by the developing fetus to feed the expanded size of the
human brain
\cite{Carter.ea2011:Evolutionofinvasiveplacentation,Robillard.ea2002:Evolutionaryadaptationstopre-eclampsiaeclampsia}.
\human have the largest brain in terms of relative size and this
differs markedly from closely related apes
\cite{Boddy.ea2012:Comparativeanalysisofencephalization}. Deep
penetration of the myometrium occurs in chimpanzees and gorillas, but
this penetration may be accelerated in humans
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}.

The near exclusivity of deep trophoblast invasion in humans has also
led many to speculate on the origin of various disorders, which are
almost exclusively symptomatic of human pregnancy. For example,
pre-eclampsia is a human disease that occurs in 4\% of pregnancies in
the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Pre-eclampsia
can be attributed to the lack of adequate invasion of maternal tissue,
and is noted by the onset of hypertension and proteinuria (or in the
absence of proteinuria, thrombocytopenia, impaired liver function,
renal insufficiency, pulmonary edema, or cerebral or visual
disturbances) %
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current,ACOG2013:hypertensionpregnancy}. Some cases of
pre-eclampsia can develop into eclampsia, a more severe condition
resulting in seizures during pregnancy and death. Lack of invasion in
pre-eclampsia may be attributed to the inefficiency of growth
regulators as well as the inability to sufficiently suppress the
maternal immune response. Although there are speculative cases of
eclampsia in great apes
\cite{Baird1981:Eclampsiainlowlandgorilla,Stout.ea1969:Glomerularcapillaryendothelialswelling,Thornton.ea1992:Convulsionsinpregnancyin},
it is exceedingly rare in these species
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}.

The identification of multiple genes (\textit{CRH}, \textit{ADAM12},
\textit{KISS1}, \textit{PAPPA}, \textit{IL1RL1}, and \textit{PLAC1};
\cref{tab:human_lineage_specific}) which are associated with
pre-eclampsia which have increased expression on the \human lineage
may explain the larger proportion of pre-eclampsia observed in humans.
However, there are still genes which are associated with pre-eclampsia
which are also more highly expressed in the \mouse lineage such as
\textit{Afp} and \textit{Gjb2}. Furthermore, the onset of
pre-eclampsia is well before term; all of our placentas are term
placentas and provide limited information about the transcriptome
during the onset of pre-eclampsia.

\subsection{Conclusion}

We analyzed the sequenced transcriptome of
\Sexpr{star.logs[,length(unique(file))]} placentas from
\Sexpr{star.logs[,length(unique(species))]} species spanning the
phylogenetic tree of mammals. This wealth of data revealed both
lineage-specific expansions and differential expression of gene
families as well as a set of core, non-housekeeping genes which are
critical for the proper organization and functioning of the placenta.

There is diversity in the placental expression of hormones, including
estrogen, progesterone, prolactin, relaxin, the growth hormone family, and
natriuretic peptides among the species and lineages analyzed.
We also identified genes differentially expressed on the \human
lineage which are implicated in pre-eclampsia, including
\textit{ADAM12}, \textit{CRH}, and \textit{KISS1}.

The core placenta transcriptome was also identified, which contains
genes which are involved in maternal-fetal tolerance, such as the
annexins, and genes responsible for cell-cell and cell-matrix
interactions, such as \textit{LGALS1}. These genes were almost
certainly present in the placenta of the common mammalian ancestor. We
hypothesize that modifications to their expression or function will
compromise the development and organization of the placenta.

This study has identified many tantalizing molecular pathways which
illuminate the vast diversity of placental morphology and may explain
some placental malfunctions in \human{}. However, the strength of our
hypotheses are limited by our limited data; more placentas from
multiple members of diverse clades and different timepoints are still
needed to come to a more complete understanding of the genomic and
molecular basis for placenta morphology and function. We look forward
to collecting and analyzing such a dataset.

%\begin{backmatter}
  \section*{Competing interests}
  The authors declare that they have no competing interests.

  \section*{Author's contributions}
  Don Armstrong analyzed all of the data and wrote the paper.

  Michael McGowan wrote the paper and performed initial analyses.

  Amy Weckle performed all of the RNA extractions and managed the
  samples.

  Jason Caravas performed initial analyses.

  Dalen Agnew provided samples.

  Kurt Benirschke provided samples.

  Sue Savage-Rumbaugh provided samples.

  Eviatar Nevo provided samples.

  Chong J. Kim provided samples.

  Günter P. Wagner provided samples and wrote the paper.

  Roberto Romero provided samples and conceived the study.

  Derek E. Wildman conceived the study, wrote the paper, and performed
  analyses.

  \section*{Acknowledgements}
  The authors would like to specially acknowledge those groups whose
  publicly available data contributed to the completeness of this
  work.

  \section*{Funding}
  Partial support for this research came from the \textit{Eunice
    Kennedy Shriver} National Institute of Child Health and Human
  Development (NICHD), National Institutes of Health, U.S. Department
  of Health and Human Services (N01-HD-3342), and the National Science
  Foundation BCS-0827546.

%  \printaddresses

%\end{backmatter}

\printbibliography

<<save_mentioned_genes,echo=FALSE,results="hide">>=
saved_mentioned_genes <- mentioned.genes
@

\end{document}