### 2016-11-29 manuscript/acronyms.tex
use long names for genes where possible

\DeclareAcronym{EV}{short=EV,long=Extracellular Vesicle}
\DeclareAcronym{SVM}{long=Support Vector Machine}
\DeclareAcronym{ROC}{long=Receiver Operating Characteristic}
\DeclareAcronym{KNN}{long=K Nearest Neighbors}
\DeclareAcronym{SRA}{long=Sequence Read Archive}
\DeclareAcronym{PRG2}{short=\textit{PRG2},long=Proteoglycan 2, bone marrow}
\DeclareAcronym{FPKM}{long=Fragments Per Kilobase of exon per Million reads}
\DeclareAcronym{CSH1}{short=\textit{CSH1},long=chorionic somatomammotropin hormone 1}
\DeclareAcronym{CSH2}{short=\textit{CSH2},long=chorionic somatomammotropin hormone 2}
\DeclareAcronym{CSHL1}{short=\textit{CSHL1},long=chorionic somatomammotropin hormone-like 1}
\DeclareAcronym{GH2}{short=\textit{GH2},long=growth hormone 2}
\DeclareAcronym{CEA}{long=carcinoembryonic antigen}
\DeclareAcronym{HSD3B1}{short=\textit{HSD3B1},long=3-β Hydroxysteroid dehydrogenase 1}
\DeclareAcronym{GSE}{long=Gene Set Enrichment}
\DeclareAcronym{OU}{long=Ornstein-Uhlenbeck}
\DeclareAcronym{OUCH}{long=\ac{OU} Comparative Hypotheses}
\DeclareAcronym{TSI}{long=Tissue Specificity Index}
\DeclareAcronym{SRA}{long=Sequence Read Archive}
\DeclareAcronym{GH}{long=Growth Hormone}
\DeclareAcronym{PSG}{long=Pregnancy-Specific Glycoprotein}
\DeclareAcronym{PAG}{long=Pregnancy-Associated Glycoprotein}
\DeclareAcronym{GO}{long=Gene Ontology}
\DeclareAcronym{FDR}{long=False Discovery Rate}
\DeclareAcronym{gsea}{short=GSEA,long=gene set enrichment analysis}
\DeclareAcronym{CRH}{short=\textit{CRH},long={corticotropin-releasing hormone}}
\DeclareAcronym{KISS1}{short=\textit{KISS1},long={KiSS-1 metastatis-supressor}}
\DeclareAcronym{PAPPA}{short=\textit{PAPPA},long={pappalysin 1}}
\DeclareAcronym{ADAM12}{short=\textit{ADAM12},long={ADAM Metallopeptidase Domain 12}}
\DeclareAcronym{ANXA1}{short=\textit{ANXA1},long={annexin A1}}
\DeclareAcronym{ANXA5}{short=\textit{ANXA5},long={annexin A5}}
\DeclareAcronym{S100A11}{short=\textit{S100A11},long={S100 calcium binding protein A11}}
\DeclareAcronym{S100A10}{short=\textit{S100A10},long={S100 calcium binding protein A10}}
\DeclareAcronym{LGALS1}{short=\textit{LGALS1},long={galectin 1}}
\DeclareAcronym{LAMC1}{short=\textit{LAMC1},long={laminin subunit gamma 1}}
\DeclareAcronym{LUM}{short=\textit{LUM},long={lumican}}
\DeclareAcronym{GRB2}{short=\textit{GRB2},long={growth factor receptor bound protein 2}}
\DeclareAcronym{RALB}{short=\textit{RALB},long={RAS like proto-oncogene B}}
\DeclareAcronym{ERVFRD-1}{short=\textit{ERVFRD-1},long={endogenous retrovirus group FRD member 1}}
\DeclareAcronym{ERVW-1}{short=\textit{ERVW-1},long={endogenous retrovirus group W member 1}}
\DeclareAcronym{MFSD2A}{short=\textit{MFSD2A},long={major facilitator superfamily domain containing 2A}}
\DeclareAcronym{PRSS16}{short=\textit{PRSS16},long={protease, serine 16}}
\DeclareAcronym{TIMP3}{short=\textit{TIMP3},long={TIMP metallopeptidase inhibitor 3}}
\DeclareAcronym{F11R}{short=\textit{F11R},long={F11 receptor}}
\DeclareAcronym{SDC4}{short=\textit{SDC4},long={syndecan 4}}
\DeclareAcronym{FN1}{short=\textit{FN1},long={fibronectin 1}}
\DeclareAcronym{ITGB1}{short=\textit{ITGB1},long={integrin subunit beta 1}}
\DeclareAcronym{ITGB1BP1}{short=\textit{ITGB1BP1},long={integrin subunit beta 1 binding protein 1}}
\DeclareAcronym{EREG}{short=\textit{EREG},long={epiregulin}}
\DeclareAcronym{UNC93A}{short=\textit{UNC93A},long={unc-93 homolog A (\textit{C. elegans})}}
\DeclareAcronym{UNC93B1}{short=\textit{UNC93B1},long={unc-93 homolog B1 (\textit{C. elegans)}}}
\DeclareAcronym{WNT5A}{short=\textit{WNT5A},long={Wnt family member 5A}}
\DeclareAcronym{IL1RL1}{short=\textit{IL1RL1},long={interleukin 1 receptor like 1}}
\DeclareAcronym{PLAC1}{short=\textit{PLAC1},long={placenta specific 1}}
\DeclareAcronym{Afp}{short=\textit{Afp},long={alpha fetoprotein}}
\DeclareAcronym{Gjb2}{short=\textit{Gjb2},long={gap junction protein beta 2}}
\DeclareAcronym{CD9}{short=\textit{CD9},long={CD9 molecule}}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manuscript.Rnw"
%%% End: