### 2016-12-17 manuscript/acronyms.tex
add links to hugo for genes

\DeclareAcronym{EV}{short=EV,long={Extracellular Vesicle}}
\DeclareAcronym{SVM}{short=SVM,long={Support Vector Machine}}
\DeclareAcronym{ROC}{short=ROC,long={Receiver Operating Characteristic}}
\DeclareAcronym{KNN}{short=KNN,long={K Nearest Neighbors}}
\DeclareAcronym{SRA}{short=SRA,long={Sequence Read Archive}}
\DeclareAcronym{CEA}{short=CEA,long={carcinoembryonic antigen}}
\DeclareAcronym{GSE}{short=GSE,long={Gene Set Enrichment}}
%\DeclareAcronym{OU}{short=OU,long=Ornstein-Uhlenbeck}
%\DeclareAcronym{OUCH}{short=OUCH,long=\ac*{OU} Comparative Hypotheses}
\DeclareAcronym{TSI}{short=TSI,long=Tissue Specificity Index}
\DeclareAcronym{GH}{short=GH,long=Growth Hormone}
\DeclareAcronym{PSG}{short=PSG,long={Pregnancy-Specific Glycoprotein}}
\DeclareAcronym{PAG}{short=PAG,long={Pregnancy-Associated Glycoprotein}}
\DeclareAcronym{GO}{short=GO,long={Gene Ontology}}
\DeclareAcronym{FDR}{short=FDR,long={False Discovery Rate}}
\DeclareAcronym{gsea}{short=GSEA,long={gene set enrichment analysis}}
\DeclareAcronym{FPKM}{short=FPKM,long={Fragments Per Kilobase of exon per Million reads}}
\DeclareAcronym{HSD3B1}{short=\textit{HSD3B1},long={hydroxy-δ-5-steroid dehydrogenase, 3 β- and steroid δ-isomerase 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=HSD3B1}{\textit{HSD3B1}} (hydroxy-δ-5-steroid dehydrogenase, 3 β- and steroid δ-isomerase 1)}}
\DeclareAcronym{CSH1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CSH1}{\textit{CSH1}},long={chorionic somatomammotropin hormone 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CSH1}{\textit{CSH1}} (chorionic somatomammotropin hormone 1)}}
\DeclareAcronym{CSH2}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CSH2}{\textit{CSH2}},long={chorionic somatomammotropin hormone 2},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CSH2}{\textit{CSH2}} (chorionic somatomammotropin hormone 2)}}
\DeclareAcronym{CSHL1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CSHL1}{\textit{CSHL1}},long={chorionic somatomammotropin hormone like 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CSHL1}{\textit{CSHL1}} (chorionic somatomammotropin hormone like 1)}}
\DeclareAcronym{GH2}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=GH2}{\textit{GH2}},long={growth hormone 2},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=GH2}{\textit{GH2}} (growth hormone 2)}}
\DeclareAcronym{PRG2}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PRG2}{\textit{PRG2}},long={proteoglycan 2, pro eosinophil major basic protein},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PRG2}{\textit{PRG2}} (proteoglycan 2, pro eosinophil major basic protein)}}
\DeclareAcronym{CRH}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CRH}{\textit{CRH}},long={corticotropin releasing hormone},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CRH}{\textit{CRH}} (corticotropin releasing hormone)}}
\DeclareAcronym{KISS1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=KISS1}{\textit{KISS1}},long={KiSS-1 metastasis-suppressor},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=KISS1}{\textit{KISS1}} (KiSS-1 metastasis-suppressor)}}
\DeclareAcronym{PAPPA}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PAPPA}{\textit{PAPPA}},long={pappalysin 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PAPPA}{\textit{PAPPA}} (pappalysin 1)}}
\DeclareAcronym{ADAM12}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ADAM12}{\textit{ADAM12}},long={ADAM metallopeptidase domain 12},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ADAM12}{\textit{ADAM12}} (ADAM metallopeptidase domain 12)}}
\DeclareAcronym{ANXA1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ANXA1}{\textit{ANXA1}},long={annexin A1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ANXA1}{\textit{ANXA1}} (annexin A1)}}
\DeclareAcronym{ANXA2}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ANXA2}{\textit{ANXA2}},long={annexin A2},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ANXA2}{\textit{ANXA2}} (annexin A2)}}
\DeclareAcronym{ANXA5}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ANXA5}{\textit{ANXA5}},long={annexin A5},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ANXA5}{\textit{ANXA5}} (annexin A5)}}
\DeclareAcronym{S100A11}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=S100A11}{\textit{S100A11}},long={S100 calcium binding protein A11 (calgizzarin)},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=S100A11}{\textit{S100A11}} (S100 calcium binding protein A11; calgizzarin)}}
\DeclareAcronym{S100A10}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=S100A10}{\textit{S100A10}},long={S100 calcium binding protein A10 (calpectin-I light polypeptide)},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=S100A10}{\textit{S100A10}} (S100 calcium binding protein A10; calpectin-I light polypeptide)}}
\DeclareAcronym{LGALS1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=LGALS1}{\textit{LGALS1}},long={galectin 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=LGALS1}{\textit{LGALS1}} (galectin 1)}}
\DeclareAcronym{LAMC1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=LAMC1}{\textit{LAMC1}},long={laminin subunit γ 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=LAMC1}{\textit{LAMC1}} (laminin subunit γ 1)}}
\DeclareAcronym{LUM}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=LUM}{\textit{LUM}},long={lumican},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=LUM}{\textit{LUM}} (lumican)}}
\DeclareAcronym{GRB2}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=GRB2}{\textit{GRB2}},long={growth factor receptor bound protein 2},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=GRB2}{\textit{GRB2}} (growth factor receptor bound protein 2)}}
\DeclareAcronym{RALB}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=RALB}{\textit{RALB}},long={RAS like proto-oncogene B},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=RALB}{\textit{RALB}} (RAS like proto-oncogene B)}}
\DeclareAcronym{ERVFRD-1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ERVFRD-1}{\textit{ERVFRD-1}},long={endogenous retrovirus group FRD member 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ERVFRD-1}{\textit{ERVFRD-1}} (endogenous retrovirus group FRD member 1)}}
\DeclareAcronym{ERVW-1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ERVW-1}{\textit{ERVW-1}},long={endogenous retrovirus group W member 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ERVW-1}{\textit{ERVW-1}} (endogenous retrovirus group W member 1)}}
\DeclareAcronym{MFSD2A}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=MFSD2A}{\textit{MFSD2A}},long={major facilitator superfamily domain containing 2A},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=MFSD2A}{\textit{MFSD2A}} (major facilitator superfamily domain containing 2A)}}
\DeclareAcronym{PRSS16}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PRSS16}{\textit{PRSS16}},long={protease, serine 16},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PRSS16}{\textit{PRSS16}} (protease, serine 16)}}
\DeclareAcronym{TIMP3}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=TIMP3}{\textit{TIMP3}},long={TIMP metallopeptidase inhibitor 3},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=TIMP3}{\textit{TIMP3}} (TIMP metallopeptidase inhibitor 3)}}
\DeclareAcronym{F11R}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=F11R}{\textit{F11R}},long={F11 receptor},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=F11R}{\textit{F11R}} (F11 receptor)}}
\DeclareAcronym{SDC4}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=SDC4}{\textit{SDC4}},long={syndecan 4},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=SDC4}{\textit{SDC4}} (syndecan 4)}}
\DeclareAcronym{FN1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=FN1}{\textit{FN1}},long={fibronectin 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=FN1}{\textit{FN1}} (fibronectin 1)}}
\DeclareAcronym{ITGB1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ITGB1}{\textit{ITGB1}},long={integrin subunit β 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ITGB1}{\textit{ITGB1}} (integrin subunit β 1)}}
\DeclareAcronym{ITGB1BP1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ITGB1BP1}{\textit{ITGB1BP1}},long={integrin subunit β 1 binding protein 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=ITGB1BP1}{\textit{ITGB1BP1}} (integrin subunit β 1 binding protein 1)}}
\DeclareAcronym{EREG}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=EREG}{\textit{EREG}},long={epiregulin},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=EREG}{\textit{EREG}} (epiregulin)}}
\DeclareAcronym{UNC93A}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=UNC93A}{\textit{UNC93A}},long={unc-93 homolog A (C. elegans)},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=UNC93A}{\textit{UNC93A}} (unc-93 homolog A (C. elegans))}}
\DeclareAcronym{UNC93B1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=UNC93B1}{\textit{UNC93B1}},long={unc-93 homolog B1 (C. elegans)},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=UNC93B1}{\textit{UNC93B1}} (unc-93 homolog B1 (C. elegans))}}
\DeclareAcronym{WNT5A}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=WNT5A}{\textit{WNT5A}},long={Wnt family member 5A},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=WNT5A}{\textit{WNT5A}} (Wnt family member 5A)}}
\DeclareAcronym{IL1RL1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=IL1RL1}{\textit{IL1RL1}},long={interleukin 1 receptor like 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=IL1RL1}{\textit{IL1RL1}} (interleukin 1 receptor like 1)}}
\DeclareAcronym{PLAC1}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PLAC1}{\textit{PLAC1}},long={placenta specific 1},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=PLAC1}{\textit{PLAC1}} (placenta specific 1)}}
\DeclareAcronym{Afp}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=Afp}{\textit{Afp}},long={α-fetoprotein},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=Afp}{\textit{Afp}} (α-fetoprotein)}}
\DeclareAcronym{Gjb2}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=Gjb2}{\textit{Gjb2}},long={gap junction protein β 2},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=Gjb2}{\textit{Gjb2}} (gap junction protein β 2)}}
\DeclareAcronym{CD9}{short=\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CD9}{\textit{CD9}},long={CD9 molecule},first-style=reversed,single={\href{http://www.genenames.org/cgi-bin/gene_symbol_report?match=CD9}{\textit{CD9}} (CD9 molecule)}}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "manuscript.Rnw"
%%% End: